### 2023-06-21 procesamiento/prod_analysis.Rmd
update1

---
title: "Análisis de datos"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output: 
  html_document: 
    toc: yes
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: true
editor_options: 
  chunk_output_type: console
---

# Setup

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(scipen=9999) # desactivar notacion cientifica
```

```{r}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,ggplot2)
load(here::here("input/data/proc/study1.RData"))

df1$country2 <- sjlabelled::to_character(df1$country) 
percentage_data <- df1 %>%
  group_by(country2, class5) %>%
  summarise(count = n()) %>%
  group_by(country2) %>%
  na.omit() %>% 
  mutate(percentage = count / sum(count)) %>% 
  arrange(country2,class5,percentage)

percentage_data$class5 <- forcats::fct_rev(percentage_data$class5)

order <- 
percentage_data %>% 
  filter(class5=="Non-skilled manual workers (VIIa+b)") %>% 
  arrange(percentage) %>% 
  select(country2)

order$orden <- 1:30
ggplot(percentage_data, aes(x=country2,y=percentage,fill=class5,group=class5))+
  geom_bar(stat = 'identity')+
  # geom_text(aes(label=number_of_errors),position = position_stack(0.5))+
  scale_y_continuous(labels = NULL) +
  scale_x_discrete(limits=order$country2) +  
  scale_fill_ordinal() +
  theme_classic() +
  theme(legend.position = 'bottom')  +
  coord_flip()
```

```{r}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,ggplot2)
load(here::here("input/data/proc/study1.RData"))

df1$country2 <- sjlabelled::to_character(df1$country) 
percentage_data <- df1 %>%
  group_by(country2, class3) %>%
  summarise(count = n()) %>%
  group_by(country2) %>%
  na.omit() %>% 
  mutate(percentage = count / sum(count)) %>% 
  arrange(country2,class3,percentage)

percentage_data$class3 <- forcats::fct_rev(percentage_data$class3)

order <- 
percentage_data %>% 
  filter(class3=="Intermediate class (III+IV)") %>% 
  arrange(percentage) %>% 
  select(country2)

order$orden <- 1:30
ggplot(percentage_data, aes(x=country2,y=percentage,fill=class3,group=class3))+
  geom_bar(stat = 'identity')+
  # geom_text(aes(label=number_of_errors),position = position_stack(0.5))+
  scale_y_continuous(labels = NULL) +
  scale_x_discrete(limits=order$country2) +  
  scale_fill_ordinal() +
  theme_classic() +
  theme(legend.position = 'bottom')  +
  coord_flip()
```