### 2023-07-26 procesamiento/prod_analysis.Rmd
modelos ok

---
title: "Análisis de datos"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output: 
  html_document: 
    toc: yes
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: true
editor_options: 
  chunk_output_type: console
---

# Setup

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(scipen=9999) # desactivar notacion cientifica
```

# Descriptive

```{r ineq-attitudes}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,ggplot2)
load(here::here("input/data/proc/study1.RData"))
  # Declare ggplot2 features
ggplot2::theme_set(
  ggplot2::theme(
    panel.background = ggplot2::element_rect(fill = "gray85",
                                             colour = "gray85"),
    panel.border = ggplot2::element_blank(),
    axis.text.y = ggplot2::element_text(size = 13,
                                        hjust = 1),
    title = ggplot2::element_text(size = 13,
                                  face = "bold"),
    legend.text = ggplot2::element_text(size = 12),
    plot.caption = ggplot2::element_text(size = 10,
                                         face = "plain",
                                         hjust = 1)
  )
)

df1 %>%
    dplyr::select(perineq,prefineq,pertransfer,redist) %>%
    na.omit() %>% 
    sjPlot::plot_likert(geom.colors = "PuBu",
                        grid.range  =  c (1.2 , 1),
                        # title = "a. Perception meritocratic",
                          geom.size = c(0.62),
                          catcount = 4,
                          expand.grid = T,
                          values  =  "sum.outside",
                          reverse.colors = T,
                          reverse.scale = T,
                          show.n = FALSE,cat.neutral = 3,
                        show.prc.sign = T,
                        title = '') +
    ggplot2::theme(legend.position = "bottom")

ggsave(plot = last_plot(),filename = "ineq-attitudes.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)
```

```{r welfare}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,ggplot2,scales)
load(here::here("input/data/proc/study1.RData"))
df1 %>%
    dplyr::select(whealth,wcarelder) %>%
    na.omit() %>% 
    sjPlot::plot_likert(geom.colors = "PuBu",
                        grid.range  =  c (1 , 1),
                        # title = "a. Perception meritocratic",
                          geom.size = c(0.62),
                          catcount = 5,
                          expand.grid = T,
                          # values  =  "sum.outside",
                          reverse.colors = T,
                          reverse.scale = T,
                          show.n = FALSE,show.prc.sign = T,
                        title = 'Who do you think should primarily provide...') +
    scale_colour_discrete(labels = function(x) str_wrap(x, width = 10))+
    ggplot2::theme(legend.position = "bottom")

ggsave(plot = last_plot(),filename = "welf-attitudes.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 16)
```

```{r perc-integration}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,ggplot2,scales)
load(here::here("input/data/proc/study1.RData"))
df1 %>%
    dplyr::select(starts_with("pi_")) %>%
    na.omit() %>% 
    sjPlot::plot_likert(geom.colors = "PuBu",
                        grid.range  =  c (0.2 , 1),
                        # title = "a. Perception meritocratic",
                          geom.size = c(0.62),
                          catcount = 5,
                          expand.grid = T,
                          # values  =  "sum.outside",
                          reverse.colors = T,
                          reverse.scale = T,
                          show.n = FALSE,show.prc.sign = T,sort.frq = "pos.asc",
                        title = 'How often in the past 4 weeks have you felt that...') +
    scale_colour_discrete(labels = function(x) str_wrap(x, width = 10))+ 
    ggplot2::theme(legend.position = "bottom")

ggsave(plot = last_plot(),filename = "perc-integration.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)

df1 %>%
    dplyr::select(starts_with("pi_")) %>% 
  psych::alpha()
```

```{r}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,ggplot2,scales)
load(here::here("input/data/proc/study1.RData"))
df1$pi_nteg<- rowMeans(df1[,c("pi_compa","pi_isola","pi_left")],na.rm = T)
dfcor <- 
df1 %>%
    dplyr::select(pi_nteg,class6,sstatus,redist) %>% 
  as.data.frame()

levels(df1$class6)
table(df1$know_total)
table(df1$n_high)
table(df1$n_middle)
table(df1$n_low)

# dfcor$hhdepriv <- as.factor(dfcor$hhdepriv)
# dfcor$sstatus <- as.numeric(dfcor$sstatus)
cormat<- polycor::hetcor(dfcor,parallel = T,thresholds = F)

jpeg(file=here::here("output/network-structure.jpeg"),units = "cm",width = 17,height = 15,res = 300)
# Set up the layout for the multi-panel plot
par(mfrow = c(1, 1))
# Create the histograms
# hist(df1$prestige, main = sjlabelled::get_label(df1$prestige),xlab = NULL)
# hist(df1$diversity, main = sjlabelled::get_label(df1$diversity),xlab = NULL)
hist(df1$homclass, main = sjlabelled::get_label(df1$homclass),xlab = NULL)
dev.off()

sjPlot::plot_frq(df1$class3,show.n = F) +
  theme_grey()
ggsave(plot = last_plot(),filename = "class6.png",device = "png",
       path = here::here("output"),width = 1.5,height = 1,units = "cm",scale = 10)

jpeg(file=here::here("output/n_known.jpeg"),units = "cm",width = 15,height = 15,res = 300)
# Set up the layout for the multi-panel plot
par(mfrow = c(2, 2))

# Create the histograms
hist(df1$n_low, main = sjlabelled::get_label(df1$n_low),xlab = NULL)
hist(df1$n_middle, main = sjlabelled::get_label(df1$n_middle),xlab = NULL)
hist(df1$n_high, main = sjlabelled::get_label(df1$n_high),xlab = NULL)
hist(df1$know_total, main = sjlabelled::get_label(df1$know_total),xlab = NULL)
# hist(df1$prestige, main = sjlabelled::get_label(df1$prestige),xlab = NULL)
# hist(df1$diversity, main = sjlabelled::get_label(df1$diversity),xlab = NULL)
dev.off()

sjPlot::plot_grpfrq(var.cnt = df1$know_total,var.grp = df1$class3,type = 'box',geom.colors = "Blues")
ggsave(plot = last_plot(),filename = "known_total_class6.png",device = "png",
       path = here::here("output"),width = 1,height = 1,units = "cm",scale = 15)

sjPlot::plot_grpfrq(var.cnt = df1$diversity,var.grp = df1$class3,type = 'box',geom.colors = "Blues")
ggsave(plot = last_plot(),filename = "diversity_class6.png",device = "png",
       path = here::here("output"),width = 1,height = 1,units = "cm",scale = 15)

sjPlot::plot_grpfrq(var.cnt = df1$prestige,var.grp = df1$class3,type = 'box',geom.colors = "Blues")
ggsave(plot = last_plot(),filename = "prestige_class6.png",device = "png",
       path = here::here("output"),width = 1,height = 1,units = "cm",scale = 15)

sjPlot::plot_grpfrq(var.cnt = df1$homclass,var.grp = df1$class3,type = 'box',geom.colors = "Blues")
ggsave(plot = last_plot(),filename = "homclass_class6.png",device = "png",
       path = here::here("output"),width = 1,height = 1,units = "cm",scale = 15)

sjlabelled::set_label(df1$pi_nteg) <- "Perceived social isolation"
sjPlot::plot_grpfrq(var.cnt = df1$pi_nteg,var.grp = df1$class3,type = 'box',geom.colors = "Blues",ylim = c(1,5))
ggsave(plot = last_plot(),filename = "pint_class3.png",device = "png",
       path = here::here("output"),width = 1,height = 1,units = "cm",scale = 15)
```

# Descriptive x country

```{r redistribution}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,ggplot2)
load(here::here("input/data/proc/study1.RData"))
df1$country<- as.numeric(df1$country)
df1$country2 <- countrycode::countrycode(sourcevar = df1$country,origin = "iso3n",destination = "country.name")
df1$redistF <- factor(df1$redist,levels = 1:5,labels = sjlabelled::get_labels(df1$redist))
percentage_data <- df1 %>%
  group_by(country2, redistF) %>%
  summarise(count = n()) %>%
  group_by(country2) %>%
  na.omit() %>% 
  mutate(percentage = count / sum(count)) %>% 
  arrange(country2,redistF,percentage) %>% 
  mutate(pct_lab = paste0(round(percentage*100,1),"%")) %>% 
  mutate(pct_lab=ifelse(!redistF %in% c("Strongly agree"),NA,pct_lab))

order <- 
percentage_data %>% 
  filter(redistF=="Strongly agree") %>% 
  arrange(percentage) %>% 
  select(country2)

order$orden <- 1:31
ggplot(percentage_data, aes(x=country2,y=percentage*100,fill=redistF,group=redistF,label = pct_lab))+
  geom_bar(stat = 'identity') +
  # geom_text(position = position_fill(vjust =100)) +
  scale_y_continuous(name=NULL,labels = scales::percent_format(scale = 1),limits = c(0,101), expand = c(0, 0)) +
  scale_x_discrete(limits=order$country2,name=NULL) +  
  scale_fill_ordinal(guide = guide_legend(reverse = TRUE)) +
  theme_classic() +
  ggtitle(label= "Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes.") + 
  theme(legend.position = 'bottom',legend.title = element_blank())  +
  coord_flip()

ggsave(plot = last_plot(),filename = "wredist_cn.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)

df1$pi_nteg<- rowMeans(df1[,c("pi_compa","pi_isola","pi_left")],na.rm = T)
resits_count <- df1 %>%
  group_by(country,country2) %>%
  summarise(redist = mean(redist,na.rm = T),
            homclass = mean(homclass,na.rm = T),
            pi_nteg = mean(pi_nteg,na.rm = T),
            pi_ntegm = median(pi_nteg,na.rm = T)
            ) %>%
  na.omit()

cor(df1[,c("redist","homclass","pi_nteg")],use = "complete.obs")

ggplot(data = resits_count,aes(x = homclass,y = redist,label=country2)) +
  xlab("Class-based homogeneity") +
  ylab("Redistributive preferences")+  
  labs(alt = "A",caption = "r=0.42") +  
  geom_smooth(method = "lm") +
  geom_text()

myorder <-  resits_count$country2[order(resits_count$homclass)]
ggplot(data = resits_count,aes(x =country2,y = homclass)) + 
  geom_segment(aes(x=country2, xend=country2, y=0, yend=homclass), color="black") +
  geom_point() +
  scale_x_discrete(limits=myorder) +
  ylab("Class-based homogeneity")+
  xlab("") +
  theme(axis.text.x = element_text(angle = 45))

ggsave(plot = last_plot(),filename = "homclass_cn.png",device = "png",
       path = here::here("output"),width = 1,height = 1,units = "cm",scale = 20)

myorder <-  resits_count$country2[order(resits_count$pi_nteg)]
ggplot(data = resits_count,aes(x =country2,y = pi_nteg)) + 
  geom_segment(aes(x=country2, xend=country2, y=0, yend=pi_nteg), color="black") +
  geom_point() +
  scale_x_discrete(limits=myorder) +
  ylab("Perceived isolation")+
  xlab("") +
  theme(axis.text.x = element_text(angle = 45))

ggsave(plot = last_plot(),filename = "p_isolat_cn.png",device = "png",
       path = here::here("output"),width = 1,height = 1,units = "cm",scale = 20)
```

```{r health}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,ggplot2)
load(here::here("input/data/proc/study1.RData"))
df1$country2 <- countrycode::countrycode(sourcevar = df1$country,origin = "iso3n",destination = "country.name")

df1$whealthF <- factor(df1$whealth,levels = 1:5,labels = sjlabelled::get_labels(df1$whealth))

percentage_data <- df1 %>%
  group_by(country2, whealthF) %>%
  summarise(count = n()) %>%
  group_by(country2) %>%
  na.omit() %>% 
  mutate(percentage = count / sum(count)) %>% 
  arrange(country2,whealthF,percentage) %>% 
  mutate(pct_lab = paste0(round(percentage*100,1),"%")) %>% 
  mutate(pct_lab=ifelse(!whealthF %in% c("Government"),NA,pct_lab))

order <- 
percentage_data %>% 
  filter(whealthF=="Government") %>% 
  arrange(percentage) %>% 
  select(country2)

order$orden <- 1:32
ggplot(percentage_data, aes(x=country2,y=percentage*100,fill=whealthF,group=whealthF,label = pct_lab))+
  geom_bar(stat = 'identity') +
  geom_text(position = position_fill(vjust = 90)) + 
  scale_y_continuous(name=NULL,labels = scales::percent_format(scale = 1),limits = c(0,101), expand = c(0, 0)) +
  scale_x_discrete(limits=order$country2,name=NULL) +  
  scale_fill_ordinal(guide = guide_legend(reverse = TRUE)) +
  theme_classic() +
  ggtitle(label= "Who do you think should primarily provide: health") + 
  theme(legend.position = 'bottom',legend.title = element_blank())  +
  coord_flip()

ggsave(plot = last_plot(),filename = "whealth_cn.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)
```

```{r elderlycare}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,ggplot2)
load(here::here("input/data/proc/study1.RData"))
df1$country2 <- countrycode::countrycode(sourcevar = df1$country,origin = "iso3n",destination = "country.name")
df1$wcarelderF <- factor(df1$wcarelder,levels = 1:5,labels = sjlabelled::get_labels(df1$wcarelder))
percentage_data <- df1 %>%
  group_by(country2, wcarelderF) %>%
  summarise(count = n()) %>%
  group_by(country2) %>%
  na.omit() %>% 
  mutate(percentage = count / sum(count)) %>% 
  arrange(country2,wcarelderF,percentage) %>% 
  mutate(pct_lab = paste0(round(percentage*100,1),"%")) %>% 
  mutate(pct_lab=ifelse(!wcarelderF %in% c("Government"),NA,pct_lab))

order <- 
percentage_data %>% 
  filter(wcarelderF=="Government") %>% 
  arrange(percentage) %>% 
  select(country2)

order$orden <- 1:32
ggplot(percentage_data, aes(x=country2,y=percentage*100,fill=wcarelderF,group=wcarelderF,label = pct_lab))+
  geom_bar(stat = 'identity') +
  geom_text(position = position_fill(vjust = 95)) + 
  scale_y_continuous(name=NULL,labels = scales::percent_format(scale = 1),limits = c(0,101), expand = c(0, 0)) +
  scale_x_discrete(limits=order$country2,name=NULL) +  
  scale_fill_ordinal(guide = guide_legend(reverse = TRUE)) +
  theme_classic() +
  ggtitle(label= "Who do you think should primarily provide: elderly care") + 
  theme(legend.position = 'bottom',legend.title = element_blank())  +
  coord_flip()

ggsave(plot = last_plot(),filename = "wcareelder_cn.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)
```

```{r class5}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,ggplot2)
load(here::here("input/data/proc/study1.RData"))
df1$country2 <- countrycode::countrycode(sourcevar = df1$country,origin = "iso3n",destination = "country.name")
percentage_data <- df1 %>%
  group_by(country2, class5) %>%
  summarise(count = n()) %>%
  group_by(country2) %>%
  na.omit() %>% 
  mutate(percentage = count / sum(count)) %>% 
  arrange(country2,class5,percentage)

percentage_data$class5 <- forcats::fct_rev(percentage_data$class5)

order <- 
percentage_data %>% 
  filter(class5=="Service class (I+II)") %>% 
  arrange(percentage) %>% 
  select(country2)

order$orden <- 1:32
ggplot(percentage_data, aes(x=country2,y=percentage*100,fill=class5,group=class5))+
  geom_bar(stat = 'identity')+
  scale_y_continuous(name=NULL,labels = scales::percent_format(scale = 1),limits = c(0,101), expand = c(0, 0)) +
  scale_x_discrete(limits=order$country2,name=NULL) +  
  scale_fill_ordinal(guide = guide_legend(reverse = TRUE)) +
  theme_classic() +
  ggtitle(label= "Social class by country") + 
  theme(legend.position = 'bottom',legend.title = element_blank())  +
  coord_flip()

ggsave(plot = last_plot(),filename = "class5_cn.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)
```

```{r class3}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,ggplot2)
load(here::here("input/data/proc/study1.RData"))

df1$country2 <- sjlabelled::to_character(df1$country) 
percentage_data <- df1 %>%
  group_by(country2, class3) %>%
  summarise(count = n()) %>%
  group_by(country2) %>%
  na.omit() %>% 
  mutate(percentage = count / sum(count)) %>% 
  arrange(country2,class3,percentage)

percentage_data$class3 <- forcats::fct_rev(percentage_data$class3)

order <- 
percentage_data %>% 
  filter(class3=="Working Class (V+VI+VII)") %>% 
  arrange(percentage) %>% 
  select(country2)

order$orden <- 1:32
ggplot(percentage_data, aes(x=country2,y=percentage,fill=class3,group=class3))+
  geom_bar(stat = 'identity')+
  # geom_text(aes(label=number_of_errors),position = position_stack(0.5))+
  scale_y_continuous(labels = NULL) +
  scale_x_discrete(limits=order$country2) +  
  scale_fill_ordinal() +
  theme_classic() +
  theme(legend.position = 'bottom')  +
  coord_flip()
```

```{r correlation-socap}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,ggplot2)
load(here::here("input/data/proc/study1.RData"))
# df1$class6 <- forcats::fct_rev(factor(df1$class6,ordered = T))
cormat1 <- 
df1 %>% 
    select('redist',wcarelder,whealth,isei_tot,isei_max,isei_range,"class6") %>% 
  na.omit() %>% 
  polycor::hetcor(parallel = T)

  corrplot::corrplot(cormat1$correlations,
    method = "color",
    addCoef.col = "#000390",
    type = "upper",
    tl.col = "black",
    col=colorRampPalette(c("white","#0068DC"))(12),
    bg = "white",
    na.label = "-")
  text(4.7, 0.6, paste0("Source: Authors calculation based on ISSP 2017",
                             " (N = ",")"),   
       cex = 0.75)
```

```{r correlation-segregation}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,ggplot2)
load(here::here("input/data/proc/study1.RData"))
# df1$class6 <- forcats::fct_rev(factor(df1$class6,ordered = T))
cormat1 <- 
df1 %>% 
    select('redist',homclass,homclass_s,homclass_w,pi_nteg,class3) %>% 
  na.omit() %>% 
  polycor::hetcor(parallel = T)

  corrplot::corrplot(cormat1$correlations,
    method = "color",
    addCoef.col = "#000390",
    type = "upper",
    tl.col = "black",
    col=colorRampPalette(c("white","#0068DC"))(12),
    bg = "white",
    na.label = "-")
  text(4.7, 0.6, paste0("Source: Authors calculation based on ISSP 2017",
                             " (N = ",")"),   
       cex = 0.75)
```

```{r}
df1$country2 <- sjlabelled::to_character(df1$country)
library(scales)

df1 %>% 
  filter(!is.na(class5)) %>%
ggplot(aes(x = class5, y = isei_avg))+
  geom_boxplot(outlier.shape = NA)+
  stat_summary(fun.y=mean, geom="point", shape=20, size=2, color="red", fill="red")+
  ylim(25, 75) +
  facet_wrap(.~country2, scales = "free")+ 
  scale_x_discrete(labels=c("I+II","IIIa+b","IVa+b+c","V+VI","VIIa+b"))+
  ggtitle(sjlabelled::get_label(df1$isei_avg))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

df1 %>% 
  # filter(!is.na(class5)) %>% 
ggplot(aes(x = class5, y = hindex_10))+
  geom_boxplot(outlier.shape = NA)+
  stat_summary(fun.y=mean, geom="point", shape=20, size=2, color="red", fill="red")+
  ylim(0, 1) +
  facet_wrap(.~country2, scales = "free")+ 
  scale_x_discrete(labels=c("I+II","IIIa+b","IVa+b+c","V+VI","VIIa+b"))+
  ggtitle(sjlabelled::get_label(df1$hindex_10))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

df1 %>% 
  filter(!is.na(class6)) %>% 
ggplot(aes(x = class6, y = hindex_10))+
  geom_boxplot(outlier.shape = NA)+
  facet_wrap(.~country2, scales = "free")+ 
  scale_x_discrete(labels=c("I","II","IIIa+b","IVa+b+c","V+VI","VIIa+b"))+
  ggtitle(sjlabelled::get_label(df1$hindex_10))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

levels(df1$class5)
```

# Modelos

dicotomizar homogeneidad en 1=50% similar,49% similar

Modelos homogeneidad

```{r class-homclass}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,sjPlot)
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"))
df1 <- df2
df1$class3 <- relevel(df1$class3,ref =  "Intermediate class (III+IV)")
df1$class5 <- factor(df1$class5,levels = c("Petty-bourgeoise (IVa+b+c)",
                                           "Routine non-manual (IIIa+b)",
                                           "Skilled manual workers and supv. (V+VI)",
                                           "Non-skilled manual workers (VIIa+b)",
                                           "Service class (I+II)"
                                           ))
df1$class6 <- factor(df1$class6,levels = c("Petty-bourgeoise (IVa+b+c)",
                                           "Routine non-manual (IIIa+b)",
                                           "Skilled manual workers and supv. (V+VI)",
                                           "Non-skilled manual workers (VIIa+b)",
                                           "Service class (II - lower grade)",
                                           "Service class (I - higher grade)" 
                                           ))
df1$pi_nteg<- rowMeans(df1[,c("pi_compa","pi_isola","pi_left")],na.rm = T)
sjlabelled::set_label(df1$pi_nteg) <- "Perceived social isolation"
df1$Q05pcm <- car::recode(df1$Q05pc,"1='Q01';2='Q02';3='Q03';4='Q04';5='Q05';NA='Missing'",
                          as.factor = T,
                          levels = c("Q01","Q02","Q03","Q04","Q05","Missing")) 
frq(df1$Q05pcm)
dfreg <- df1 %>% dplyr::select(redist,
                        class6,class3,class5,
                        # tr_inte,
                        # eindex,
                        # know_total,
                        homclass,
                        homclass_s,homclass_w,
                        # prestige,
                        # diversity,
                        # pi_nteg,
                        # isei_range,
                        # hindex_sta,
                        # hindex_10,
                        Q05pcm,labst,union,
                        # sstatus,
                        gini_disp,
                        gini_mkt,
                        # palmawd,
                        rgdpna,
                        # giniwb,
                        # giniwd,
                        ednum,female,agenum,country) %>% 
  mutate(logrgdpna=log(rgdpna)) %>% 
  # filter(country!=233) %>% 
  na.omit()

dfreg$homclass_bi <- ifelse(dfreg$homclass>0.5,1,0)
frq(dfreg$homclass_bi)

base <- lmer(redist~1 + (1|country),data=dfreg);performance::icc(base)
demo <- lmer(redist~1 +female+agenum + labst + union + (1|country),data=dfreg)
homclass <- lmer(redist~1 + female+agenum + labst + union+  homclass + (1|country),data=dfreg)
class <- lmer(redist~1 +female+agenum + labst + union+ class3+ (1|country),data=dfreg)
class_status <- lmer(redist~1 +female+agenum + labst + union+ class3+ednum+ Q05pcm + (1|country),data=dfreg)
full <- lmer(redist~1 +female+agenum + labst + union+ class3 +ednum+homclass +  Q05pcm + (1|country),data=dfreg)
full_m <- lmer(redist~1 +female+agenum + labst + union+ class3 +ednum+homclass +  Q05pcm +  gini_disp + logrgdpna + (1|country),data=dfreg)

# Interactions segregation x class
full_int1 <- lmer(redist~1 +female+agenum + homclass*class3+ ednum + Q05pcm + (1|country),data=dfreg)
# full_int3 <- lmer(redist~1 +female+agenum + labst + union+ homclass+pi_nteg*class3+ pi_nteg+ ednum + Q05pcm + tr_inte + (1|country),data=dfreg)

models <- list(homclass,class,class_status,full,full_int1)
texreg::knitreg(models)

sjPlot::plot_model(full,show.values = T,rm.terms = c("ednum","femaleFemale","agenum")) + geom_hline(yintercept = 0)
ggsave(plot = last_plot(),filename = "model_full.png",device = "png",
       path = here::here("output"),width = 1.5,height = 1,units = "cm",scale = 15)

# Conditional Marginal effect
interplot <- 
interplot::interplot(m = full_int1,var1 = "homclass",var2="class3")
df_int <- interplot$data
df_int <- df_int[c(1,2,4),]
df_int$value <- c("Intermediate class (III+IV)","Service Class (I+II)","Working Class (V+VI+VII)")
df_int
df_int$value <- factor(df_int$value,ordered = T,
                       levels = c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"))

df_int %>% 
ggplot() +
  geom_point(aes(x = value,y = coef1)) + 
  geom_errorbar(aes(x = value, ymin = lb, ymax = ub),color="black",width=0.1,size=1) +
  labs(caption = paste0("N=",nobs(full),", Countries=",lme4::ngrps(full))) +
  ylab("Class-based homogeneity coefficient")+
  xlab("Social Class")+
  ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
  geom_hline(yintercept = 0) +
  labs(caption = paste0("N=",nobs(full),", Countries=",lme4::ngrps(full))) +  
  theme(plot.title = element_text(size = 8))

ggsave(plot = last_plot(),filename = "mar_homo_class.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 10)

# Predicted values 
pred1<- plot_model(full_int1,type = "int")
pred_data <- data.frame(pred1$data)
pred_data <- pred_data %>% arrange(pred_data$group)
pred_data$group <- factor(pred_data$group,ordered = T,
                       levels = c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"))

pred_data %>% 
ggplot() +
  geom_point(aes(x = x,y = predicted)) +
  geom_line(aes(x = x,y = predicted))+
  geom_errorbar(aes(x = x, ymin = conf.low, ymax = conf.high),color="black",width=0.05,size=1) +
  facet_grid(. ~ group) +
  labs(caption = paste0("N=",nobs(full),", Countries=",lme4::ngrps(full))) +
  xlab("Network homogeneity")+
  ylab("Redistributive preferences") + 
  theme_minimal()

ggsave(plot = last_plot(),filename = "pred_homo_class.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 10)

# isolation <- 
# interplot::interplot(m = full_int3,var1 = "pi_nteg",var2="class3")
# df_int2 <- isolation$data
# df_int2 <- df_int2[c(1,2,4),]
# df_int2$value <- c("Intermediate class (III+IV)","Service Class (I+II)","Working Class (V+VI+VII)")
# df_int2
# df_int2$value <- factor(df_int2$value,ordered = T,
#                         levels = c("Working Class (V+VI+VII)",
#                                    "Intermediate class (III+IV)",
#                                    "Service Class (I+II)"))
# 
# df_int2 %>% 
# ggplot() +
#   geom_point(aes(x = value,y = coef1)) + 
#   geom_errorbar(aes(x = value, ymin = lb, ymax = ub),color="black",width=0.1,size=1) +
#   ylab("Perceived isolation coefficient")+
#   xlab("Social Class")+
#   ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
#   geom_hline(yintercept = 0)
# 
# ggsave(plot = last_plot(),filename = "margins_isolationxclass.png",device = "png",
#        path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)

```

```{r homclass-macro}
# Interactions segregation x inequality
full_int4 <- lmer(redist~1 +female+agenum + labst + union+ class3 + homclass*gini_disp+ednum + Q05pcm  +logrgdpna + (1+homclass|country),data=dfreg)
# full_int5 <- lmer(redist~1 +female+agenum + labst + union+ class3 + homclass+pi_nteg*gini_disp+pi_nteg+ednum + Q05pcm + tr_inte +logrgdpna + (1+pi_nteg|country),data=dfreg)
models_m <- list(full_int4)
texreg::knitreg(models_m)

interplot::interplot(m = full_int4,var1 = "homclass",var2="gini_disp") +
  geom_point()+
  ylab("Class-based homogeneity coefficient")+
  xlab("Economic Inequality")+
  ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
  geom_hline(yintercept = 0) +
  labs(caption = paste0("N=",nobs(full_int4),", Countries=",lme4::ngrps(full_int4))) +  
  theme(plot.title = element_text(size = 8))

ggsave(plot = last_plot(),filename = "mar_homclass_gini.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 10)

# interplot::interplot(m = full_int5,var1 = "pi_nteg",var2="gini_disp") +
#   geom_point()+
#   ylab("Perceived isolation coefficient")+
#   xlab("Economic Inequality")+
#   ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
#   geom_hline(yintercept = 0)
# ggsave(plot = last_plot(),filename = "margins_isolationxgini.png",device = "png",
#        path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)
```

Modelos strong ties

```{r strong-ties}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,ggplot2,interplot)
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"))
df1 <- df2
df1$class3 <- relevel(df1$class3,ref =  "Intermediate class (III+IV)")
df1$class5 <- factor(df1$class5,levels = c("Petty-bourgeoise (IVa+b+c)",
                                           "Routine non-manual (IIIa+b)",
                                           "Skilled manual workers and supv. (V+VI)",
                                           "Non-skilled manual workers (VIIa+b)",
                                           "Service class (I+II)"
                                           ))
df1$class6 <- factor(df1$class6,levels = c("Petty-bourgeoise (IVa+b+c)",
                                           "Routine non-manual (IIIa+b)",
                                           "Skilled manual workers and supv. (V+VI)",
                                           "Non-skilled manual workers (VIIa+b)",
                                           "Service class (II - lower grade)",
                                           "Service class (I - higher grade)" 
                                           ))
df1$pi_nteg<- rowMeans(df1[,c("pi_compa","pi_isola","pi_left")],na.rm = T)
sjlabelled::set_label(df1$pi_nteg) <- "Perceived social isolation"
df1$Q05pcm <- car::recode(df1$Q05pc,"1='Q01';2='Q02';3='Q03';4='Q04';5='Q05';NA='Missing'",
                          as.factor = T,
                          levels = c("Q01","Q02","Q03","Q04","Q05","Missing")) 
frq(df1$Q05pcm)
dfreg <- df1 %>% dplyr::select(redist,
                        class6,class3,class5,
                        # tr_inte,
                        # eindex,
                        # know_total,
                        homclass,
                        homclass_s,
                        homclass_w,
                        # prestige,
                        # diversity,
                        # pi_nteg,
                        # isei_range,
                        # hindex_sta,
                        # hindex_10,
                        Q05pcm,labst,union,
                        # sstatus,
                        gini_disp,
                        gini_mkt,
                        # palmawd,
                        rgdpna,
                        # giniwb,
                        # giniwd,
                        ednum,female,agenum,country) %>% 
  mutate(logrgdpna=log(rgdpna)) %>% 
  # filter(country!=233) %>% 
  na.omit()
dim(dfreg)
base <- lmer(redist~1 + (1|country),data=dfreg);performance::icc(base)
demo <- lmer(redist~1 +female+agenum + labst + union + (1|country),data=dfreg)
homclass <- lmer(redist~1 + female+agenum + labst + union+  homclass_s + (1|country),data=dfreg)
class <- lmer(redist~1 +female+agenum + labst + union+ class3+ (1|country),data=dfreg)
class_status <- lmer(redist~1 +female+agenum + labst + union+ class3+ednum+ Q05pcm + (1|country),data=dfreg)
full <- lmer(redist~1 +female+agenum + labst + union+ class3 +ednum+homclass_s + Q05pcm + (1|country),data=dfreg)
full_m <- lmer(redist~1 +female+agenum + labst + union+ class3 +ednum+homclass_s + Q05pcm + gini_disp + logrgdpna + (1|country),data=dfreg)

# Interactions segregation x class
full_int1 <- lmer(redist~1 +female+agenum + labst + union+ homclass_s*class3+ ednum + Q05pcm + (1|country),data=dfreg)
# full_int3 <- lmer(redist~1 +female+agenum + labst + union+ homclass_s+pi_nteg*class3+ pi_nteg+ ednum + Q05pcm + tr_inte + (1|country),data=dfreg)

models <- list(homclass,class,class_status,full,full_m,full_int1)
texreg::knitreg(models)

# Conditional Marginal effect---------------------------------------------------
interplot <- 
interplot::interplot(m = full_int1,var1 = "homclass_s",var2="class3")
df_int <- interplot$data
df_int <- df_int[c(1,2,4),]
df_int$value <- c("Intermediate class (III+IV)","Service Class (I+II)","Working Class (V+VI+VII)")
df_int
df_int$value <- factor(df_int$value,ordered = T,
                       levels = c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"))

df_int %>% 
ggplot() +
  geom_point(aes(x = value,y = coef1)) + 
  geom_errorbar(aes(x = value, ymin = lb, ymax = ub),color="black",width=0.1,size=1) +
  ylab("Class-based homogeneity (s) coefficient")+
  xlab("Social Class")+
  ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
  geom_hline(yintercept = 0) +
  labs(caption = paste0("N=",nobs(full),", Countries=",lme4::ngrps(full))) +  
  theme(plot.title = element_text(size = 8))

ggsave(plot = last_plot(),filename = "mar_strong-ties_class.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 10)

# Predicted values -------------------------------------------------------------
pred1<- plot_model(full_int1,type = "int")
pred_data <- data.frame(pred1$data)
pred_data <- pred_data %>% arrange(pred_data$group)
pred_data$group <- factor(pred_data$group,ordered = T,
                       levels = c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"))

pred_data %>% 
ggplot() +
  geom_point(aes(x = x,y = predicted)) +
  geom_line(aes(x = x,y = predicted))+
  geom_errorbar(aes(x = x, ymin = conf.low, ymax = conf.high),color="black",width=0.05,size=1) +
  facet_grid(. ~ group) +
  labs(caption = paste0("N=",nobs(full_int1),", Countries=",lme4::ngrps(full_int1))) +
  xlab("Network homogeneity (strong ties)")+
  ylab("Redistributive preferences") + 
  theme_minimal()

ggsave(plot = last_plot(),filename = "pred_strong-ties_class.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 10)

# isolation <- 
# interplot::interplot(m = full_int3,var1 = "pi_nteg",var2="class3")
# df_int2 <- isolation$data
# df_int2 <- df_int2[c(1,2,4),]
# df_int2$value <- c("Intermediate class (III+IV)","Service Class (I+II)","Working Class (V+VI+VII)")
# df_int2
# df_int2$value <- factor(df_int2$value,ordered = T,
#                         levels = c("Working Class (V+VI+VII)",
#                                    "Intermediate class (III+IV)",
#                                    "Service Class (I+II)"))

# df_int2 %>% 
# ggplot() +
#   geom_point(aes(x = value,y = coef1)) + 
#   geom_errorbar(aes(x = value, ymin = lb, ymax = ub),color="black",width=0.1,size=1) +
#   ylab("Perceived isolation coefficient")+
#   xlab("Social Class")+
#   ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
#   geom_hline(yintercept = 0)

# ggsave(plot = last_plot(),filename = "margins_isolationxclass.png",device = "png",
#        path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)

```

```{r strong-ties-macro}
# Interactions segregation x inequality
full_int4 <- lmer(redist~1 +female+agenum + labst + union+ class3 + homclass_s*gini_disp+ednum + Q05pcm +logrgdpna + (1+homclass_s|country),data=dfreg)
# full_int5 <- lmer(redist~1 +female+agenum + labst + union+ class3 + homclass_s+pi_nteg*gini_disp+ednum + Q05pcm + tr_inte +logrgdpna + (1+pi_nteg|country),data=dfreg)

models_m <- list(full_int4)
texreg::knitreg(models_m)

interplot::interplot(m = full_int4,var1 = "homclass_s",var2="gini_disp") +
  geom_point()+
  ylab("Class-based homogeneity (s) coefficient")+
  xlab("Economic Inequality")+
  ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
  geom_hline(yintercept = 0) +
  labs(caption = paste0("N=",nobs(full_int4),", Countries=",lme4::ngrps(full_int4))) +  
  theme(plot.title = element_text(size = 8))

ggsave(plot = last_plot(),filename = "mar_strong-ties_gini.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 10)

# interplot::interplot(m = full_int5,var1 = "pi_nteg",var2="gini_disp") +
#   geom_point()+
#   ylab("Perceived isolation coefficient")+
#   xlab("Economic Inequality")+
#   ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
#   geom_hline(yintercept = 0)
# 
# ggsave(plot = last_plot(),filename = "margins_isolationxgini.png",device = "png",
#        path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)
```

Modelos weak ties

```{r weak-ties}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,ggplot2,interplot)
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"))
df1 <- df2
df1$class3 <- relevel(df1$class3,ref =  "Intermediate class (III+IV)")
df1$class5 <- factor(df1$class5,levels = c("Petty-bourgeoise (IVa+b+c)",
                                           "Routine non-manual (IIIa+b)",
                                           "Skilled manual workers and supv. (V+VI)",
                                           "Non-skilled manual workers (VIIa+b)",
                                           "Service class (I+II)"
                                           ))
df1$class6 <- factor(df1$class6,levels = c("Petty-bourgeoise (IVa+b+c)",
                                           "Routine non-manual (IIIa+b)",
                                           "Skilled manual workers and supv. (V+VI)",
                                           "Non-skilled manual workers (VIIa+b)",
                                           "Service class (II - lower grade)",
                                           "Service class (I - higher grade)" 
                                           ))
df1$pi_nteg<- rowMeans(df1[,c("pi_compa","pi_isola","pi_left")],na.rm = T)
sjlabelled::set_label(df1$pi_nteg) <- "Perceived social isolation"
df1$Q05pcm <- car::recode(df1$Q05pc,"1='Q01';2='Q02';3='Q03';4='Q04';5='Q05';NA='Missing'",
                          as.factor = T,
                          levels = c("Q01","Q02","Q03","Q04","Q05","Missing")) 
frq(df1$Q05pcm)
dfreg <- df1 %>% dplyr::select(redist,
                        class6,class3,class5,
                        # tr_inte,
                        # eindex,
                        # know_total,
                        homclass,
                        homclass_s,
                        homclass_w,
                        # prestige,
                        # diversity,
                        # pi_nteg,
                        # isei_range,
                        # hindex_sta,
                        # hindex_10,
                        Q05pcm,labst,union,
                        # sstatus,
                        gini_disp,
                        gini_mkt,
                        # palmawd,
                        rgdpna,
                        # giniwb,
                        # giniwd,
                        ednum,female,agenum,country) %>% 
  mutate(logrgdpna=log(rgdpna)) %>% 
  # filter(country!=233) %>% 
  na.omit()
dim(dfreg)
base <- lmer(redist~1 + (1|country),data=dfreg);performance::icc(base)
demo <- lmer(redist~1 +female+agenum + labst + union + (1|country),data=dfreg)
homclass <- lmer(redist~1 + female+agenum + labst + union+  homclass_w + (1|country),data=dfreg)
class <- lmer(redist~1 +female+agenum + labst + union+ class3+ (1|country),data=dfreg)
class_status <- lmer(redist~1 +female+agenum + labst + union+ class3+ednum+ Q05pcm + (1|country),data=dfreg)
full <- lmer(redist~1 +female+agenum + labst + union+ class3 +ednum+homclass_w + Q05pcm + (1|country),data=dfreg)
full_m <- lmer(redist~1 +female+agenum + labst + union+ class3 +ednum+homclass_w + Q05pcm + gini_disp + logrgdpna + (1|country),data=dfreg)

# Interactions segregation x class
full_int1 <- lmer(redist~1 +female+agenum + labst + union+ homclass_w*class3+ ednum + Q05pcm + (1|country),data=dfreg)
# full_int3 <- lmer(redist~1 +female+agenum + labst + union+ homclass_s+pi_nteg*class3+ pi_nteg+ ednum + Q05pcm + tr_inte + (1|country),data=dfreg)

models <- list(homclass,class,class_status,full,full_m,full_int1)
texreg::knitreg(models)

# Conditional Marginal effect---------------------------------------------------
interplot <- 
interplot::interplot(m = full_int1,var1 = "homclass_w",var2="class3")
df_int <- interplot$data
df_int <- df_int[c(1,2,4),]
df_int$value <- c("Intermediate class (III+IV)","Service Class (I+II)","Working Class (V+VI+VII)")
df_int
df_int$value <- factor(df_int$value,ordered = T,
                       levels = c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"))

df_int %>% 
ggplot() +
  geom_point(aes(x = value,y = coef1)) + 
  geom_errorbar(aes(x = value, ymin = lb, ymax = ub),color="black",width=0.1,size=1) +
  ylab("Class-based homogeneity (w) coefficient")+
  xlab("Social Class")+
  ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
  geom_hline(yintercept = 0) +
  labs(caption = paste0("N=",nobs(full),", Countries=",lme4::ngrps(full))) +  
  theme(plot.title = element_text(size = 8))

ggsave(plot = last_plot(),filename = "mar_weak-ties_class.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 10)

# Predicted values -------------------------------------------------------------
pred1<- plot_model(full_int1,type = "int")
pred_data <- data.frame(pred1$data)
pred_data <- pred_data %>% arrange(pred_data$group)
pred_data$group <- factor(pred_data$group,ordered = T,
                       levels = c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"))

pred_data %>% 
ggplot() +
  geom_point(aes(x = x,y = predicted)) +
  geom_line(aes(x = x,y = predicted))+
  geom_errorbar(aes(x = x, ymin = conf.low, ymax = conf.high),color="black",width=0.05,size=1) +
  facet_grid(. ~ group) +
  labs(caption = paste0("N=",nobs(full_int1),", Countries=",lme4::ngrps(full_int1))) +
  xlab("Network homogeneity (weak ties)")+
  ylab("Redistributive preferences") + 
  theme_minimal()

ggsave(plot = last_plot(),filename = "pred_weak-ties_class.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 10)

# ggsave(plot = last_plot(),filename = "margins_isolationxclass.png",device = "png",
#        path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)

```

```{r weak-ties-macro}
# Interactions segregation x inequality
full_int4 <- lmer(redist~1 +female+agenum + labst + union+ class3 + homclass_w*gini_disp+ednum + Q05pcm +logrgdpna + (1+homclass_w|country),data=dfreg)
# full_int5 <- lmer(redist~1 +female+agenum + labst + union+ class3 + homclass_s+pi_nteg*gini_disp+ednum + Q05pcm + tr_inte +logrgdpna + (1+pi_nteg|country),data=dfreg)

models_m <- list(full_int4)
texreg::knitreg(models_m)

interplot::interplot(m = full_int4,var1 = "homclass_w",var2="gini_disp") +
  geom_point()+
  ylab("Class-based homogeneity (w) coefficient")+
  xlab("Economic Inequality")+
  ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
  geom_hline(yintercept = 0) +
  labs(caption = paste0("N=",nobs(full_int4),", Countries=",lme4::ngrps(full_int4))) +  
  theme(plot.title = element_text(size = 8))

ggsave(plot = last_plot(),filename = "mar_weak-ties_gini.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 10)

# ggsave(plot = last_plot(),filename = "margins_isolationxgini.png",device = "png",
#        path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)
```

# Modelos alternativos

```{r isei}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(sjmisc,ggplot2,lme4,interplot)
load(here::here("input/data/proc/study1.RData"))
# remove.packages("Matrix")
# install.packages("Matrix")
df1$class3<-forcats::fct_rev(df1$class3)
df1$class5<-forcats::fct_rev(df1$class5)
df1$class6<-forcats::fct_rev(df1$class6)
dfreg <- df1 %>% dplyr::select(redist,
                        isei08,
                        eindex,
                        prestige,
                        diversity,
                        # isei_range,
                        # hindex_sta,
                        # hindex_10,
                        ednum,female,agenum,country) %>% na.omit()

dfreg$country <- to_factor(dfreg$country)
dim(dfreg)

base <- lmer(redist~1 + (1|country),data=dfreg)
demo <- lmer(redist~1 +female+agenum + (1|country),data=dfreg)
prestige <- lmer(redist~1 + female+agenum + prestige + (1|country),data=dfreg)
diversity <- lmer(redist~1 + female+agenum + prestige + diversity+ eindex + (1|country),data=dfreg)

class <- lmer(redist~1 +female+agenum + isei08+ (1|country),data=dfreg)
class_status <- lmer(redist~1 +female+agenum + isei08+ednum+ (1|country),data=dfreg)
full <- lmer(redist~1+female+agenum+isei08+ednum+prestige+diversity+eindex+(1|country),data=dfreg)
full_int1 <- lmer(redist~1+female+agenum+eindex*isei08+(1|country),data=dfreg)
full_int2 <- lmer(redist~1+female+agenum+prestige+diversity+eindex*isei08+ednum+(1+diversity|country),data=dfreg)
models <- list(prestige,diversity,class,class_status,full,full_int1,full_int2)
texreg::knitreg(models)

texreg::plotreg(full,omit.coef = "(Intercept)")

interplot::interplot(m = full_int1,var1 = "eindex",var2="isei08") +
  ylab("Network diversity coefficient")+
  xlab("SES Respondent")+
  ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
  geom_hline(yintercept = 0)

ggsave(plot = last_plot(),filename = "margins_diversityxisei.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)

#-----------
dfreg <- df1 %>% select(redist,
                        isei08,
                        eindex,
                        prestige,
                        # diversity,
                        # isei_range,
                        # hindex_sta,
                        # hindex_10,
                        ednum,female,agenum,country) %>% na.omit()

dim(dfreg)
base <- lmer(redist~1 + (1|country),data=dfreg)
demo <- lmer(redist~1 +female+agenum + (1|country),data=dfreg)
prestige <- lmer(redist~1 + female+agenum + prestige + (1|country),data=dfreg)
diversity <- lmer(redist~1 + female+agenum + prestige + eindex+ (1|country),data=dfreg)

class <- lmer(redist~1+female+agenum+isei08+(1|country),data=dfreg)
class_status <- lmer(redist~1+female+agenum+isei08+ednum+(1|country),data=dfreg)
full <- lmer(redist~1 +female+agenum + isei08 +ednum +  prestige + eindex +(1|country),data=dfreg)
fullrs <- lmer(redist~1 +female+agenum + isei08 +ednum +  prestige + eindex +(1+eindex|country),data=dfreg)
full_int1 <- lmer(redist~1+female+agenum+prestige+eindex*isei08+ednum+(1|country),data=dfreg)
full_int2 <- lmer(redist~1+female+agenum+prestige+eindex*isei08+ednum+(1+eindex|country),data=dfreg)
models <- list(prestige,diversity,class,class_status,full,full_int1,full_int2)
texreg::knitreg(models)

texreg::plotreg(full, omit.coef = "(Intercept)")

interplot::interplot(m = full_int2,var1 = "eindex",var2="isei08") +
  ylab("Network heterogeneity coefficient")+
  xlab("SES Respondent")+
  ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
  geom_hline(yintercept = 0)

ggsave(plot = last_plot(),filename = "margins_diversityxisei.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)
# en la medida que aumenta el status, el efecto de la diversidad de la red se vuelve positivo.
# en personas de estatus bajo la diversidad hace que demanden menos redistribucion
# en personas de estatus alto la diversidad hace que demanden más redistribucion
```

```{r class-diversity}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,ggplot2,lme4)
load(here::here("input/data/proc/study1.RData"))

levels(df1$class3)
df1$class3 <- relevel(df1$class3,ref =  "Intermediate class (III+IV)")
levels(df1$class3)
levels(df1$class5)
df1$class5 <- factor(df1$class5,levels = c("Petty-bourgeoise (IVa+b+c)",
                                           "Routine non-manual (IIIa+b)",
                                           "Skilled manual workers and supv. (V+VI)",
                                           "Non-skilled manual workers (VIIa+b)",
                                           "Service class (I+II)"
                                           ))

df1$class6 <- factor(df1$class6,levels = c("Petty-bourgeoise (IVa+b+c)",
                                           "Routine non-manual (IIIa+b)",
                                           "Skilled manual workers and supv. (V+VI)",
                                           "Non-skilled manual workers (VIIa+b)",
                                           "Service class (II - lower grade)",
                                           "Service class (I - higher grade)" 
                                           ))
levels(df1$class6)

dfreg <- df1 %>% select(redist,
                        class6,
                        class3,
                        class5,
                        # eindex,
                        prestige,
                        diversity,
                        # isei_range,
                        # hindex_sta,
                        # hindex_10,
                        ednum,female,agenum,country) %>% na.omit()
dim(dfreg)

base <- lmer(redist~1 + (1|country),data=dfreg)
demo <- lmer(redist~1 +female+agenum + (1|country),data=dfreg)
prestige <- lmer(redist~1 + female+agenum + prestige + (1|country),data=dfreg)
diversity <- lmer(redist~1 + female+agenum + prestige + diversity+ (1|country),data=dfreg)

class <- lmer(redist~1 +female+agenum + class6+ (1|country),data=dfreg)
class_status <- lmer(redist~1 +female+agenum + class6+ednum+ (1|country),data=dfreg)
full <- lmer(redist~1 +female+agenum + class6 +ednum +  prestige + diversity +(1|country),data=dfreg)
full_int1 <- lmer(redist~1 +female+agenum + prestige+diversity*class6+ ednum +(1|country),data=dfreg)
full_int2 <- lmer(redist~1 +female+agenum + prestige+diversity*class6+ ednum + (1+diversity|country),data=dfreg)
models <- list(prestige,diversity,class,class_status,full,full_int1,full_int2)
texreg::knitreg(models)

base <- lmer(redist~1 + (1|country),data=dfreg)
demo <- lmer(redist~1 +female+agenum + (1|country),data=dfreg)
prestige <- lmer(redist~1 + female+agenum + prestige + (1|country),data=dfreg)
diversity <- lmer(redist~1 + female+agenum + prestige + diversity+ (1|country),data=dfreg)

class <- lmer(redist~1 +female+agenum + class5+ (1|country),data=dfreg)
class_status <- lmer(redist~1 +female+agenum + class5+ednum+ (1|country),data=dfreg)
full <- lmer(redist~1 +female+agenum + class5 +ednum +  prestige + diversity +(1|country),data=dfreg)
full_int1 <- lmer(redist~1 +female+agenum + prestige+diversity*class5+ ednum +(1|country),data=dfreg)
full_int2 <- lmer(redist~1 +female+agenum + prestige+diversity*class5+ ednum + (1+diversity|country),data=dfreg)
models <- list(prestige,diversity,class,class_status,full,full_int1,full_int2)
texreg::knitreg(models)

base <- lmer(redist~1 + (1|country),data=dfreg)
demo <- lmer(redist~1 +female+agenum + (1|country),data=dfreg)
prestige <- lmer(redist~1 + female+agenum + prestige + (1|country),data=dfreg)
diversity <- lmer(redist~1 + female+agenum + prestige + diversity+ (1|country),data=dfreg)

class <- lmer(redist~1 +female+agenum + class3+ (1|country),data=dfreg)
class_status <- lmer(redist~1 +female+agenum + class3+ednum+ (1|country),data=dfreg)
full <- lmer(redist~1 +female+agenum + class3 +ednum +  prestige + diversity +(1|country),data=dfreg)
full_int1 <- lmer(redist~1 +female+agenum + prestige+diversity*class3+ ednum +(1|country),data=dfreg)
full_int2 <- lmer(redist~1 +female+agenum + prestige+diversity*class3+ ednum + (1+diversity|country),data=dfreg)
models <- list(prestige,diversity,class,class_status,full,full_int1,full_int2)
texreg::knitreg(models)

interplot::interplot(m = full_int1,var1 = "diversity",var2="class3") +
  ylab("Network diversity coefficient")+
  xlab("SES Respondent")+
  ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
  geom_hline(yintercept = 0)

```

```{r eindex}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,ggplot2,lme4)
load(here::here("input/data/proc/study1.RData"))

levels(df1$class3)
df1$class3 <- relevel(df1$class3,ref =  "Intermediate class (III+IV)")
levels(df1$class3)
levels(df1$class5)
df1$class5 <- factor(df1$class5,levels = c("Petty-bourgeoise (IVa+b+c)",
                                           "Routine non-manual (IIIa+b)",
                                           "Skilled manual workers and supv. (V+VI)",
                                           "Non-skilled manual workers (VIIa+b)",
                                           "Service class (I+II)"
                                           ))

df1$class6 <- factor(df1$class6,levels = c("Petty-bourgeoise (IVa+b+c)",
                                           "Routine non-manual (IIIa+b)",
                                           "Skilled manual workers and supv. (V+VI)",
                                           "Non-skilled manual workers (VIIa+b)",
                                           "Service class (II - lower grade)",
                                           "Service class (I - higher grade)" 
                                           ))
levels(df1$class6)

dfreg <- df1 %>% select(redist,
                        class6,
                        class3,
                        class5,
                        eindex,
                        homclass,
                        # homclass_s,homclass_w,
                        prestige,
                        diversity,
                        # isei_range,
                        # hindex_sta,
                        # hindex_10,
                        ednum,female,agenum,country) %>% na.omit()

dim(dfreg)
cor(dfreg[,c("eindex","prestige","diversity")])

base <- lmer(redist~1 + (1|country),data=dfreg)
demo <- lmer(redist~1 +female+agenum + (1|country),data=dfreg)
prestige <- lmer(redist~1 + female+agenum + prestige + (1|country),data=dfreg)
eindex <- lmer(redist~1 + female+agenum + prestige +diversity+eindex+ (1|country),data=dfreg)

class <- lmer(redist~1 +female+agenum + class6+ (1|country),data=dfreg)
class_status <- lmer(redist~1 +female+agenum + class6+ednum+ (1|country),data=dfreg)
full <- lmer(redist~1 +female+agenum + class6 +ednum + prestige +diversity+ eindex +(1|country),data=dfreg)
full_int1 <- lmer(redist~1 +female+agenum + prestige+diversity+eindex*class6+ ednum +(1|country),data=dfreg)
full_int2 <- lmer(redist~1 +female+agenum + prestige+diversity+eindex*class6+ ednum + (1+eindex|country),data=dfreg)
models <- list(prestige,eindex,class,class_status,full,full_int1,full_int2)
texreg::knitreg(models)

base <- lmer(redist~1 + (1|country),data=dfreg)
demo <- lmer(redist~1 +female+agenum + (1|country),data=dfreg)
prestige <- lmer(redist~1 + female+agenum + prestige + (1|country),data=dfreg)
eindex <- lmer(redist~1 + female+agenum + prestige + eindex+ (1|country),data=dfreg)

class <- lmer(redist~1 +female+agenum + class5+ (1|country),data=dfreg)
class_status <- lmer(redist~1 +female+agenum + class5+ednum+ (1|country),data=dfreg)
full <- lmer(redist~1 +female+agenum + class5 +ednum +  prestige + eindex +(1|country),data=dfreg)
full_int1 <- lmer(redist~1 +female+agenum + prestige+eindex*class5+ ednum +(1|country),data=dfreg)
full_int2 <- lmer(redist~1 +female+agenum + prestige+eindex*class5+ ednum + (1+eindex|country),data=dfreg)
models <- list(prestige,eindex,class,class_status,full,full_int1,full_int2)
texreg::knitreg(models)

base <- lmer(redist~1 + (1|country),data=dfreg)
demo <- lmer(redist~1 +female+agenum + (1|country),data=dfreg)
prestige <- lmer(redist~1 + female+agenum + prestige + (1|country),data=dfreg)
eindex <- lmer(redist~1 + female+agenum + prestige + eindex+ (1|country),data=dfreg)

class <- lmer(redist~1 +female+agenum + class3+ (1|country),data=dfreg)
class_status <- lmer(redist~1 +female+agenum + class3+ednum+ (1|country),data=dfreg)
full <- lmer(redist~1 +female+agenum + class3 +ednum +  prestige + eindex +(1|country),data=dfreg)
full_int1 <- lmer(redist~1 +female+agenum + prestige+eindex*class3+ ednum +(1|country),data=dfreg)
full_int2 <- lmer(redist~1 +female+agenum + prestige+eindex*class3+ ednum + (1+eindex|country),data=dfreg)
models <- list(prestige,eindex,class,class_status,full,full_int1,full_int2)
texreg::knitreg(models)

interplot <- 
interplot::interplot(m = full_int2,var1 = "eindex",var2="class3") +
  ylab("Network heterophily coefficient")+
  xlab("Social Class")+
  ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
  geom_hline(yintercept = 0)

df_int <- interplot$data
df_int <- df_int[c(1,2,4),]
df_int$value <- c("Intermediate class (III+IV)","Service Class (I+II)","Working Class (V+VI+VII)")
df_int
df_int$value <- factor(df_int$value,ordered = T,levels = c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"))

df_int %>% 
ggplot() +
  geom_point(aes(x = value,y = coef1)) + 
  geom_errorbar(aes(x = value, ymin = lb, ymax = ub),color="black",width=0.1,size=1) +
  ylab("Network heterophily coefficient")+
  xlab("Social Class")+
  ggtitle("Government's responsibility: reduce the differences in income between people with high incomes and those with low incomes") +
  geom_hline(yintercept = 0)

ggsave(plot = last_plot(),filename = "margins_hetxclass.png",device = "png",
       path = here::here("output"),width = 2,height = 1,units = "cm",scale = 15)
```