### 2023-10-09 procesamiento/prod_prep.Rmd

---
title: "Data preparation"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output: 
  html_document: 
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
      number_sections: yes
    code_folding: show  
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

# Setup

```{r setup}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE 
                      )
options(scipen=9999) # desactivar notacion cientifica
```

```{r}
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,GGally,sjlabelled,tidyverse,MLMusingR)
df1 <- haven::read_dta(here::here("input/data/original/ZA6980_v2-0-0.dta"))
dfTR <- haven::read_dta(here::here("input/data/original/ZA5521_v1-0-0.dta"))
dfES <- haven::read_dta(here::here("input/data/original/ZA5962_v1-0-0.dta"))

relig <- haven::read_dta(here::here("input/data/original/ZA7570_v2-1-0.dta"))
religTR <- relig %>% filter(country==792)

df1 <- df1 %>% rename(topbot=TOPBOT)
dfES <- dfES %>% rename(topbot=TOPBOT)
#fix problem with variable names
dfTR <- dfTR %>% rename(SEX=sex,
                        DEGREE=degree,
                        EDUCYRS=educyrs,
                        AGE=age,
                        MAINSTAT=mainstat,
                        UNION=union,
                        WEIGHT=weight,
                        URBRURAL=urbrural,
                        WORK=work,
                        HOMPOP=hompop)  
dfES$c_alphan <- NULL
dfTR$c_alphan <- NULL
df2 <- bind_rows(dfES,dfTR)
df1<- bind_rows(df1,df2)
```

# Welfare attitudes

```{r}
#frq(df1$v13)
df1$redist <- car::recode(df1$v13,recodes = "c(0,8,9)=NA")
df1$redist <- set_na(df1$redist,na = c(0,8,9))
df1$redist <- sjmisc::rec(x = df1$redist,rec = "rev")
sjlabelled::set_label(df1$redist) <- "Preferences goverment redistribution"
#frq(df1$redist)

#frq(df1$v15)
df1$whealth <- car::recode(df1$v15,recodes = "c(0,8,9)=NA")
df1$whealth <- set_na(df1$whealth,na = c(0,8,9))
df1$whealth <- sjmisc::rec(x = df1$whealth,rec = "rev")
sjlabelled::set_label(df1$whealth) <- "Welfare provision: health"
#frq(df1$whealth)

#frq(df1$v16)
df1$wcarelder <- car::recode(df1$v16,recodes = "c(0,8,9)=NA")
df1$wcarelder <- set_na(df1$wcarelder,na = c(0,8,9))
df1$wcarelder <- sjmisc::rec(x = df1$wcarelder,rec = "rev")
sjlabelled::set_label(df1$wcarelder) <- "Welfare provision: elderly care"
#frq(df1$wcarelder)
# cor(df1[,c("redist","whealth","wcarelder")],use = "complete.obs")
```

# Inequality perceptions

```{r}
#frq(df1$v59)
df1$hhdepriv <- car::recode(df1$v59,recodes = "c(0,8,9)=NA")
df1$hhdepriv <- set_na(df1$hhdepriv,na = c(0,8,9))
# df1$hhdepriv <- sjmisc::rec(x = df1$hhdepriv,rec = "rev")
sjlabelled::set_label(df1$hhdepriv) <- "Household economic deprivation (household difficulties)"
#frq(df1$hhdepriv)

#frq(df1$v11)
df1$perineq <- car::recode(df1$v11,recodes = "c(0,8,9)=NA")
df1$perineq <- set_na(df1$perineq,na = c(0,8,9))
df1$perineq <- sjmisc::rec(x = df1$perineq,rec = "rev")
sjlabelled::set_label(df1$perineq) <- "Perceived economic inequality (Inc. diff. too large)"
#frq(df1$perineq)

#frq(df1$v12)
df1$prefineq <- car::recode(df1$v12,recodes = "c(0,8,9)=NA")
df1$prefineq <- set_na(df1$prefineq,na = c(0,8,9))
df1$prefineq <- sjmisc::rec(x = df1$prefineq,rec = "rev")
sjlabelled::set_label(df1$prefineq) <- "Preference inequality (society has to be fair)"
#frq(df1$prefineq)

#frq(df1$v14)
df1$pertransfer <- car::recode(df1$v14,recodes = "c(0,8,9)=NA")
df1$pertransfer <- set_na(df1$pertransfer,na = c(0,8,9))
df1$pertransfer <- sjmisc::rec(x = df1$pertransfer,rec = "rev")
sjlabelled::set_label(df1$pertransfer) <- "Perception about welfare (people become lazy)"
#frq(df1$pertransfer)
```

# Perceived integration

```{r}
#frq(df1$v31)
df1$pi_compa <- car::recode(df1$v31,recodes = "c(0,8,9)=NA")
df1$pi_compa <- set_na(df1$pi_compa,na = c(0,8,9))
# df1$pi_compa <- sjmisc::rec(x = df1$pi_compa,rec = "rev")
sjlabelled::set_label(df1$pi_compa) <- "Perceived integration: companionship lacking"
#frq(df1$pi_compa)

#frq(df1$v32)
df1$pi_isola <- car::recode(df1$v32,recodes = "c(0,8,9)=NA")
df1$pi_isola <- set_na(df1$pi_isola,na = c(0,8,9))
# df1$pi_isola <- sjmisc::rec(x = df1$pi_isola,rec = "rev")
sjlabelled::set_label(df1$pi_isola) <- "Perceived integration: isolated from others"
#frq(df1$pi_compa)

#frq(df1$v33)
df1$pi_left <- car::recode(df1$v33,recodes = "c(0,8,9)=NA")
df1$pi_left <- set_na(df1$pi_left,na = c(0,8,9))
# df1$pi_left <- sjmisc::rec(x = df1$pi_left,rec = "rev")
sjlabelled::set_label(df1$pi_left) <- "Perceived integration: left out from activities"
#frq(df1$pi_left)

df1$isolation<- rowMeans(df1[,c("pi_compa","pi_isola","pi_left")],na.rm = T)
sjlabelled::set_label(df1$isolation) <- "Perceived isolation"
```

# Trust

```{r}
df1$tr_inte <- car::recode(df1$v35,recodes = "c(0,8,9)=NA")
df1$tr_inte <- set_na(df1$tr_inte,na = c(0,8,9))
df1$tr_inte <- sjmisc::rec(x = df1$tr_inte,rec = "rev")
sjlabelled::set_label(df1$tr_inte) <- "Interpersonal trust"
frq(df1$tr_inte)

df1$tr_courts <- car::recode(df1$v36,recodes = "c(98,99)=NA")
df1$tr_courts <- set_na(df1$tr_courts,na = c(98,99))
# df1$tr_courts <- sjmisc::rec(x = df1$tr_courts,rec = "rev")
sjlabelled::set_label(df1$tr_courts) <- "Trust in courts"
frq(df1$tr_courts)
```

# Social status

```{r income-hhold}
income_country <- names(df1 %>% dplyr::select(dplyr::ends_with("_INC")))
# lapply(df1[,income_country], table)
for (i in income_country) {
  # df1[[i]] <- car::recode(var = df1[[i]],"999990:99999999=NA")
  df1[[i]] <- as.numeric(df1[[i]])
  }

df1$AT_INC<- car::recode( df1$AT_INC,"999990:999999=NA")
df1$AU_INC<- car::recode( df1$AU_INC,"9999990:9999999=NA")
df1$CH_INC<- car::recode( df1$CH_INC,"999990:999999=NA")
df1$CN_INC<- car::recode( df1$CN_INC,"9999990:9999999=NA")
df1$CZ_INC<- car::recode( df1$CZ_INC,"999990:999999=NA")
df1$DE_INC<- car::recode( df1$DE_INC,"999990:999999=NA")
df1$DK_INC<- car::recode( df1$DK_INC,"9999990:9999999=NA")
df1$ES_INC<- car::recode( df1$ES_INC,"999990:999999=NA")
df1$FI_INC<- car::recode( df1$FI_INC,"999990:999999=NA")
df1$FR_INC<- car::recode( df1$FR_INC,"999990:999999=NA")
df1$GB_INC<- car::recode( df1$GB_INC,"999990:999999=NA")
df1$HR_INC<- car::recode( df1$HR_INC,"999990:999999=NA")
df1$HU_INC<- car::recode( df1$HU_INC,"9999990:9999999=NA")
df1$IL_INC<- car::recode( df1$IL_INC,"999990:999999=NA")
df1$IN_INC<- car::recode( df1$IN_INC,"999990:999999=NA")
df1$IS_INC<- car::recode( df1$IS_INC,"99999990:99999999=NA")
df1$JP_INC<- car::recode( df1$JP_INC,"99999990:99999999=NA")
df1$LT_INC<- car::recode( df1$LT_INC,"999990:999999=NA")
df1$MX_INC<- car::recode( df1$MX_INC,"999990:999999=NA")
df1$NZ_INC<- car::recode( df1$NZ_INC,"999990:999999=NA")
df1$PH_INC<- car::recode( df1$PH_INC,"999990:999999=NA")
df1$RU_INC<- car::recode( df1$RU_INC,"999990:999999=NA")
df1$SE_INC<- car::recode( df1$SE_INC,"999990:999999=NA")
df1$SI_INC<- car::recode( df1$SI_INC,"999990:999999=NA")
df1$SK_INC<- car::recode( df1$SK_INC,"999990:999999=NA")
df1$SR_INC<- car::recode( df1$SR_INC,"999990:999999=NA")
df1$TH_INC<- car::recode( df1$TH_INC,"999990:999999=NA")
df1$TW_INC<- car::recode( df1$TW_INC,"9999990:9999999=NA")
df1$US_INC<- car::recode( df1$US_INC,"999990:999999=NA")
df1$ZA_INC<- car::recode( df1$ZA_INC,"999990:999999=NA")

frq(df1$EE_INC)

c((200 )/1,
    (201 + 300)/2,
    (301 + 400)/2,
    (401 + 500)/2,
    (501 + 650)/2,
    (651 + 800)/2,
    (801 + 1000)/2,
    (1001 +1300)/2,
    (1301 + 1600)/2,
    (1601 +2100)/2,
    (2500)/1)

df1$EE_INC<- car::recode( df1$EE_INC,"1=200.0;2=250.5;3=350.5;4=450.5;5=575.5;6=725.5;7=900.5;8=1150.5;9=1450.5;10=1850.5;11=2500.0;999990:999999=NA")

frq(df1$TR_INC)
df1$TR_INC<- car::recode( df1$TR_INC,"999990:999997=NA")
# lapply(df1[,income_country], table)  # Ok el recode de ingresos familiares 
# Crear variable ingresos unica 
df1 <- df1 %>% 
  mutate(income = coalesce(AT_INC,AU_INC,CH_INC,CN_INC,CZ_INC,DE_INC,DK_INC,ES_INC,
                           FI_INC,FR_INC,GB_INC,HR_INC,HU_INC,IL_INC,IN_INC,IS_INC,
                           JP_INC,LT_INC,MX_INC,NZ_INC,PH_INC,RU_INC,SE_INC,SI_INC,
                           SK_INC,SR_INC,TH_INC,TW_INC,US_INC,ZA_INC,EE_INC,TR_INC))

table(df1$c_alphan,is.na(df1$income))

# summary(df1$JP_RINC)
# summary(df1$JP_RINC)
# 
# summary(df1$EE_INC)
# summary(df1$TR_INC)

#frq(df1$income==0)
frq(df1$HOMPOP)
df1$hhpop <- as.numeric(df1$HOMPOP) 
df1$hhpop[df1$hhpop %in%c(0,97,99)] <- NA
# frq(df1$hhpop)
df1$incomepcap <- as.numeric(df1$income/df1$hhpop) # Ingreso per capita del hogar

df1 <- df1 %>%group_by(country)  %>% mutate(D10pc=ntile(incomepcap,10),
                                              Q05pc=ntile(incomepcap,5),
                                              Q03pc=ntile(incomepcap,3),
                                              zincpc=scale(incomepcap,center = TRUE),
                                              logincpc=log(incomepcap)) %>% ungroup()

df1$Q05pcm <- car::recode(df1$Q05pc,"1='Q01';2='Q02';3='Q03';4='Q04';5='Q05';NA='Missing'",
                          as.factor = T,
                          levels = c("Q01","Q02","Q03","Q04","Q05","Missing")) 
frq(df1$Q05pc)

df1$Q03pcm <- car::recode(df1$Q03pc,"1='T01';2='T02';3='T03';NA='Missing'",
                          as.factor = T,
                          levels = c("T01","T02","T03","Missing")) 
frq(df1$Q03pcm)

table(as_character(df1$country),df1$Q03pcm)

sjlabelled::set_label(df1$D10pc) <- "Household income (D10)"
sjlabelled::set_label(df1$Q05pc) <- "Household income (Q05)"
sjlabelled::set_label(df1$zincpc) <- "Household income (zscore)"
sjlabelled::set_label(df1$logincpc) <- "Household income (log)"
```

```{r income-respondent}
income_country <- names(df1 %>% dplyr::select(dplyr::ends_with("_RINC")))
# lapply(df1[,income_country], table)
for (i in income_country) {
  # df1[[i]] <- car::recode(var = df1[[i]],"999990:99999999=NA")
  df1[[i]] <- as.numeric(df1[[i]])
  }

df1$AT_RINC<- car::recode( df1$AT_RINC,"999990:999999=NA")
df1$AU_RINC<- car::recode( df1$AU_RINC,"999990:999999=NA")
df1$CH_RINC<- car::recode( df1$CH_RINC,"999990:999999=NA")
df1$CN_RINC<- car::recode( df1$CN_RINC,"9999990:9999999=NA")
df1$CZ_RINC<- car::recode( df1$CZ_RINC,"999990:999999=NA")
df1$DE_RINC<- car::recode( df1$DE_RINC,"999990:999999=NA")
df1$DK_RINC<- car::recode( df1$DK_RINC,"999990:999999=NA")
df1$ES_RINC<- car::recode( df1$ES_RINC,"999990:999999=NA")
df1$FI_RINC<- car::recode( df1$FI_RINC,"999990:999999=NA")
df1$FR_RINC<- car::recode( df1$FR_RINC,"999990:999999=NA")
df1$GB_RINC<- car::recode( df1$GB_RINC,"999990:999999=NA")
df1$HR_RINC<- car::recode( df1$HR_RINC,"999990:999999=NA")
df1$HU_RINC<- car::recode( df1$HU_RINC,"999990:999999=NA")
df1$IL_RINC<- car::recode( df1$IL_RINC,"999990:999999=NA")
df1$IN_RINC<- car::recode( df1$IN_RINC,"999990:999999=NA")
df1$IS_RINC<- car::recode( df1$IS_RINC,"9999990:9999999=NA")
df1$JP_RINC<- car::recode( df1$JP_RINC,"99999990:99999999=NA")
df1$LT_RINC<- car::recode( df1$LT_RINC,"999990:999999=NA")
df1$MX_RINC<- car::recode( df1$MX_RINC,"999990:999999=NA")
df1$NZ_RINC<- car::recode( df1$NZ_RINC,"999990:999999=NA")
df1$PH_RINC<- car::recode( df1$PH_RINC,"999990:999999=NA")
df1$RU_RINC<- car::recode( df1$RU_RINC,"999990:999999=NA")
df1$SE_RINC<- car::recode( df1$SE_RINC,"999990:999999=NA")
df1$SI_RINC<- car::recode( df1$SI_RINC,"999990:999999=NA")
df1$SK_RINC<- car::recode( df1$SK_RINC,"999990:999999=NA")
df1$SR_RINC<- car::recode( df1$SR_RINC,"999990:999999=NA")
df1$TH_RINC<- car::recode( df1$TH_RINC,"999990:999999=NA")
df1$TW_RINC<- car::recode( df1$TW_RINC,"999990:999999=NA")
df1$US_RINC<- car::recode( df1$US_RINC,"999990:999999=NA")
df1$ZA_RINC<- car::recode( df1$ZA_RINC,"999990:999999=NA")

frq(df1$EE_RINC)
df1$EE_RINC<- car::recode( df1$EE_RINC,"1=200.0;2=250.5;3=350.5;4=450.5;5=575.5;6=725.5;7=900.5;8=1150.5;9=1450.5;10=1850.5;11=2500.0;999990:999999=NA")
frq(df1$EE_RINC)

frq(df1$TR_RINC)
df1$TR_RINC<- car::recode( df1$TR_RINC,"999990:999997=NA")
frq(df1$TR_RINC)

# lapply(df1[,income_country], table)  # Ok el recode de ingresos familiares 
# Crear variable ingresos unica 
df1 <- df1 %>% mutate(incomer = coalesce(AT_RINC,AU_RINC,
                                        CH_RINC, CN_RINC,
                                        CZ_RINC, DE_RINC,
                                        DK_RINC, ES_RINC,
                                        FI_RINC, FR_RINC,
                                        GB_RINC, HR_RINC,
                                        HU_RINC, IL_RINC,
                                        IN_RINC, IS_RINC,
                                        JP_RINC, LT_RINC,
                                        MX_RINC, NZ_RINC,
                                        PH_RINC, RU_RINC,
                                        SE_RINC, SI_RINC,
                                        SK_RINC, SR_RINC,
                                        TH_RINC, TW_RINC,
                                        US_RINC, ZA_RINC, 
                                        EE_RINC,TR_RINC))
table(df1$c_alphan,is.na(df1$income))
prop.table(table(df1$c_alphan,is.na(df1$income)),margin =1)

df1 <- df1 %>%group_by(country) %>% mutate(D10r=ntile(incomer,10),
                                              Q05r=ntile(incomer,5),
                                           Q03r=ntile(incomer,3),
                                              zincpcr=scale(incomer,center = TRUE),
                                              logincr=log(incomer)) %>% ungroup()

df1$Q05rm <- car::recode(df1$Q05r,"1='Q01';2='Q02';3='Q03';4='Q04';5='Q05';NA='Missing'",
                          as.factor = T,
                          levels = c("Q01","Q02","Q03","Q04","Q05","Missing"))

df1$Q03rm <- car::recode(df1$Q03r,"1='T01';2='T02';3='T03';NA='Missing'",
                          as.factor = T,
                          levels = c("T01","T02","T03","Missing"))

table(as_character(df1$country),df1$Q03rm)

# frq(df1$incomer)
# frq(df1$incomepcap)
# Single household 
df1$incomerep <- ifelse(df1$hhpop==1 & is.na(df1$incomepcap),df1$incomer,df1$incomepcap)
#frq(is.na(df1$incomerep))
#frq(df1$incomerep==0)
df1[,c("incomer","hhpop","incomepcap","incomerep")] %>% arrange(incomepcap,hhpop) %>% head()
df1$incomerep <- ifelse(df1$hhpop==1 & df1$incomerep==0,df1$incomer,df1$incomerep)
#frq(df1$incomerep==0)
df1[,c("incomer","hhpop","incomepcap","incomerep")] %>% arrange(incomepcap,hhpop) %>% head()

summary(df1[,c("incomer","hhpop","incomepcap","incomerep")])
```

```{r}
#frq(df1$EDUCYRS)
df1$edyears <- as.numeric(car::recode(df1$EDUCYRS,recodes = "98:99=NA")) 
sjlabelled::set_label(df1$edyear) <- "Years of schooling"
#frq(df1$edyears)

#frq(df1$DEGREE)
df1$ednum <- car::recode(df1$DEGREE+1,"10=NA")
sjlabelled::set_label(df1$ednum) <- "Educational level"
#frq(df1$ednum)
df1$edcat <- car::recode(df1$ednum,recodes = "1:2='Primary or less';
                         3:4='Secondary';
                         5='Terciary short-cycle';
                         6:7='Terciary or higher';
                         9=NA",
                         levels =c('Primary or less',
                                   'Secondary',
                                   'Post-secondary non-terciary',
                                   'Terciary or higher'),as.factor = T)
sjlabelled::set_label(df1$edcat) <- "Educational level"
#frq(df1$edcat)
```

```{r subjstatus}
df1$sstatus <- car::recode(df1$topbot, "c(0,98,99)=NA") 
#frq(df1$sstatus)
```

# Occupational variables

```{r labour-status}
#frq(df1$MAINSTAT)
df1$labst <- car::recode(
  as.numeric(df1$MAINSTAT),
  recodes = "1='In labour force';2:9='Not in labour force';c(0,99)='No information'",
  levels = c('In labour force', 'Not in labour force', 'No information'),
  as.factor = T
)
sjlabelled::set_label(df1$labst) <- "Occupational status"
#frq(df1$labst)
```

```{r paid-work}
frq(df1$WORK)
df1$workst <- car::recode(
  as.numeric(df1$WORK),
  recodes = "1='In paid work';2:3='Not in paid work';c(9)=NA",
  levels = c('In paid work', 'Not in paid work'),
  as.factor = T
)
sjlabelled::set_label(df1$workst) <- "Work status"
frq(df1$workst)
```

```{r union}
frq(df1$UNION)
df1$union <- car::recode(as.numeric(df1$UNION),
                         recodes= "1:2='Yes';c(0,3,7)='No';8:9=NA",
                         levels = c("No","Yes"))
frq(df1$union)
```

```{r isei}
df1$ISCO08[df1$ISCO08 %in% c(9998,9999)] <- NA
df1$isei08 <- occupar::isco08toISEI08(isco08 = df1$ISCO08,display.nas = T)
# summary(df1$isei08)
sjlabelled::set_label(df1$isei08) <- "ISEI respondent"

#frq(df1$SPISCO08)
df1$SPISCO08[df1$SPISCO08 %in% c(9998,9999)] <- NA
df1$isei08sp <- occupar::isco08toISEI08(isco08 = df1$SPISCO08,display.nas = T)
# summary(df1$isei08sp)
sjlabelled::set_label(df1$isei08sp) <- "ISEI spouse"
```

```{r class}
## Ocupación
# Ocupación u oficio actual ISCO 08
#frq(df1$SPISCO08) # isco08 
# Recode CIUO_encuestado desde ISCO 2008 a ISCO 1988
df1$isco88spo <- occupar::isco08to88(df1$SPISCO08)
#frq(df1$isco88spo)

#Relación de Empleo encuestado
#frq(df1$SPEMPREL)

df1$sempspo <- df1$SPEMPREL
df1$sempspo <- as.numeric(car::recode(df1$sempspo, "c(2,3,4,5)=1;c(1)=2;c(0,8,9)=NA"))
#frq(df1$sempspo)

df1$sempspo <- sjlabelled::set_labels(x = df1$sempspo,labels=c('Self-employed','Employee'))
#frq(df1$sempspo)

# superivison
#frq(df1$SPWRKSUP) #supervision spouse
df1$supvisspo <- as.numeric(car::recode(df1$SPWRKSUP,"c(0,2)=0;1=9;c(8,9)=NA"))
#frq(df1$supvisspo)

## Ocupación
# Ocupación u oficio actual ISCO 08
#frq(df1$ISCO08) # isco08 
# Recode CIUO_encuestado desde ISCO 2008 a ISCO 1988
df1$isco88r <- occupar::isco08to88(df1$ISCO08)
#frq(df1$isco88r)

#Note: the package generate identical recodes as my own manual recoding

#Employment relation encuestado
#frq(df1$EMPREL)
#frq(df1$WORK)
df1$semp <- df1$EMPREL
df1$semp <- as.numeric(car::recode(df1$semp, "c(2,3,4)=1;c(1)=2;c(0,8,9)=NA"))
#frq(df1$semp)
df1$semp <- sjlabelled::set_labels(x = df1$semp,labels=c('Self-employed','Employee'))
#frq(df1$semp)

# superivison
#frq(df1$NEMPLOY) #supervision respondent
df1$supvis <- as.numeric(car::recode(df1$NEMPLOY,"c(9998,9999)=NA"))
#frq(df1$supvis)
# save the occupational variables for using them in stata-----------------------
df1$id <- NULL
df1 <- tibble::rowid_to_column(df1, "id")
haven::write_dta(data = df1[c("id","isco88r","semp","supvis",
                              "isco88spo","sempspo","supvisspo")],
                 path = here::here("input/data/proc/ocup_respondent-spouse.dta"))
```

```{r class-stata}
# Load the data created in stata------------------------------------------------
df_class <- haven::read_dta(file = here::here("input/data/proc/class_respondent-spouse.dta"))
df1 <- left_join(df1,df_class,by='id')

# Class R´s and Spouse----------------------------------------------------------
df1$class10 <- car::recode(df1$class10, "7=8; 8=9; 9=10; 10=11; 11=7")
df1$class10 <- sjlabelled::set_labels(df1$class10, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11),
                        labels = c("I", "II", "IIIa", "IIIb", "IVa", "IVb", "IVc", "V", "VI", "VIIa", "VIIb"))

sjlabelled::set_label(df1$class10) <- "Social Class - respondent"

#frq(df1$class10)

df1$class10spo <- car::recode(df1$class10spo, "7=8; 8=9; 9=10; 10=11; 11=7")
df1$class10spo <- sjlabelled::set_labels(df1$class10spo, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11),
                        labels = c("I", "II", "IIIa", "IIIb", "IVa", "IVb", "IVc", "V", "VI", "VIIa", "VIIb"))

sjlabelled::set_label(df1$class10spo) <- "Social Class - spouse"

#frq(df1$class10spo)

#if NA in class of respondent, replace with household

#frq(df1$class10)
#frq(df1$class10spo)

df1$class11 <- ifelse(is.na(df1$class10),df1$class10spo,df1$class10)
#frq(df1$class11)

# if NA in respondents it is assumed to be lower than spouse
# know, with the cases in we have information for respondent and spouse
# we replace the value by the highest one, just for those

# Replace the highest class by dominance criteria 
# Keep the highest of two values:
df1$class11d <- pmax(df1$class11, #social class respondent + spouse vs:
                       df1$class10spo,#social class spouse
                       na.rm = T) # if NA, do nothing

df1$class11d <- 
  sjlabelled::set_labels(df1$class11d,
                         levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11),
                         labels =  c("I", "II", "IIIa", "IIIb", "IVa", "IVb", 
                                     "IVc", "V", "VI", "VIIa", "VIIb")) 
#frq(df1$class11d)
sjlabelled::set_label(df1$class11d) <- "Social Class"

# Class Scheme 3 classes
df1$class3 <- car::recode(df1$class11d,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV)';
                             8:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$class3) <- "Social Class"
#frq(df1$class3)

# Class Scheme 5 classes
df1$class5 <- car::recode(
  df1$class11d,
  as.factor = T,
  levels = c(
    'Service class (I+II)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1:2='Service class (I+II)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

sjlabelled::set_label(df1$class5) <- "Social Class"
#frq(df1$class5)

# Class Scheme 6 classes
df1$class6 <- car::recode(
  df1$class11d,
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

# Class only R's --------------------------------------------------------------
df1$class10

# Class Scheme 3 classes
df1$class3res <- car::recode(as.numeric(df1$class10),as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV)';
                             8:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$class3res) <- "Social Class (R's)"
# frq(df1$class3res)

# Class Scheme 5 classes
df1$class5res <- car::recode(
  as.numeric(df1$class10),
  as.factor = T,
  levels = c(
    'Service class (I+II)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1:2='Service class (I+II)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

sjlabelled::set_label(df1$class5res) <- "Social Class (R's)"
# frq(df1$class5res)

# Class Scheme 6 classes
df1$class6res <- car::recode(
  as.numeric(df1$class10),
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

sjlabelled::set_label(df1$class6res) <- "Social Class (R's)"
frq(df1$class6res)
# Class only Spouse --------------------------------------------------------------
df1$class10spo

# Class Scheme 3 classes
df1$class3spo <- car::recode(as.numeric(df1$class10spo),as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV)';
                             8:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$class3spo) <- "Social Class (Spouse)"
frq(df1$class3spo)

# Class Scheme 5 classes
df1$class5spo <- car::recode(
  as.numeric(df1$class10spo),
  as.factor = T,
  levels = c(
    'Service class (I+II)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1:2='Service class (I+II)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

sjlabelled::set_label(df1$class5spo) <- "Social Class (Spouse)"
# frq(df1$class5res)

# Class Scheme 6 classes
df1$class6spo <- car::recode(
  as.numeric(df1$class10spo),
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)
sjlabelled::set_label(df1$class6spo) <- "Social Class (Spouse)"
```

# Networks

```{r generador}
ocupations <- names(dplyr::select(df1,v1:v10))
for (i in ocupations) {
  df1[[i]] <- car::recode(df1[[i]],"c(8,9)=NA")
  df1[[i]] <- sjlabelled::set_na(df1[[i]],na = c(8,9))
}

#frq(df1$v1)
df1$busdriver <- as.numeric(car::recode(df1$v1,"1:3=1;4=0")) 
sjlabelled::set_label(df1$busdriver) <- "Know: bus/lorry driver" 
#frq(df1$busdriver)

#frq(df1$v2)
df1$ceo <- as.numeric(car::recode(df1$v2,"1:3=1;4=0")) 
sjlabelled::set_label(df1$ceo) <- "Know: CEO large company" 
#frq(df1$ceo)

#frq(df1$v3)
df1$cleaner <- as.numeric(car::recode(df1$v3,"1:3=1;4=0")) 
sjlabelled::set_label(df1$cleaner) <- "Know: home or office cleaner" 
#frq(df1$cleaner)

#frq(df1$v4)
df1$hairbarber <- as.numeric(car::recode(df1$v4,"1:3=1;4=0")) 
sjlabelled::set_label(df1$hairbarber) <- "Know: hairdresser/barber" 
#frq(df1$hairbarber)

#frq(df1$v5)
df1$hrmanager <- as.numeric(car::recode(df1$v5,"1:3=1;4=0")) 
sjlabelled::set_label(df1$hrmanager) <- "Know: human resource manager" 
#frq(df1$hrmanager)

#frq(df1$v6)
df1$lawyer <- as.numeric(car::recode(df1$v6,"1:3=1;4=0")) 
sjlabelled::set_label(df1$lawyer) <- "Know: lawyer" 
#frq(df1$lawyer)

#frq(df1$v7)
df1$mechanic <- as.numeric(car::recode(df1$v7,"1:3=1;4=0")) 
sjlabelled::set_label(df1$mechanic) <- "Know: car mechanic" 
#frq(df1$mechanic)

#frq(df1$v8)
df1$nurse <- as.numeric(car::recode(df1$v8,"1:3=1;4=0")) 
sjlabelled::set_label(df1$nurse) <- "Know: nurse" 
#frq(df1$mechanic)

#frq(df1$v9)
df1$police <- as.numeric(car::recode(df1$v9,"1:3=1;4=0")) 
sjlabelled::set_label(df1$police) <- "Know: police officer" 
#frq(df1$police)

#frq(df1$v10)
df1$teacher <- as.numeric(car::recode(df1$v10,"1:3=1;4=0")) 
sjlabelled::set_label(df1$teacher) <- "Know: school teacher" 
#frq(df1$teacher)

# Total of known occupations
isei_ocupnames<- names(select(df1,busdriver:teacher))
df1$know_total <- rowSums(x = select(df1,isei_ocupnames),na.rm = T)
# summary(df1$know_total)
sjlabelled::set_label(df1$know_total) <- 'Total of known occupations'
```

```{r strenght-ties}
ocupations <- names(dplyr::select(df1,v1:v10))
vl<- dplyr::select(df1,busdriver:teacher) %>% get_label() %>% stringr::str_replace_all(pattern = "Know:",replacement = "Tie strenght:")
vl2<- dplyr::select(df1,busdriver:teacher) %>% get_label() %>% stringr::str_replace_all(pattern = "Know:",replacement = "Strong tie:")
vl3<- dplyr::select(df1,busdriver:teacher) %>% get_label() %>% stringr::str_replace_all(pattern = "Know:",replacement = "Weak tie:")

for (i in 1:length(ocupations)) {
  df1[[i]] <- car::recode(df1[[i]],"c(8,9)=NA")
  df1[[i]] <- sjlabelled::set_na(df1[[i]],na = c(8,9))
}
lb <- c("Absent"=0,"Weak"=1,"Strong"=2)
lb2 <- c("Strong"=1,"Weak"=0)
lb3 <- c("Strong"=0,"Weak"=1)
r <- "4=0;3=1;1:2=2"
r2 <- "1:2=1;3=0;4=NA"
r3 <- "1:2=0;3=1;4=NA"

df1 <- df1 %>%
  # Tie strenght
  mutate("busdriver_ts" = set_labels(car::recode(df1$v1,r),labels = lb),
         "ceo_ts" =  set_labels(car::recode(df1$v2,r),labels = lb),
         "cleaner_ts" = set_labels(car::recode(df1$v3,r),labels = lb),
         "hairbarber_ts"  = set_labels(car::recode(df1$v4,r),labels = lb),
         "hrmanager_ts" = set_labels(car::recode(df1$v5,r),labels = lb),
         "lawyer_ts"  = set_labels(car::recode(df1$v6,r),labels = lb),
         "mechanic_ts" = set_labels(car::recode(df1$v7,r),labels = lb),
         "nurse_ts" = set_labels(car::recode(df1$v8,r),labels = lb),
         "police_ts"  = set_labels(car::recode(df1$v9,r),labels = lb),
         "teacher_ts" = set_labels(car::recode(df1$v10, r, as.factor = T), labels = lb)) %>%
  # strong ties  
  mutate("busdriver_s" = set_labels(car::recode(df1$v1,r2),labels = lb2),
         "ceo_s" =  set_labels(car::recode(df1$v2,r2),labels = lb2),
         "cleaner_s" = set_labels(car::recode(df1$v3,r2),labels = lb2),
         "hairbarber_s"  = set_labels(car::recode(df1$v4,r2),labels = lb2),
         "hrmanager_s" = set_labels(car::recode(df1$v5,r2),labels = lb2),
         "lawyer_s"  = set_labels(car::recode(df1$v6,r2),labels = lb2),
         "mechanic_s" = set_labels(car::recode(df1$v7,r2),labels = lb2),
         "nurse_s" = set_labels(car::recode(df1$v8,r2),labels = lb2),
         "police_s"  = set_labels(car::recode(df1$v9,r2),labels = lb2),
         "teacher_s" = set_labels(car::recode(df1$v10, r2), labels = lb2)) %>% 
  # weak ties
  mutate("busdriver_w" = set_labels(car::recode(df1$v1,r3),labels = lb3),
         "ceo_w" =  set_labels(car::recode(df1$v2,r3),labels = lb3),
         "cleaner_w" = set_labels(car::recode(df1$v3,r3),labels = lb3),
         "hairbarber_w"  = set_labels(car::recode(df1$v4,r3),labels = lb3),
         "hrmanager_w" = set_labels(car::recode(df1$v5,r3),labels = lb3),
         "lawyer_w"  = set_labels(car::recode(df1$v6,r3),labels = lb3),
         "mechanic_w" = set_labels(car::recode(df1$v7,r3),labels = lb3),
         "nurse_w" = set_labels(car::recode(df1$v8,r3),labels = lb3),
         "police_w"  = set_labels(car::recode(df1$v9,r3),labels = lb3),
         "teacher_w" = set_labels(car::recode(df1$v10, r3), labels = lb3))

v = names(dplyr::select(df1,ends_with("_ts")))
df1[,v]<- set_label(x = df1[,v],label = vl) #label Tie strenght

s = names(dplyr::select(df1,ends_with("_s")))
df1[,s]<- set_label(x = df1[,s],label = vl2) #label strong tie

w = names(dplyr::select(df1,ends_with("_w")))
df1[,w]<- set_label(x = df1[,w],label = vl3) #label weak tie

df1$know_total_s <- rowSums(x = select(df1,ends_with("_s")),na.rm = T)
sjlabelled::set_label(df1$know_total_s) <- 'Total of known occupations (strong ties)'

df1$know_total_w <- rowSums(x = select(df1,ends_with("_w")),na.rm = T)
sjlabelled::set_label(df1$know_total_s) <- 'Total of known occupations (weak ties)'

```

```{r strenght-num}
# strong ties
df1$n_high_s<-rowSums(x = df1[c("lawyer_s","hrmanager_s","ceo_s")],na.rm = T)
set_label(df1$n_high_s) <- 'Number of high status contacts (strong ties)'

df1$n_middle_s<-rowSums(x = df1[c("teacher_s","police_s","nurse_s")],na.rm = T)
set_label(df1$n_middle_s) <- 'Number of middle status contacts (strong ties)'

df1$n_low_s <- rowSums(x = df1[c("mechanic_s","busdriver_s","hairbarber_s","cleaner_s")],na.rm = T)
set_label(df1$n_low_s) <- 'Number of low status contacts (strong ties)'

# weak ties
df1$n_high_w<-rowSums(x = df1[c("lawyer_w","hrmanager_w","ceo_w")],na.rm = T)
set_label(df1$n_high_w) <- 'Number of high status contacts (weak ties)'

df1$n_middle_w<-rowSums(x = df1[c("teacher_w","police_w","nurse_w")],na.rm = T)
set_label(df1$n_middle_w) <- 'Number of middle status contacts (weak ties)'

df1$n_low_w <- rowSums(x = df1[c("mechanic_w","busdriver_w","hairbarber_w","cleaner_w")],na.rm = T)
set_label(df1$n_low_w) <- 'Number of low status contacts (weak ties)'
```

```{r isei-generador}
df1$isei_busdriver <- as.numeric(ifelse(df1$busdriver ==1,37,NA)) 
sjlabelled::set_label(df1$isei_busdriver) <- "ISEI: bus/lorry driver" 
#frq(df1$isei_busdriver)

df1$isei_ceo <- as.numeric(ifelse(df1$ceo==1,70,NA)) 
sjlabelled::set_label(df1$isei_ceo) <- "ISEI: CEO large company" 
#frq(df1$isei_ceo)

df1$isei_cleaner <- as.numeric(ifelse(df1$cleaner==1,17,NA)) 
sjlabelled::set_label(df1$isei_cleaner) <- "ISEI: home or office cleaner" 
#frq(df1$isei_cleaner)

df1$isei_hairbarber <- as.numeric(ifelse(df1$hairbarber==1,32,NA)) 
sjlabelled::set_label(df1$isei_hairbarber) <- "ISEI: hairdresser/barber" 
#frq(df1$isei_hairbarber)

df1$isei_hrmanager <- as.numeric(ifelse(df1$hrmanager==1,68,NA)) 
sjlabelled::set_label(df1$isei_hrmanager) <- "ISEI: human resource manager" 
#frq(df1$isei_hrmanager)

df1$isei_lawyer <- as.numeric(ifelse(df1$lawyer==1,85,NA)) 
sjlabelled::set_label(df1$isei_lawyer) <- "ISEI: lawyer" 
#frq(df1$isei_lawyer)

df1$isei_mechanic <- as.numeric(ifelse(df1$mechanic==1,38,NA)) 
sjlabelled::set_label(df1$isei_mechanic) <- "ISEI: car mechanic" 
#frq(df1$isei_mechanic)

df1$isei_nurse <- as.numeric(ifelse(df1$nurse==1,48,NA)) 
sjlabelled::set_label(df1$isei_nurse) <- "ISEI: nurse" 
#frq(df1$isei_nurse)

df1$isei_police <- as.numeric(ifelse(df1$police==1,53,NA)) 
sjlabelled::set_label(df1$isei_police) <- "ISEI: police officer" 
#frq(df1$isei_police)

df1$isei_teacher <- as.numeric(ifelse(df1$teacher==1,63,NA)) 
sjlabelled::set_label(df1$isei_teacher) <- "ISEI: school teacher" 
#frq(df1$isei_teacher)
```

```{r social-capital}
# Social capital measures
isei_ocupnames<- names(select(df1,starts_with("isei_")))

# number of high status contacts
df1$n_high <- 
  rowSums(x = select(df1,c("lawyer","hrmanager","ceo")),na.rm = T)
sjlabelled::set_label(df1$n_high) <- 'Number of high status contacts'
# sjPlot::plot_#frq(df1$n_high,type = 'density')

# percentage of high status contacts

df1$pct_high <- (df1$n_high/df1$know_total)*100
sjlabelled::set_label(df1$df1$pct_high ) <- 'Percentage of high status contacts'
# sjPlot::plot_#frq(df1$pct_high,type = 'density')

# number of middle status contacts
df1$n_middle <-
  rowSums(x = select(df1,c("teacher","police","nurse")),na.rm = T)
sjlabelled::set_label(df1$n_middle) <- 'Number of middle status contacts'
# sjPlot::plot_#frq(df1$n_middle,type = 'density')

# percentage of middle status contacts
df1$pct_middle <- (df1$n_middle/df1$know_total)*100
sjlabelled::set_label(df1$pct_middle) <- 'Percentage of middle status contacts'
# sjPlot::plot_#frq(df1$pct_middle,type = 'density')

# number of low status contacts
# df1$n_low <- 
#   rowSums(x = select(df1,c("n_waiter","n_carmech","n_taxi","n_stvend","n_cleaner")),na.rm = T)
df1$n_low <- 
  rowSums(x = select(df1,c("mechanic","busdriver","hairbarber","cleaner")),na.rm = T)
sjlabelled::set_label(df1$n_low) <- 'Number of low status contacts'
# sjPlot::plot_#frq(df1$n_low,type = 'density')

# percentage of low status contacts
df1$pct_low <- (df1$n_low/df1$know_total)*100
sjlabelled::set_label(df1$pct_low) <- 'Percentage of low status contacts'
# sjPlot::plot_#frq(df1$pct_low,type = 'density')

# Average ISEI 
df1$isei_avg <- rowMeans(x = select(df1,isei_ocupnames),na.rm = T) #REVISAR NO DEBERIA CONTENER 0
sjlabelled::set_label(df1$isei_avg) <- 'Average contact prestige'
summary(df1$isei_avg)

# sjPlot::plot_#frq(df1$isei_avg,type = 'density')

# Total ISEI 
df1$isei_tot <- rowSums(x = select(df1,isei_ocupnames),na.rm = T)
sjlabelled::set_label(df1$isei_tot) <- 'Total contact prestige'
summary(df1$isei_tot)
# sjPlot::plot_#frq(df1$isei_tot,type = 'density')

# skimr::skim(df1$isei_tot/df1$n_total)
# skimr::skim(df1$isei_tot/df1$know_total)
# skimr::skim(df1$isei_avg)

# max ISEI 
df1 <- df1 %>% 
mutate(isei_max = do.call(pmax, c(select(., isei_ocupnames), na.rm = TRUE)))
sjlabelled::set_label(df1$isei_max) <- 'Maximum contact prestige'
# sjmisc::frq(df1$isei_max)
# sjPlot::plot_#frq(df1$isei_max)

# min ISEI
df1 <- df1 %>% 
mutate(isei_min = do.call(pmin, c(select(., isei_ocupnames), na.rm = TRUE)))
sjlabelled::set_label(df1$isei_min) <- 'Minimum contact prestige'
# sjmisc::frq(df1$isei_min)
# sjPlot::plot_#frq(df1$isei_min)

# range ISEI
table(df1$know_total>=2)
df1$isei_range <- ifelse(df1$know_total>=2,(df1$isei_max - df1$isei_min),no = NA)

sjlabelled::set_label(df1$isei_range) <- 'Range contact prestige'
# sjmisc::frq(df1$isei_range)
# sjPlot::plot_#frq(df1$isei_range)
# df1 %>% select(starts_with("isei_")) %>%View

cor(df1[,c("isei_avg","isei_max","isei_range",
                 "know_total","n_high","n_middle","n_low",
                 "pct_high","pct_middle","pct_low")],use = "complete.obs")

```

```{r heterogeneity}
isei_ocupnames<- names(select(df1,busdriver:teacher))

df1$hindex_sta<- 
as.numeric(
  1 - (
    #p1^2
    (df1$n_high / df1$know_total)^2 +
    #p2^2
    (df1$n_middle / df1$know_total)^2 +
    #p3^2
    (df1$n_low / df1$know_total)^2
    )
  )

sjlabelled::set_label(df1$hindex_sta) <- "D-index: status groups"

df1$hindex_10<- 
as.numeric(
  1 - (
    #p1^2
    (df1$busdriver/df1$know_total)^2 +
    #p2^2
    (df1$ceo / df1$know_total)^2 +
    #p3^2
    (df1$cleaner / df1$know_total)^2 +
    #p3^2
    (df1$hairbarber / df1$know_total)^2 +
     #p3^2
    (df1$hrmanager / df1$know_total)^2 +
    #p3^2
    (df1$lawyer / df1$know_total)^2+
    #p3^2
    (df1$mechanic / df1$know_total)^2+
    #p3^2
    (df1$nurse / df1$know_total)^2+
    #p3^2
    (df1$police / df1$know_total)^2+    
    #p3^2
    (df1$teacher / df1$know_total)^2        
    )
  )

sjlabelled::set_label(df1$hindex_10) <- "D-index: 10 ocupations"
```

```{r homogeneity}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(class3,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$class3=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$class3=="Intermediate class (III+IV)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$class3=="Working Class (V+VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

df1$outgroup <- NA 
df1 %>% select(class3,n_high,n_middle,n_low,outgroup)
df1$outgroup <- ifelse(df1$class3=="Service Class (I+II)",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$class3=="Intermediate class (III+IV)",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$class3=="Working Class (V+VI+VII)",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup)
#frq(df1$outgroup)

# df1$eindex<- (df1$outgroup-df1$ingroup)/(df1$outgroup+df1$ingroup)
df1$eindex<- (df1$ingroup-df1$outgroup)/(df1$outgroup+df1$ingroup)
sjlabelled::set_label(df1$eindex) <- "EI homophily index"
#frq(df1$eindex)
hist.default(df1$eindex)
df1$eindex_gc <- group_center(x = df1$eindex,grp = df1$country)
hist.default(df1$eindex_gc)

df1 %>% select(class3,n_high,n_middle,n_low,outgroup,ingroup,eindex)
df1$homclass <- as.numeric(df1$ingroup/df1$know_total)
sjlabelled::set_label(df1$homclass) <- "Class homogeneity"
df1$homclass_gc <- group_center(x = df1$homclass,grp = df1$country)
# View(df1[,c("homclass","homclass_gc")])
# Weighted EI index
# https://arxiv.org/pdf/2109.00348.pdf

# Thus, an EI score of -1 means complete homophily- the individual only has relationships with actors of the same “type” as they themselves are. An EI score of 1 means complete heterophily- all the alters are of a different “type” than they themselves are. Finally, an EI score of 0 means that an equal number of alters are of both the same “type” as the ego, and different types
```

```{r homogeneity-strong}
# EI homophily index
df1$ingroup_s<- NA 
df1 %>% select(class3,n_high_s,n_middle_s,n_low_s,ingroup_s)
df1$ingroup_s <- ifelse(df1$class3=="Service Class (I+II)",df1$n_high_s,df1$ingroup_s)
df1$ingroup_s <- ifelse(df1$class3=="Intermediate class (III+IV)",df1$n_middle_s,df1$ingroup_s)
df1$ingroup_s <- ifelse(df1$class3=="Working Class (V+VI+VII)",df1$n_low_s,df1$ingroup_s)
#frq(df1$ingroup_s)

df1$outgroup_s <- NA 
df1 %>% select(class3,n_high_s,n_middle_s,n_low_s,outgroup_s)
df1$outgroup_s <- ifelse(df1$class3=="Service Class (I+II)",rowSums(df1[,c("n_low_s","n_middle_s")],na.rm = T),df1$outgroup_s)
df1$outgroup_s <- ifelse(df1$class3=="Intermediate class (III+IV)",rowSums(df1[,c("n_low_s","n_high_s")],na.rm = T),df1$outgroup_s)
df1$outgroup_s <- ifelse(df1$class3=="Working Class (V+VI+VII)",rowSums(df1[,c("n_middle_s","n_high_s")],na.rm=T),df1$outgroup_s)
#frq(df1$outgroup_s)

df1$eindex_s<- (df1$ingroup_s-df1$outgroup_s)/(df1$outgroup_s+df1$ingroup_s)
sjlabelled::set_label(df1$eindex_s) <- "EI homophily index (strong ties)"
#frq(df1$eindex_s)
hist.default(df1$eindex_s)

df1 %>% select(class3,n_high_s,n_middle_s,n_low_s,outgroup_s,ingroup_s,eindex_s)
df1$homclass_s <- as.numeric(df1$ingroup_s/df1$know_total_s)
sjlabelled::set_label(df1$homclass_s) <- "Class homogeneity (strong ties)"
hist.default(df1$homclass_s)
```

```{r homogeneity-weak}
# EI homophily index
df1$ingroup_w<- NA 
df1 %>% select(class3,n_high_w,n_middle_w,n_low_w,ingroup_w)
df1$ingroup_w <- ifelse(df1$class3=="Service Class (I+II)",df1$n_high_w,df1$ingroup_w)
df1$ingroup_w <- ifelse(df1$class3=="Intermediate class (III+IV)",df1$n_middle_w,df1$ingroup_w)
df1$ingroup_w <- ifelse(df1$class3=="Working Class (V+VI+VII)",df1$n_low_w,df1$ingroup_w)
# frq(df1$ingroup_w)

df1$outgroup_w <- NA 
df1 %>% select(class3,n_high_w,n_middle_w,n_low_w,outgroup_w)
df1$outgroup_w <- ifelse(df1$class3=="Service Class (I+II)",rowSums(df1[,c("n_low_w","n_middle_w")],na.rm = T),df1$outgroup_w)
df1$outgroup_w <- ifelse(df1$class3=="Intermediate class (III+IV)",rowSums(df1[,c("n_low_w","n_high_w")],na.rm = T),df1$outgroup_w)
df1$outgroup_w <- ifelse(df1$class3=="Working Class (V+VI+VII)",rowSums(df1[,c("n_middle_w","n_high_w")],na.rm=T),df1$outgroup_w)
# frq(df1$outgroup_w)

df1$eindex_w<- (df1$ingroup_w-df1$outgroup_w)/(df1$outgroup_w+df1$ingroup_w)
sjlabelled::set_label(df1$eindex_w) <- "EI homophily index (weak ties)"
frq(df1$eindex_w)
hist.default(df1$eindex_w)

df1 %>% select(class3,n_high_w,n_middle_w,n_low_w,outgroup_w,ingroup_w,eindex_w)
df1$homclass_w <- as.numeric(df1$ingroup_w/df1$know_total_w)
sjlabelled::set_label(df1$homclass_w) <- "Class homogeneity (weak ties)"
hist.default(df1$homclass_s)
```

```{r social distance}
df1$dis_busdriver <- df1$isei08 - df1$isei_busdriver
df1$dis_ceo <- df1$isei08 - df1$isei_ceo
df1$dis_cleaner <- df1$isei08 - df1$isei_cleaner
df1$dis_hairbarber <- df1$isei08 - df1$isei_hairbarber
df1$dis_hrmanager <- df1$isei08 - df1$isei_hrmanager
df1$dis_lawyer <- df1$isei08 - df1$isei_lawyer
df1$dis_mechanic <- df1$isei08 - df1$isei_mechanic
df1$dis_nurse <- df1$isei08 - df1$isei_nurse
df1$dis_police <- df1$isei08 - df1$isei_police
df1$dis_teacher <- df1$isei08 - df1$isei_teacher

# 88 - 37  = 51
# 88 - 17  = 71

df1$socdist<- rowMeans(x = df1 %>% select(starts_with("dis_")),na.rm = T)
sjlabelled::set_label(df1$socdist) <- "Social distance: ISEI (respondent - contacts)"
# sjPlot::plot_#frq(df1$socdist,type = "histogram")

# df1$socdist_tot<- (df1$isei08- df1$isei_avg)
# sjPlot::plot_#frq(df1$socdist_tot,type = "histogram")
# sjlabelled::set_label(df1$socdist_tot) <- "Social distance: ISEI (respondent - contacts)"

#higher values represent higher socioeconomic distance from my contacts. In other words, higher values means that in average, my contacts are more different from me in terms of socioeconomic status (ISEI), while lower values represent higher homogeneity.
```

```{r homogeneity-resp}
# EI homophily index
df1$ingroup_res <- NA 
df1 %>% select(class3res,n_high,n_middle,n_low,ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res=="Service Class (I+II)",df1$n_high,df1$ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res=="Intermediate class (III+IV)",df1$n_middle,df1$ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res=="Working Class (V+VI+VII)",df1$n_low,df1$ingroup_res)
# frq(df1$ingroup_res)

df1$outgroup_res <- NA 
df1 %>% select(class3res,n_high,n_middle,n_low,outgroup_res)
df1$outgroup_res <- ifelse(df1$class3res=="Service Class (I+II)",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup_res)
df1$outgroup_res <- ifelse(df1$class3res=="Intermediate class (III+IV)",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup_res)
df1$outgroup_res <- ifelse(df1$class3res=="Working Class (V+VI+VII)",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup_res)
# frq(df1$outgroup_res)

df1$eindex_res<- (df1$ingroup_res-df1$outgroup_res)/(df1$outgroup_res+df1$ingroup_res)
sjlabelled::set_label(df1$eindex_res) <- "EI homophily index (R's)"
# frq(df1$eindex)
# hist.default(df1$eindex)

df1 %>% select(class3res,n_high,n_middle,n_low,outgroup_res,ingroup_res,eindex_res)
df1$homclass_res <- as.numeric(df1$ingroup_res/df1$know_total)
sjlabelled::set_label(df1$homclass_res) <- "Class homogeneity (R's)"
# Weighted EI index
# https://arxiv.org/pdf/2109.00348.pdf

# Thus, an EI score of -1 means complete homophily- the individual only has relationships with actors of the same “type” as they themselves are. An EI score of 1 means complete heterophily- all the alters are of a different “type” than they themselves are. Finally, an EI score of 0 means that an equal number of alters are of both the same “type” as the ego, and different types
```

(i)  the highest status job accessed (upper reachability),
(ii) the range between highest and lowest occupations accessed (resource heterogeneity)
(iii) the total number of occupations accessed (network extensity),
(iv) the total accessed prestige (resource volume),
(v) the average status of jobs accessed (network prestige),
(vi) the total number of higher-status positions accessed,
(vii) the total number of medium-status positions accessed,
(viii) the total number of lower-status positions accessed.

```{r pca}
library(psych)
library(dplyr)
isei_ocupnames<- names(dplyr::select(df1,starts_with("isei_")))

# Calculate the standard deviation for each column
df1$isei_sd <- apply(df1[isei_ocupnames], 1, sd, na.rm = TRUE)

# Print the standard deviation for each column
# df1 %>% dplyr::select(isei_ocupnames,isei_sd) %>% View()

# reduce dimensions in 2: diversity and prestige
df_cs<- df1[,c("isei_max","isei_avg","isei_range","know_total","isei_sd")]
summary(df_cs)
dim(df_cs)

# df_cs <- na.omit(df_cs)
cor<- corr.test(df_cs)
psych::KMO(r = cor$r)
cortest.bartlett(cor$r,n = dim(df_cs)[1])
plot(scree(df_cs,factors = 2))

# Porcentaje de varianza explicada
fit1<- psych::pca(df_cs,nfactors = 2,rotate = "varimax")
fit1$loadings
plot(fit1,labels = c(names(df_cs)))
cor(fit1$scores)

fit2<- psych::fa(df_cs,nfactors = 2)
plot(fit2,labels = c(names(df_cs)))
fit2$scores
cor(fit2$scores)
fs<- as.data.frame(fit2$scores)
names(fs) <- c("diversity","prestige")

df1 <- bind_cols(df1,fs)
sjlabelled::set_label(df1$prestige) <- "Network Contacts Prestige"
sjlabelled::set_label(df1$diversity) <- "Network Contacts Diversity"
```

# Community participation

```{r}
# frq(df1$v17)
df1$act_lei <- car::recode(df1$v17,recodes = "c(0,8,9)=NA")
df1$act_lei <- set_na(df1$act_lei,na = c(0,8,9))
df1$act_lei <- sjmisc::rec(x = df1$act_lei,rec = "rev")
sjlabelled::set_label(df1$act_lei) <- "Activities: groups for leisure, sports or culture"
# frq(df1$act_lei)

# frq(df1$v18)
df1$act_pol <- car::recode(df1$v18,recodes = "c(0,8,9)=NA")
df1$act_pol <- set_na(df1$act_pol,na = c(0,8,9))
df1$act_pol <- sjmisc::rec(x = df1$act_pol,rec = "rev")
sjlabelled::set_label(df1$act_pol) <- "Activities: political parties, groups, associations"
# frq(df1$act_pol)

# frq(df1$v19)
df1$act_vol <- car::recode(df1$v19,recodes = "c(0,8,9)=NA")
df1$act_vol <- set_na(df1$act_vol,na = c(0,8,9))
df1$act_vol <- sjmisc::rec(x = df1$act_vol,rec = "rev")
sjlabelled::set_label(df1$act_vol) <- "Activities: charitable or religious - voluntary"
# frq(df1$act_vol)

# cor(df1[,c("act_lei","act_pol","act_vol")],use="complete.obs")
# psych::alpha(df1[,c("act_lei","act_pol","act_vol")])

df1$activities<- rowMeans(df1[,c("act_lei","act_pol","act_vol")],na.rm = T)
sjlabelled::set_label(df1$activities) <- "Participation in social activities"
# summary(df1$activities)
```

# Demographics

```{r demographics}
df1$agenum <- as.numeric(car::recode(df1$AGE,recodes = "999=NA",as.numeric = T))
sjlabelled::set_label(df1$agenum) <- "Age in years"
#frq(df1$agenum)

#frq(df1$SEX)
df1$female <- car::recode(as.numeric(df1$SEX),"2='Female';1='Male';9=NA",
                          levels = c('Male','Female'),as.factor = T)
#frq(df1$female)
```

# Save

```{r}
save(df1,file = here::here("input/data/proc/study1.RData"))
```