### 2023-10-20 procesamiento/prod_analysis-full-sample.Rmd
Update prod_analysis-full-sample.Rmd

---
title: "Multilevel regression models"
subtitle: "Full sample"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output: 
  html_document: 
    toc: yes
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(scipen=9999) # desactivar notacion cientifica
```

# Binary logit

## Full network

```{r mod-homclass-binary,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
df1$redistB <- car::recode(df1$redist,"4:5=1;1:3=0",as.factor = T)
dfreg <- df1 %>% dplyr::select(redist=redistB,
                        class3,
                        homclass,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  na.omit()

# Models 
base <- glmer(redist~1 + (1|country2),data=dfreg,weights = WEIGHT,family = "binomial"); icc1<- performance::icc(base)
homclass<- update(base, . ~ . +homclass)
full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
rob1 <- update(full1, . ~ . + class3+edyears+ Q03pcm+union+workst)
rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# anova(rob1,rob2)
# Interactions segregation x class
int_homo <- update(rob1, . ~ . +class3*homclass)

cap1 <- "Multilevel regression for redistributive preferences, network homogeneity and social class" 
coefmap <-  list(homclass ="Class-based network homogeneity",
                 'class3Intermediate class (III+IV)' ='Intermediate Class',
                 'class3Working Class (V+VI+VII)' ='Working Class',
                 "Q03pcmT02" = "Middle",
                 "Q03pcmT03" = "High",
                 "edyears" = 'Education in years',
                 "workstNot in paid work"="Not in labor force",
                 "unionYes" = 'Union membership: Yes',
                 'homclass:class3Intermediate class (III+IV)'='Homogeneity x Intermediate Class',
                 'homclass:class3Working Class (V+VI+VII)'='Homogeneity x Working Class'
                 )
coefgroup <- list("Social Class (ref: Service Class)"=2:3,
                  "Income Tercile (ref: Low)"=4:5,
                  "Homogeneity x Social Class"=9:10
                  )
gofrow <-  list(Controls=c("No","Yes","Yes","Yes"))
gof1<-  c(NA,NA,"Num. Countries:","Var: Group",NA,NA)

models <- list(homclass,full1,rob1,int_homo)
save( models,file = here::here("output/fullsample-logit-homclass-full-individual.RData"))
texreg::knitreg(models,caption=paste("(\\#tab:modredist)",cap1),caption.above=T,
                                float.pos="h!",
                  include.thresholds=FALSE,
                  custom.coef.map=coefmap,
                  groups = coefgroup,
                  custom.gof.rows=gofrow,
                  include.loglik = FALSE,include.aic = FALSE,
                  # custom.gof.names = gof1
                )
```

```{r mod-intermacro-binary,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,sjPlot,sjlabelled,haven,stringr,gridExtra)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
df1$redistB <- car::recode(df1$redist,"4:5=1;1:3=0",as.factor = T)
dfreg <- df1 %>% dplyr::select(redist=redistB,
                        class3,
                        homclass,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country,
                        palmaratio,
                        top10,
                        middle50,
                        d10d1,
                        giniindex,
                        ttheilge1
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  na.omit()

# Homoclass x Gini--------------------------------------------------------------
main <- "redist~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+(1|country2)"
base <- glmer(formula(main),data=dfreg,weights = WEIGHT,family = "binomial")
giniD_gdp<- update(base, . ~ . +gini_disp+logrgdpna)
giniD_rs <- update(giniD_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniD_gdp,giniD_rs)
giniD_int<- update(giniD_rs, . ~ . + homclass*gini_disp + (homclass|country2))
summary(giniD_int)

giniM_gdp<- update(base, . ~ . +gini_mkt+logrgdpna)
giniM_rs <- update(giniM_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniM_gdp,giniM_rs)
giniM_int <- update(giniM_rs, . ~ . + homclass*gini_mkt + (homclass|country2))
summary(giniM_int)
models <- list(giniD_gdp,giniD_rs,giniD_int,
               giniM_gdp,giniM_rs,giniM_int)

texreg::knitreg(models,
                caption=paste("(\\#tab:modinter1)","Cross-level interaction for Network homogeneity and Economic Inequality"),
                caption.above=T,float.pos="h!",
                include.thresholds=FALSE,
                custom.coef.map=list(homclass="Network Homogeneity",
                                     logrgdpna="log GDP","gini_disp"="Gini (Disposable) ",
                                     "homclass:gini_disp"="Homogeneity x Gini (D)",
                                     "gini_mkt"="Gini (Market) ",
                                     "homclass:gini_mkt"="Homogeneity x Gini (M)"),
                  groups = coefgroup,
                  custom.gof.rows=gofrow,
                  include.loglik = FALSE,include.aic = FALSE,
                  custom.gof.names = gof1
                )
# Homoclass x Other Economic Inequality-----------------------------------------
main <- "redist~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+logrgdpna+(homclass|country2)"
base <- glmer(formula(main),data=dfreg,weights = WEIGHT,family = "binomial")
int_palma <- update(base, . ~ . +homclass*palmaratio)
int_pc10  <- update(base, . ~ . +homclass*top10)
int_pc50  <- update(base, . ~ . +homclass*middle50)
int_d10d1 <- update(base, . ~ . +homclass*d10d1)
int_gini_wiid <- update(base, . ~ . +homclass*giniindex)
int_theil  <- update(base, . ~ . +homclass*ttheilge1)

models_m <- list(int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil
                 )
cap1=c("Cross-level interaction for Network homogeneity and Economic Inequality")
customcoef <- 
  list(
    homclass = "Class-based network homogeneity",
    logrgdpna = "log GDP",
    palmaratio = "Palma Ratio",
    "homclass:palmawd" = "Homogeneity x Palma Ratio",
    top10 = "Share National Income - top 10%",
    "homclass:top10"= "Homogeneity x top 10",
    middle50 = "Share National Income -  bottom 50%",
    "homclass:middle50" = "Homogeneity x bottom 50",
    d10d1 = "Ratio D10/D01",
    "homclass:d10d1" = "Homogeneity x Ratio D10/D01",
    giniindex = "Gini (World Bank)",
    "homclass:giniindex" = "Homogeneity x Gini (WB)",
    ttheilge1 = "Theil Index",
    "homclass:ttheilge1" = "Homogeneity x Theil Index"    
  )
texreg::knitreg(models_m,scalebox = 0.75,
                float.pos="h!",                
                custom.coef.map = customcoef,
                caption.above = T,caption = cap1)

save( list(giniD_gdp,giniD_rs,giniD_int,
                 giniM_gdp,giniM_rs,giniM_int,
                 int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil),
     file = here::here("output/fullsample-logit-homclass-full-macro.RData"))
```

## Strong ties

```{r mod-homclass-binary-strong-ties,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
df1$redistB <- car::recode(df1$redist,"4:5=1;1:3=0",as.factor = T)
dfreg <- df1 %>% dplyr::select(redist=redistB,
                        class3,
                        homclass=homclass_s,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  na.omit()

# Models 
base <- glmer(redist~1 + (1|country2),data=dfreg,weights = WEIGHT,family = "binomial"); icc1<- performance::icc(base)
homclass<- update(base, . ~ . +homclass)
full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
rob1 <- update(full1, . ~ . + class3+edyears+ Q03pcm+union+workst)
rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# anova(rob1,rob2)
# Interactions segregation x class
int_homo <- update(rob1, . ~ . +class3*homclass)

cap1 <- "Multilevel regression for redistributive preferences, network homogeneity and social class" 
coefmap <-  list(homclass ="Class-based network homogeneity",
                 'class3Intermediate class (III+IV)' ='Intermediate Class',
                 'class3Working Class (V+VI+VII)' ='Working Class',
                 "Q03pcmT02" = "Middle",
                 "Q03pcmT03" = "High",
                 "edyears" = 'Education in years',
                 "workstNot in paid work"="Not in labor force",
                 "unionYes" = 'Union membership: Yes',
                 'homclass:class3Intermediate class (III+IV)'='Homogeneity x Intermediate Class',
                 'homclass:class3Working Class (V+VI+VII)'='Homogeneity x Working Class'
                 )
coefgroup <- list("Social Class (ref: Service Class)"=2:3,
                  "Income Tercile (ref: Low)"=4:5,
                  "Homogeneity x Social Class"=9:10
                  )
gofrow <-  list(Controls=c("No","Yes","Yes","Yes"))
gof1<-  c(NA,NA,"Num. Countries:","Var: Group",NA,NA)

models <- list(homclass,full1,rob1,int_homo)
save( models,file = here::here("output/fullsample-logit-homclass-strong-individual.RData"))
texreg::knitreg(models,caption=paste("(\\#tab:modredist)",cap1),caption.above=T,
                                float.pos="h!",
                  include.thresholds=FALSE,
                  custom.coef.map=coefmap,
                  groups = coefgroup,
                  custom.gof.rows=gofrow,
                  include.loglik = FALSE,include.aic = FALSE,
                  # custom.gof.names = gof1
                )
```

```{r mod-intermacro-binary-strong-ties,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,sjPlot,sjlabelled,haven,stringr,gridExtra)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
df1$redistB <- car::recode(df1$redist,"4:5=1;1:3=0",as.factor = T)
dfreg <- df1 %>% dplyr::select(redist=redistB,
                        class3,
                        homclass=homclass_s,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country,
                        palmaratio,
                        top10,
                        middle50,
                        d10d1,
                        giniindex,
                        ttheilge1
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  na.omit()

# Homoclass x Gini--------------------------------------------------------------
main <- "redist~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+(1|country2)"
base <- glmer(formula(main),data=dfreg,weights = WEIGHT,family = "binomial")
giniD_gdp<- update(base, . ~ . +gini_disp+logrgdpna)
giniD_rs <- update(giniD_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniD_gdp,giniD_rs)
giniD_int<- update(giniD_rs, . ~ . + homclass*gini_disp + (homclass|country2))
summary(giniD_int)

giniM_gdp<- update(base, . ~ . +gini_mkt+logrgdpna)
giniM_rs <- update(giniM_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniM_gdp,giniM_rs)
giniM_int <- update(giniM_rs, . ~ . + homclass*gini_mkt + (homclass|country2))
summary(giniM_int)
models <- list(giniD_gdp,giniD_rs,giniD_int,
               giniM_gdp,giniM_rs,giniM_int)

texreg::knitreg(models,
                caption=paste("(\\#tab:modinter1)","Cross-level interaction for Network homogeneity and Economic Inequality"),
                caption.above=T,float.pos="h!",
                include.thresholds=FALSE,
                custom.coef.map=list(homclass="Network Homogeneity",
                                     logrgdpna="log GDP","gini_disp"="Gini (Disposable) ",
                                     "homclass:gini_disp"="Homogeneity x Gini (D)",
                                     "gini_mkt"="Gini (Market) ",
                                     "homclass:gini_mkt"="Homogeneity x Gini (M)"),
                  groups = coefgroup,
                  custom.gof.rows=gofrow,
                  include.loglik = FALSE,include.aic = FALSE,
                  custom.gof.names = gof1
                )

# Homoclass x Other Economic Inequality-----------------------------------------
main <- "redist~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+logrgdpna+(homclass|country2)"
base <- glmer(formula(main),data=dfreg,weights = WEIGHT,family = "binomial")
int_palma <- update(base, . ~ . +homclass*palmaratio)
int_pc10  <- update(base, . ~ . +homclass*top10)
int_pc50  <- update(base, . ~ . +homclass*middle50)
int_d10d1 <- update(base, . ~ . +homclass*d10d1)
int_gini_wiid <- update(base, . ~ . +homclass*giniindex)
int_theil  <- update(base, . ~ . +homclass*ttheilge1)

models_m <- list(int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil
                 )
cap1=c("Cross-level interaction for Network homogeneity and Economic Inequality")
customcoef <- 
  list(
    homclass = "Class-based network homogeneity",
    logrgdpna = "log GDP",
    palmaratio = "Palma Ratio",
    "homclass:palmawd" = "Homogeneity x Palma Ratio",
    top10 = "Share National Income - top 10%",
    "homclass:top10"= "Homogeneity x top 10",
    middle50 = "Share National Income -  bottom 50%",
    "homclass:middle50" = "Homogeneity x bottom 50",
    d10d1 = "Ratio D10/D01",
    "homclass:d10d1" = "Homogeneity x Ratio D10/D01",
    giniindex = "Gini (World Bank)",
    "homclass:giniindex" = "Homogeneity x Gini (WB)",
    ttheilge1 = "Theil Index",
    "homclass:ttheilge1" = "Homogeneity x Theil Index"    
  )
texreg::knitreg(models_m,scalebox = 0.75,
                float.pos="h!",                
                custom.coef.map = customcoef,
                caption.above = T,caption = cap1)

save( list(giniD_gdp,giniD_rs,giniD_int,
                 giniM_gdp,giniM_rs,giniM_int,
                 int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil),
     file = here::here("output/fullsample-logit-homclass-strong-macro.RData"))
```

## Weak ties

```{r mod-homclass-binary-weak-ties,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
df1$redistB <- car::recode(df1$redist,"4:5=1;1:3=0",as.factor = T)
dfreg <- df1 %>% dplyr::select(redist=redistB,
                        class3,
                        homclass=homclass_w,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  na.omit()

# Models 
base <- glmer(redist~1 + (1|country2),data=dfreg,weights = WEIGHT,family = "binomial"); icc1<- performance::icc(base)
homclass<- update(base, . ~ . +homclass)
full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
rob1 <- update(full1, . ~ . + class3+edyears+ Q03pcm+union+workst)
rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# anova(rob1,rob2)
# Interactions segregation x class
int_homo <- update(rob1, . ~ . +class3*homclass)

cap1 <- "Multilevel regression for redistributive preferences, network homogeneity and social class" 
coefmap <-  list(homclass ="Class-based network homogeneity",
                 'class3Intermediate class (III+IV)' ='Intermediate Class',
                 'class3Working Class (V+VI+VII)' ='Working Class',
                 "Q03pcmT02" = "Middle",
                 "Q03pcmT03" = "High",
                 "edyears" = 'Education in years',
                 "workstNot in paid work"="Not in labor force",
                 "unionYes" = 'Union membership: Yes',
                 'homclass:class3Intermediate class (III+IV)'='Homogeneity x Intermediate Class',
                 'homclass:class3Working Class (V+VI+VII)'='Homogeneity x Working Class'
                 )
coefgroup <- list("Social Class (ref: Service Class)"=2:3,
                  "Income Tercile (ref: Low)"=4:5,
                  "Homogeneity x Social Class"=9:10
                  )
gofrow <-  list(Controls=c("No","Yes","Yes","Yes"))
gof1<-  c(NA,NA,"Num. Countries:","Var: Group",NA,NA)

models <- list(homclass,full1,rob1,int_homo)
save( models,file = here::here("output/fullsample-logit-homclass-weak-individual.RData"))
texreg::knitreg(models,caption=paste("(\\#tab:modredist)",cap1),caption.above=T,
                                float.pos="h!",
                  include.thresholds=FALSE,
                  custom.coef.map=coefmap,
                  groups = coefgroup,
                  custom.gof.rows=gofrow,
                  include.loglik = FALSE,include.aic = FALSE,
                  # custom.gof.names = gof1
                )
```

```{r mod-intermacro-binary-weak-ties,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,sjPlot,sjlabelled,haven,stringr,gridExtra)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
df1$redistB <- car::recode(df1$redist,"4:5=1;1:3=0",as.factor = T)
dfreg <- df1 %>% dplyr::select(redist=redistB,
                        class3,
                        homclass=homclass_w,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country,
                        palmaratio,
                        top10,
                        middle50,
                        d10d1,
                        giniindex,
                        ttheilge1
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  na.omit()

# Homoclass x Gini--------------------------------------------------------------
main <- "redist~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+(1|country2)"
base <- glmer(formula(main),data=dfreg,weights = WEIGHT,family = "binomial")
giniD_gdp<- update(base, . ~ . +gini_disp+logrgdpna)
giniD_rs <- update(giniD_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniD_gdp,giniD_rs)
giniD_int<- update(giniD_rs, . ~ . + homclass*gini_disp + (homclass|country2))
summary(giniD_int)

giniM_gdp<- update(base, . ~ . +gini_mkt+logrgdpna)
giniM_rs <- update(giniM_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniM_gdp,giniM_rs)
giniM_int <- update(giniM_rs, . ~ . + homclass*gini_mkt + (homclass|country2))
summary(giniM_int)
models <- list(giniD_gdp,giniD_rs,giniD_int,
               giniM_gdp,giniM_rs,giniM_int)

texreg::knitreg(models,
                caption=paste("(\\#tab:modinter1)","Cross-level interaction for Network homogeneity and Economic Inequality"),
                caption.above=T,float.pos="h!",
                include.thresholds=FALSE,
                custom.coef.map=list(homclass="Network Homogeneity",
                                     logrgdpna="log GDP","gini_disp"="Gini (Disposable) ",
                                     "homclass:gini_disp"="Homogeneity x Gini (D)",
                                     "gini_mkt"="Gini (Market) ",
                                     "homclass:gini_mkt"="Homogeneity x Gini (M)"),
                  groups = coefgroup,
                  custom.gof.rows=gofrow,
                  include.loglik = FALSE,include.aic = FALSE,
                  custom.gof.names = gof1
                )

# Homoclass x Other Economic Inequality-----------------------------------------
main <- "redist~class3+homclass+edyears+Q03pcm+workst+union+female+agenum+age2+logrgdpna+(homclass|country2)"
base <- glmer(formula(main),data=dfreg,weights = WEIGHT,family = "binomial")
int_palma <- update(base, . ~ . +homclass*palmaratio)
int_pc10  <- update(base, . ~ . +homclass*top10)
int_pc50  <- update(base, . ~ . +homclass*middle50)
int_d10d1 <- update(base, . ~ . +homclass*d10d1)
int_gini_wiid <- update(base, . ~ . +homclass*giniindex)
int_theil  <- update(base, . ~ . +homclass*ttheilge1)

models_m <- list(int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil
                 )
cap1=c("Cross-level interaction for Network homogeneity and Economic Inequality")
customcoef <- 
  list(
    homclass = "Class-based network homogeneity",
    logrgdpna = "log GDP",
    palmaratio = "Palma Ratio",
    "homclass:palmawd" = "Homogeneity x Palma Ratio",
    top10 = "Share National Income - top 10%",
    "homclass:top10"= "Homogeneity x top 10",
    middle50 = "Share National Income -  bottom 50%",
    "homclass:middle50" = "Homogeneity x bottom 50",
    d10d1 = "Ratio D10/D01",
    "homclass:d10d1" = "Homogeneity x Ratio D10/D01",
    giniindex = "Gini (World Bank)",
    "homclass:giniindex" = "Homogeneity x Gini (WB)",
    ttheilge1 = "Theil Index",
    "homclass:ttheilge1" = "Homogeneity x Theil Index"    
  )
texreg::knitreg(models_m,scalebox = 0.75,
                float.pos="h!",                
                custom.coef.map = customcoef,
                caption.above = T,caption = cap1)

save( list(giniD_gdp,giniD_rs,giniD_int,
                 giniM_gdp,giniM_rs,giniM_int,
                 int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil),
     file = here::here("output/fullsample-logit-homclass-weak-macro.RData"))
```

# Ordinal

## Full network

```{r mod-homclass-ordinal,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,haven,ordinal,brms)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2

df1$redistOR <- ordered(x = df1$redist)
dfreg <- df1 %>% dplyr::select(redistOR,
                        class3,
                        homclass,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  na.omit()

# Models 
# base <- clmm(redistF~1 + (1|country2),data=dfreg,weights = WEIGHT); icc1<- performance::icc(base)
# homclass<- update(base, . ~ . +homclass)
# full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
# rob1 <- update(full1, . ~ . + edyears+ Q03pcm+union+workst)
# rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# Bayesian ordinal logit
# brms::brm()
base <- brm(redistOR~1 + (1|country2),data=dfreg,family=cumulative("logit"))
homclass<- update(base, . ~ . +homclass)
full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
rob1 <- update(full1, . ~ . + edyears+ Q03pcm+union+workst)
rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# Interactions segregation x class
int_homo <- update(rob1, . ~ . +class3*homclass)

cap1 <- "Multilevel regression for redistributive preferences, network homogeneity and social class" 
coefmap <-  list(homclass ="Class-based network homogeneity",
                 'class3Intermediate class (III+IV)' ='Intermediate Class',
                 'class3Working Class (V+VI+VII)' ='Working Class',
                 "Q03pcmT02" = "Middle",
                 "Q03pcmT03" = "High",
                 "edyears" = 'Education in years',
                 "workstNot in paid work"="Not in labor force",
                 "unionYes" = 'Union membership: Yes',
                 'homclass:class3Intermediate class (III+IV)'='Homogeneity x Intermediate Class',
                 'homclass:class3Working Class (V+VI+VII)'='Homogeneity x Working Class'
                 )
coefgroup <- list("Social Class (ref: Service Class)"=2:3,
                  "Income Tercile (ref: Low)"=4:5,
                  "Homogeneity x Social Class"=9:10
                  )
gofrow <-  list(Controls=c("No","Yes","Yes","Yes"))
gof1<-  c(NA,NA,"Num. Countries:","Var: Group",NA,NA)

models <- list(homclass,full1,rob1,int_homo)
save( models,file = here::here("output/fullsample-ologit-homclass-full-individual.RData"))
texreg::knitreg(models,caption=paste("(\\#tab:modredist)",cap1),caption.above=T,
                                float.pos="h!",
                  include.thresholds=FALSE,
                  custom.coef.map=coefmap,
                  groups = coefgroup,
                  custom.gof.rows=gofrow,
                  include.loglik = FALSE,include.aic = FALSE,
                  # custom.gof.names = gof1
                )
```

```{r mod-intermacro-ordinal,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,sjPlot,sjlabelled,haven,stringr,gridExtra)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
df1$redistOR <- ordered(x = df1$redist)
dfreg <- df1 %>% dplyr::select(redistOR,
                        class3,
                        homclass,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country,
                        palmaratio,
                        top10,
                        middle50,
                        d10d1,
                        giniindex,
                        ttheilge1
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  na.omit()

# Homoclass x Gini--------------------------------------------------------------
main <- "redist~class3+ homclass+ edyears+ Q03pcm+ workst+ union+ female+agenum +age2+ (1|country2)"
base <- clmm(formula(main),data=dfreg,weights = WEIGHT,Hess = T)
giniD_gdp<- update(base, . ~ . +gini_disp+logrgdpna)
giniD_rs <- update(giniD_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniD_gdp,giniD_rs)
giniD_int<- update(giniD_rs, . ~ . + homclass*gini_disp + (homclass|country2))
# summary(giniD_int)

giniM_gdp<- update(base, . ~ . +gini_mkt+logrgdpna)
giniM_rs <- update(giniM_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniM_gdp,giniM_rs)
giniM_int <- update(giniM_rs, . ~ . + homclass*gini_mkt + (homclass|country2))
summary(giniM_int)
models <- list(giniD_gdp,giniD_rs,giniD_int,
               giniM_gdp,giniM_rs,giniM_int)

texreg::knitreg(models,
                caption=paste("(\\#tab:modinter1)","Cross-level interaction for Network homogeneity and Economic Inequality"),
                caption.above=T,float.pos="h!",
                include.thresholds=FALSE,
                custom.coef.map=list(homclass="Network Homogeneity",
                                     logrgdpna="log GDP","gini_disp"="Gini (Disposable) ",
                                     "homclass:gini_disp"="Homogeneity x Gini (D)",
                                     "gini_mkt"="Gini (Market) ",
                                     "homclass:gini_mkt"="Homogeneity x Gini (M)"),
                  # groups = coefgroup,
                  # custom.gof.rows=gofrow,
                  include.loglik = FALSE,include.aic = FALSE,
                  # custom.gof.names = gof1
                )

# Homoclass x Other Economic Inequality-----------------------------------------
main <- "redist~class3+ homclass+ edyears+ Q03pcm+ workst+ union+ female+agenum +age2+ logrgdpna +(homclass|country2)"
base <- clmm(redistOR,data=dfreg,weights = WEIGHT)
int_palma <- update(base, . ~ . +homclass*palmaratio)
int_pc10  <- update(base, . ~ . +homclass*top10)
int_pc50  <- update(base, . ~ . +homclass*middle50)
int_d10d1 <- update(base, . ~ . +homclass*d10d1)
int_gini_wiid <- update(base, . ~ . +homclass*giniindex)
int_theil  <- update(base, . ~ . +homclass*ttheilge1)

models_m <- list(int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil
                 )

cap1=c("Cross-level interaction for Network homogeneity and Economic Inequality")
customcoef <- 
  list(
    homclass = "Class-based network homogeneity",
    logrgdpna = "log GDP",
    palmaratio = "Palma Ratio",
    "homclass:palmawd" = "Homogeneity x Palma Ratio",
    top10 = "Share National Income - top 10%",
    "homclass:top10"= "Homogeneity x top 10",
    middle50 = "Share National Income -  bottom 50%",
    "homclass:middle50" = "Homogeneity x bottom 50",
    d10d1 = "Ratio D10/D01",
    "homclass:d10d1" = "Homogeneity x Ratio D10/D01",
    giniindex = "Gini (World Bank)",
    "homclass:giniindex" = "Homogeneity x Gini (WB)",
    ttheilge1 = "Theil Index",
    "homclass:ttheilge1" = "Homogeneity x Theil Index"    
  )
texreg::knitreg(models_m,scalebox = 0.75,
                float.pos="h!",                
                custom.coef.map = customcoef,
                caption.above = T,caption = cap1)
save( list(giniD_gdp,giniD_rs,giniD_int,
                 giniM_gdp,giniM_rs,giniM_int,
                 int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil),
     file = here::here("output/fullsample-ologit-homclass-full-macro.RData"))
```

## Strong ties

```{r mod-homclass-ordinal-strong-ties,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,haven,ordinal,brms)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2

df1$redistOR <- ordered(x = df1$redist)
dfreg <- df1 %>% dplyr::select(redistOR,
                        class3,
                        homclass=homclass_s,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  na.omit()

# Models 
# base <- clmm(redistF~1 + (1|country2),data=dfreg,weights = WEIGHT); icc1<- performance::icc(base)
# homclass<- update(base, . ~ . +homclass)
# full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
# rob1 <- update(full1, . ~ . + edyears+ Q03pcm+union+workst)
# rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# Bayesian ordinal logit
# brms::brm()
base <- brm(redistOR~1 + (1|country2),data=dfreg,family=cumulative("logit"))
homclass<- update(base, . ~ . +homclass)
full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
rob1 <- update(full1, . ~ . + edyears+ Q03pcm+union+workst)
rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# Interactions segregation x class
int_homo <- update(rob1, . ~ . +class3*homclass)

cap1 <- "Multilevel regression for redistributive preferences, network homogeneity and social class" 
coefmap <-  list(homclass ="Class-based network homogeneity",
                 'class3Intermediate class (III+IV)' ='Intermediate Class',
                 'class3Working Class (V+VI+VII)' ='Working Class',
                 "Q03pcmT02" = "Middle",
                 "Q03pcmT03" = "High",
                 "edyears" = 'Education in years',
                 "workstNot in paid work"="Not in labor force",
                 "unionYes" = 'Union membership: Yes',
                 'homclass:class3Intermediate class (III+IV)'='Homogeneity x Intermediate Class',
                 'homclass:class3Working Class (V+VI+VII)'='Homogeneity x Working Class'
                 )
coefgroup <- list("Social Class (ref: Service Class)"=2:3,
                  "Income Tercile (ref: Low)"=4:5,
                  "Homogeneity x Social Class"=9:10
                  )
gofrow <-  list(Controls=c("No","Yes","Yes","Yes"))
gof1<-  c(NA,NA,"Num. Countries:","Var: Group",NA,NA)

models <- list(homclass,full1,rob1,int_homo)
save( models,file = here::here("output/fullsample-ologit-homclass-strong-individual.RData"))
texreg::knitreg(models,caption=paste("(\\#tab:modredist)",cap1),caption.above=T,
                                float.pos="h!",
                  include.thresholds=FALSE,
                  custom.coef.map=coefmap,
                  groups = coefgroup,
                  custom.gof.rows=gofrow,
                  include.loglik = FALSE,include.aic = FALSE,
                  # custom.gof.names = gof1
                )
```

```{r mod-intermacro-ordinal-strong-ties,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,sjPlot,sjlabelled,haven,stringr,gridExtra)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
df1$redistOR <- ordered(x = df1$redist)
dfreg <- df1 %>% dplyr::select(redistOR,
                        class3,
                        homclass=homclass_s,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country,
                        palmaratio,
                        top10,
                        middle50,
                        d10d1,
                        giniindex,
                        ttheilge1
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  na.omit()

# Homoclass x Gini--------------------------------------------------------------
main <- "redist~class3+ homclass+ edyears+ Q03pcm+ workst+ union+ female+agenum +age2+ (1|country2)"
base <- clmm(formula(main),data=dfreg,weights = WEIGHT,Hess = T)
giniD_gdp<- update(base, . ~ . +gini_disp+logrgdpna)
giniD_rs <- update(giniD_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniD_gdp,giniD_rs)
giniD_int<- update(giniD_rs, . ~ . + homclass*gini_disp + (homclass|country2))
# summary(giniD_int)

giniM_gdp<- update(base, . ~ . +gini_mkt+logrgdpna)
giniM_rs <- update(giniM_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniM_gdp,giniM_rs)
giniM_int <- update(giniM_rs, . ~ . + homclass*gini_mkt + (homclass|country2))
summary(giniM_int)
models <- list(giniD_gdp,giniD_rs,giniD_int,
               giniM_gdp,giniM_rs,giniM_int)

texreg::knitreg(models,
                caption=paste("(\\#tab:modinter1)","Cross-level interaction for Network homogeneity and Economic Inequality"),
                caption.above=T,float.pos="h!",
                include.thresholds=FALSE,
                custom.coef.map=list(homclass="Network Homogeneity",
                                     logrgdpna="log GDP","gini_disp"="Gini (Disposable) ",
                                     "homclass:gini_disp"="Homogeneity x Gini (D)",
                                     "gini_mkt"="Gini (Market) ",
                                     "homclass:gini_mkt"="Homogeneity x Gini (M)"),
                  # groups = coefgroup,
                  # custom.gof.rows=gofrow,
                  include.loglik = FALSE,include.aic = FALSE,
                  # custom.gof.names = gof1
                )

# Homoclass x Other Economic Inequality-----------------------------------------
main <- "redist~class3+ homclass+ edyears+ Q03pcm+ workst+ union+ female+agenum +age2+ logrgdpna +(homclass|country2)"
base <- clmm(redistOR,data=dfreg,weights = WEIGHT)
int_palma <- update(base, . ~ . +homclass*palmaratio)
int_pc10  <- update(base, . ~ . +homclass*top10)
int_pc50  <- update(base, . ~ . +homclass*middle50)
int_d10d1 <- update(base, . ~ . +homclass*d10d1)
int_gini_wiid <- update(base, . ~ . +homclass*giniindex)
int_theil  <- update(base, . ~ . +homclass*ttheilge1)

models_m <- list(int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil
                 )

cap1=c("Cross-level interaction for Network homogeneity and Economic Inequality")
customcoef <- 
  list(
    homclass = "Class-based network homogeneity",
    logrgdpna = "log GDP",
    palmaratio = "Palma Ratio",
    "homclass:palmawd" = "Homogeneity x Palma Ratio",
    top10 = "Share National Income - top 10%",
    "homclass:top10"= "Homogeneity x top 10",
    middle50 = "Share National Income -  bottom 50%",
    "homclass:middle50" = "Homogeneity x bottom 50",
    d10d1 = "Ratio D10/D01",
    "homclass:d10d1" = "Homogeneity x Ratio D10/D01",
    giniindex = "Gini (World Bank)",
    "homclass:giniindex" = "Homogeneity x Gini (WB)",
    ttheilge1 = "Theil Index",
    "homclass:ttheilge1" = "Homogeneity x Theil Index"    
  )
texreg::knitreg(models_m,scalebox = 0.75,
                float.pos="h!",                
                custom.coef.map = customcoef,
                caption.above = T,caption = cap1)

save( list(giniD_gdp,giniD_rs,giniD_int,
                 giniM_gdp,giniM_rs,giniM_int,
                 int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil),
     file = here::here("output/fullsample-ologit-homclass-strong-macro.RData"))
```

## Weak ties

```{r mod-homclass-ordinal-weak-ties,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,haven,ordinal,brms)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2

df1$redistOR <- ordered(x = df1$redist)
dfreg <- df1 %>% dplyr::select(redistOR,
                        class3,
                        homclass=homclass_w,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  na.omit()

# Models 
# base <- clmm(redistF~1 + (1|country2),data=dfreg,weights = WEIGHT); icc1<- performance::icc(base)
# homclass<- update(base, . ~ . +homclass)
# full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
# rob1 <- update(full1, . ~ . + edyears+ Q03pcm+union+workst)
# rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# Bayesian ordinal logit
# brms::brm()
base <- brm(redistOR~1 + (1|country2),data=dfreg,family=cumulative("logit"))
homclass<- update(base, . ~ . +homclass)
full1 <- update(homclass, . ~ . + class3+female+agenum+age2)
rob1 <- update(full1, . ~ . + edyears+ Q03pcm+union+workst)
rob2 <- update(rob1, . ~ . -(1|country2) + (homclass|country2))
# Interactions segregation x class
int_homo <- update(rob1, . ~ . +class3*homclass)

cap1 <- "Multilevel regression for redistributive preferences, network homogeneity and social class" 
coefmap <-  list(homclass ="Class-based network homogeneity",
                 'class3Intermediate class (III+IV)' ='Intermediate Class',
                 'class3Working Class (V+VI+VII)' ='Working Class',
                 "Q03pcmT02" = "Middle",
                 "Q03pcmT03" = "High",
                 "edyears" = 'Education in years',
                 "workstNot in paid work"="Not in labor force",
                 "unionYes" = 'Union membership: Yes',
                 'homclass:class3Intermediate class (III+IV)'='Homogeneity x Intermediate Class',
                 'homclass:class3Working Class (V+VI+VII)'='Homogeneity x Working Class'
                 )
coefgroup <- list("Social Class (ref: Service Class)"=2:3,
                  "Income Tercile (ref: Low)"=4:5,
                  "Homogeneity x Social Class"=9:10
                  )
gofrow <-  list(Controls=c("No","Yes","Yes","Yes"))
gof1<-  c(NA,NA,"Num. Countries:","Var: Group",NA,NA)

models <- list(homclass,full1,rob1,int_homo)
save( models,file = here::here("output/fullsample-ologit-homclass-weak-individual.RData"))
texreg::knitreg(models,caption=paste("(\\#tab:modredist)",cap1),caption.above=T,
                                float.pos="h!",
                  include.thresholds=FALSE,
                  custom.coef.map=coefmap,
                  groups = coefgroup,
                  custom.gof.rows=gofrow,
                  include.loglik = FALSE,include.aic = FALSE,
                  # custom.gof.names = gof1
                )
```

```{r mod-intermacro-ordinal-weak-ties,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,sjPlot,sjlabelled,haven,stringr,gridExtra)  
rm(list=ls())
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
df1$redistOR <- ordered(x = df1$redist)
dfreg <- df1 %>% dplyr::select(redistOR,
                        class3,
                        homclass=homclass_w,
                        Q03pcm,
                        edyears,
                        female,
                        agenum,
                        union,
                        workst,
                        WEIGHT,
                        region,
                        "gini_disp",
                        "gini_mkt",
                        rgdpna,
                        oecd,
                        country2,country,
                        palmaratio,
                        top10,
                        middle50,
                        d10d1,
                        giniindex,
                        ttheilge1
                        ) %>% 
  mutate(logrgdpna=log(rgdpna),
         edyears2=edyears^2,
         age2=agenum^2) %>%
  na.omit()

# Homoclass x Gini--------------------------------------------------------------
main <- "redist~class3+ homclass+ edyears+ Q03pcm+ workst+ union+ female+agenum +age2+ (1|country2)"
base <- clmm(formula(main),data=dfreg,weights = WEIGHT,Hess = T)
giniD_gdp<- update(base, . ~ . +gini_disp+logrgdpna)
giniD_rs <- update(giniD_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniD_gdp,giniD_rs)
giniD_int<- update(giniD_rs, . ~ . + homclass*gini_disp + (homclass|country2))
# summary(giniD_int)

giniM_gdp<- update(base, . ~ . +gini_mkt+logrgdpna)
giniM_rs <- update(giniM_gdp, . ~ . - (1|country2) + (homclass|country2))
# anova(giniM_gdp,giniM_rs)
giniM_int <- update(giniM_rs, . ~ . + homclass*gini_mkt + (homclass|country2))
summary(giniM_int)
models <- list(giniD_gdp,giniD_rs,giniD_int,
               giniM_gdp,giniM_rs,giniM_int)

texreg::knitreg(models,
                caption=paste("(\\#tab:modinter1)","Cross-level interaction for Network homogeneity and Economic Inequality"),
                caption.above=T,float.pos="h!",
                include.thresholds=FALSE,
                custom.coef.map=list(homclass="Network Homogeneity",
                                     logrgdpna="log GDP","gini_disp"="Gini (Disposable) ",
                                     "homclass:gini_disp"="Homogeneity x Gini (D)",
                                     "gini_mkt"="Gini (Market) ",
                                     "homclass:gini_mkt"="Homogeneity x Gini (M)"),
                  # groups = coefgroup,
                  # custom.gof.rows=gofrow,
                  include.loglik = FALSE,include.aic = FALSE,
                  # custom.gof.names = gof1
                )

# Homoclass x Other Economic Inequality-----------------------------------------
main <- "redist~class3+ homclass+ edyears+ Q03pcm+ workst+ union+ female+agenum +age2+ logrgdpna +(homclass|country2)"
base <- clmm(redistOR,data=dfreg,weights = WEIGHT)
int_palma <- update(base, . ~ . +homclass*palmaratio)
int_pc10  <- update(base, . ~ . +homclass*top10)
int_pc50  <- update(base, . ~ . +homclass*middle50)
int_d10d1 <- update(base, . ~ . +homclass*d10d1)
int_gini_wiid <- update(base, . ~ . +homclass*giniindex)
int_theil  <- update(base, . ~ . +homclass*ttheilge1)

models_m <- list(int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil
                 )

cap1=c("Cross-level interaction for Network homogeneity and Economic Inequality")
customcoef <- 
  list(
    homclass = "Class-based network homogeneity",
    logrgdpna = "log GDP",
    palmaratio = "Palma Ratio",
    "homclass:palmawd" = "Homogeneity x Palma Ratio",
    top10 = "Share National Income - top 10%",
    "homclass:top10"= "Homogeneity x top 10",
    middle50 = "Share National Income -  bottom 50%",
    "homclass:middle50" = "Homogeneity x bottom 50",
    d10d1 = "Ratio D10/D01",
    "homclass:d10d1" = "Homogeneity x Ratio D10/D01",
    giniindex = "Gini (World Bank)",
    "homclass:giniindex" = "Homogeneity x Gini (WB)",
    ttheilge1 = "Theil Index",
    "homclass:ttheilge1" = "Homogeneity x Theil Index"    
  )
texreg::knitreg(models_m,scalebox = 0.75,
                float.pos="h!",                
                custom.coef.map = customcoef,
                caption.above = T,caption = cap1)

save( list(giniD_gdp,giniD_rs,giniD_int,
                 giniM_gdp,giniM_rs,giniM_int,
                 int_palma,int_pc10,int_pc50,int_d10d1,
                 int_gini_wiid,int_theil),
     file = here::here("output/fullsample-ologit-homclass-weak-macro.RData"))
```