### 2023-10-30 procesamiento/prod_analysis-tables-macro.Rmd
lunes

---
title: "Additional analyses"
subtitle: "Cross-level interactions"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/css/custom.css" # custom CSS para html
output: 
  html_document: 
    toc: false
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(scipen=9999) # desactivar notacion cientifica
```

```{css}
body {
    text-align: initial;
    font-size: 20px;
    font-family: "Times New Roman", Times, serif;
}
```

## All countries

```{r full-lmer-macro,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())
coef_map <-
  list(
    homclass = "Homogeneity",
    logrgdpna = "log GDP",
    "gini_disp" = "GiniD ",
    "homclass:gini_disp" = "Hom:Gini(D)",
    "gini_mkt" = "Gini(M) ",
    "homclass:gini_mkt" = "Hom:Gini(M)",
    palmaratio               = "Palma",
    "homclass:palmawd"      = "Hom:Palma",
    top10             = "Top 10%",
    "homclass:top10"    = "Hom:top10",
    middle50             = "Bott.50%",
    "homclass:middle50"    = "Hom:bott50",
    d10d1             = "D10/D01",
    "homclass:d10d1"    = "Hom:D10D01",
    giniindex             = "Gini(WIID)",
    "homclass:giniindex"    = "Hom:Gini(WD)",
    ttheilge1             = "Theil",
    "homclass:ttheilge1"    = "Hom:Theil"
  )

load(here::here("output/fullsample-numeric-homclass-full-macro.RData"));fsample_num_full_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)
load(here::here("output/fullsample-numeric-homclass-strong-macro.RData"));fsample_num_strong_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)
load(here::here("output/fullsample-numeric-homclass-weak-macro.RData")); fsample_num_weak_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)

htmlreg(
  c(fsample_num_full_mac,fsample_num_strong_mac,fsample_num_weak_mac),
  caption = "Multilevel linear regression: full country sample cross-level interaction",
  caption.above = T,
  custom.coef.map= coef_map,
  custom.header = 
    list(
      "Full network"=1:8,
      "Strong ties"=9:16,
      "Weak ties"=17:24
      ))
```

```{r full-clmm-macro,results='asis'}
htmltools::includeHTML("../output/fullsample-ologit-homclass-full-strong-weak-macro.html")
```

```{r full-lmer-macro,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())
coef_map <-
  list(
    homclass = "Homogeneity",
    logrgdpna = "log GDP",
    "gini_disp" = "GiniD ",
    "homclass:gini_disp" = "Hom:Gini(D)",
    "gini_mkt" = "Gini(M) ",
    "homclass:gini_mkt" = "Hom:Gini(M)",
    palmaratio               = "Palma",
    "homclass:palmawd"      = "Hom:Palma",
    top10             = "Top 10%",
    "homclass:top10"    = "Hom:top10",
    middle50             = "Bott.50%",
    "homclass:middle50"    = "Hom:bott50",
    d10d1             = "D10/D01",
    "homclass:d10d1"    = "Hom:D10D01",
    giniindex             = "Gini(WIID)",
    "homclass:giniindex"    = "Hom:Gini(WD)",
    ttheilge1             = "Theil",
    "homclass:ttheilge1"    = "Hom:Theil"
  )

load(here::here("output/fullsample-egalitarian-homclass-full-macro.RData"));fsample_egal_full_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)
load(here::here("output/fullsample-egalitarian-homclass-strong-macro.RData"));fsample_egal_strong_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)
load(here::here("output/fullsample-egalitarian-homclass-weak-macro.RData")); fsample_egal_weak_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)

htmlreg(
  c(fsample_egal_full_mac,fsample_egal_strong_mac,fsample_egal_weak_mac),
  caption = "Multilevel linear regression: full country sample cross-level interaction - egalitarian index",
  caption.above = T,
  custom.coef.map= coef_map,
  custom.header = 
    list(
      "Full network"=1:8,
      "Strong ties"=9:16,
      "Weak ties"=17:24
      ))
```

## Non-confusian

```{r nonconfusian-lmer-macro,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())
load(here::here("output/nonconfusian-numeric-homclass-full-macro.RData")); nconf_num_full_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)
load(here::here("output/nonconfusian-numeric-homclass-strong-macro.RData"));nconf_num_strong_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)
load(here::here("output/nonconfusian-numeric-homclass-weak-macro.RData")); nconf_num_weak_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)

coef_map <-
  list(
    homclass = "Homogeneity",
    logrgdpna = "log GDP",
    "gini_disp" = "GiniD ",
    "homclass:gini_disp" = "Hom:Gini(D)",
    "gini_mkt" = "Gini(M) ",
    "homclass:gini_mkt" = "Hom:Gini(M)",
    palmaratio               = "Palma",
    "homclass:palmawd"      = "Hom:Palma",
    top10             = "Top 10%",
    "homclass:top10"    = "Hom:top10",
    middle50             = "Bott.50%",
    "homclass:middle50"    = "Hom:bott50",
    d10d1             = "D10/D01",
    "homclass:d10d1"    = "Hom:D10D01",
    giniindex             = "Gini(WIID)",
    "homclass:giniindex"    = "Hom:Gini(WD)",
    ttheilge1             = "Theil",
    "homclass:ttheilge1"    = "Hom:Theil"
  )

htmlreg(
  c(nconf_num_full_mac,nconf_num_strong_mac,nconf_num_weak_mac),
  caption = "Multilevel linear regression: Non-confusian country sample cross-level interaction",
  caption.above = T,
  custom.coef.map= coef_map,
  custom.header = 
    list(
      "Full network"=1:8,
      "Strong ties"=9:16,
      "Weak ties"=17:24
      ))
```

```{r nonconfusian-clmm-macro,results='asis'}

```

```{r nonconfusian-lmer-egalitarian-macro,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())
coef_map <-
  list(
    homclass = "Homogeneity",
    logrgdpna = "log GDP",
    "gini_disp" = "GiniD ",
    "homclass:gini_disp" = "Hom:Gini(D)",
    "gini_mkt" = "Gini(M) ",
    "homclass:gini_mkt" = "Hom:Gini(M)",
    palmaratio               = "Palma",
    "homclass:palmawd"      = "Hom:Palma",
    top10             = "Top 10%",
    "homclass:top10"    = "Hom:top10",
    middle50             = "Bott.50%",
    "homclass:middle50"    = "Hom:bott50",
    d10d1             = "D10/D01",
    "homclass:d10d1"    = "Hom:D10D01",
    giniindex             = "Gini(WIID)",
    "homclass:giniindex"    = "Hom:Gini(WD)",
    ttheilge1             = "Theil",
    "homclass:ttheilge1"    = "Hom:Theil"
  )

load(here::here("output/nonconfusian-egalitarian-homclass-full-macro.RData"));nconf_num_full_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)
load(here::here("output/nonconfusian-egalitarian-homclass-strong-macro.RData"));nconf_num_strong_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)
load(here::here("output/nonconfusian-egalitarian-homclass-weak-macro.RData")); nconf_num_weak_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)

htmlreg(
  c(nconf_num_full_mac,nconf_num_strong_mac,nconf_num_weak_mac),
  caption = "Multilevel linear regression: full country sample cross-level interaction - egalitarian index",
  caption.above = T,
  custom.coef.map= coef_map,
  custom.header = 
    list(
      "Full network"=1:8,
      "Strong ties"=9:16,
      "Weak ties"=17:24
      ))
```

## OECD

```{r ocde-lmer-macro,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())
load(here::here("output/ocde-numeric-homclass-full-macro.RData")); ocde_num_full_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)
load(here::here("output/ocde-numeric-homclass-strong-macro.RData"));ocde_num_strong_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)
load(here::here("output/ocde-numeric-homclass-weak-macro.RData")); ocde_num_weak_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)

coef_map <-
  list(
    homclass = "Homogeneity",
    logrgdpna = "log GDP",
    "gini_disp" = "GiniD ",
    "homclass:gini_disp" = "Hom:Gini(D)",
    "gini_mkt" = "Gini(M) ",
    "homclass:gini_mkt" = "Hom:Gini(M)",
    palmaratio               = "Palma",
    "homclass:palmawd"      = "Hom:Palma",
    top10             = "Top 10%",
    "homclass:top10"    = "Hom:top10",
    middle50             = "Bott.50%",
    "homclass:middle50"    = "Hom:bott50",
    d10d1             = "D10/D01",
    "homclass:d10d1"    = "Hom:D10D01",
    giniindex             = "Gini(WIID)",
    "homclass:giniindex"    = "Hom:Gini(WD)",
    ttheilge1             = "Theil",
    "homclass:ttheilge1"    = "Hom:Theil"
  )

htmlreg(
  c(ocde_num_full_mac,ocde_num_strong_mac,ocde_num_weak_mac),
  caption = "Multilevel linear regression: OECD country sample cross-level interaction",
  custom.coef.map= coef_map,
  caption.above = T,
  custom.header = 
    list(
      "Full network"=1:8,
      "Strong ties"=9:16,
      "Weak ties"=17:24
      ))
```

```{r oecd-clmm-macro,results='asis'}

```

```{r ocde-lmer-egalitarian-macro,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())
coef_map <-
  list(
    homclass = "Homogeneity",
    logrgdpna = "log GDP",
    "gini_disp" = "GiniD ",
    "homclass:gini_disp" = "Hom:Gini(D)",
    "gini_mkt" = "Gini(M) ",
    "homclass:gini_mkt" = "Hom:Gini(M)",
    palmaratio               = "Palma",
    "homclass:palmawd"      = "Hom:Palma",
    top10             = "Top 10%",
    "homclass:top10"    = "Hom:top10",
    middle50             = "Bott.50%",
    "homclass:middle50"    = "Hom:bott50",
    d10d1             = "D10/D01",
    "homclass:d10d1"    = "Hom:D10D01",
    giniindex             = "Gini(WIID)",
    "homclass:giniindex"    = "Hom:Gini(WD)",
    ttheilge1             = "Theil",
    "homclass:ttheilge1"    = "Hom:Theil"
  )

load(here::here("output/ocde-egalitarian-homclass-full-macro.RData"));ocde_num_full_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)
load(here::here("output/ocde-egalitarian-homclass-strong-macro.RData"));ocde_num_strong_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)
load(here::here("output/ocde-egalitarian-homclass-weak-macro.RData")); ocde_num_weak_mac<- list(giniD_int,giniM_int,int_palma,int_pc10,int_pc50,int_d10d1,int_gini_wiid,int_theil)

htmlreg(
  c(ocde_num_full_mac,ocde_num_strong_mac,ocde_num_weak_mac),
  caption = "Multilevel linear regression: full country sample cross-level interaction - egalitarian index",
  caption.above = T,
  custom.coef.map= coef_map,
  custom.header = 
    list(
      "Full network"=1:8,
      "Strong ties"=9:16,
      "Weak ties"=17:24
      ))
```