### 2023-11-02 procesamiento/prod_analysis-plots-models.Rmd
ajuste

---
title: "Estimations plots"
subtitle: "Full sample"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output: 
  bookdown::html_document2: 
    toc: yes
    code_folding: hide
    toc_float: 
      collapsed: true
      smooth_scroll: false
      number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(scipen=9999) # desactivar notacion cientifica
```

```{r}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(theme_bw())
ggplot2::theme_update(text=element_text(size=15,  family="serif"))  
```

# Preferences for Goverment redistribution

```{r full-redist-est}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg,broom.mixed,ggplot2,haven)  
rm(list=ls())

load(here::here("output/fullsample-numeric-homclass-full-individual.RData")); fsample_num_full_ind<- models
load(here::here("output/fullsample-numeric-homclass-strong-individual.RData"));fsample_num_strong_ind<- models
load(here::here("output/fullsample-numeric-homclass-weak-individual.RData")); fsample_num_weak_ind<- models

coef <- c("homclass","class3Intermediate class (III+IV)","class3Working Class (V+VI+VII)")

df_p_full<- 
bind_rows(
  fsample_num_full_ind[[3]] %>% 
    broom.mixed::tidy(conf.int = T) %>% 
    filter(term %in% coef) %>% 
    mutate(group ="Full network"),
  fsample_num_strong_ind[[3]] %>% 
    broom.mixed::tidy(conf.int = T) %>% 
    filter(term %in% coef) %>% 
    mutate(group ="Strong ties"),
  fsample_num_weak_ind[[3]] %>%
    broom.mixed::tidy(conf.int = T) %>%
    filter(term %in% coef) %>%
    mutate(group ="Weak ties")
)
df_p_full$depvar <- "Full sample"
```

```{r nonconfusian-redist-est}
load(here::here("output/nonconfusian-numeric-homclass-full-individual.RData")); nconf_num_full_ind<- models
load(here::here("output/nonconfusian-numeric-homclass-strong-individual.RData"));nconf_num_strong_ind<- models
load(here::here("output/nonconfusian-numeric-homclass-weak-individual.RData")); nconf_num_weak_ind<- models

coef <- c("homclass","class3Intermediate class (III+IV)","class3Working Class (V+VI+VII)")

df_p_nconf<- 
bind_rows(
  nconf_num_full_ind[[3]] %>% 
    broom.mixed::tidy(conf.int = T) %>% 
    filter(term %in% coef) %>% 
    mutate(group ="Full network"),
  nconf_num_strong_ind[[3]] %>% 
    broom.mixed::tidy(conf.int = T) %>% 
    filter(term %in% coef) %>% 
    mutate(group ="Strong ties"),
  nconf_num_weak_ind[[3]] %>%
    broom.mixed::tidy(conf.int = T) %>%
    filter(term %in% coef) %>%
    mutate(group ="Weak ties")
)
df_p_nconf$depvar <- "Non-Confusian"
```

```{r oecd-redist-est}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
load(here::here("output/ocde-numeric-homclass-full-individual.RData")); ocde_num_full_ind<- models
load(here::here("output/ocde-numeric-homclass-strong-individual.RData"));ocde_num_strong_ind<- models
load(here::here("output/ocde-numeric-homclass-weak-individual.RData")); ocde_num_weak_ind<- models

df_p_ocde<- 
bind_rows(
  ocde_num_full_ind[[3]] %>% 
    broom.mixed::tidy(conf.int = T) %>% 
    filter(term %in% coef) %>% 
    mutate(group ="Full network"),
  ocde_num_strong_ind[[3]] %>% 
    broom.mixed::tidy(conf.int = T) %>% 
    filter(term %in% coef) %>% 
    mutate(group ="Strong ties"),
  ocde_num_weak_ind[[3]] %>%
    broom.mixed::tidy(conf.int = T) %>%
    filter(term %in% coef) %>%
    mutate(group ="Weak ties")
)
df_p_ocde$depvar <- "OECD"
```

```{r coefplot-redist}
df_p<- bind_rows(df_p_full,df_p_nconf,df_p_ocde)
df_p$term <- 
  factor(df_p$term,
         levels = c("class3Working Class (V+VI+VII)",
                    "class3Intermediate class (III+IV)",
                    "homclass"))
df_p$group <- factor(df_p$group,levels = c("Weak ties","Strong ties","Full network"))

ggplot(df_p,
       aes(x = estimate, y = term, color = group)) +
  geom_vline(xintercept = 0) +
  geom_errorbar(
    aes(xmin = conf.low, xmax = conf.high),
    size = 1,
    width = 0.2,
    position = position_dodge(0.5)
  ) +
  geom_point(
    shape = 21,
    size = 3,
    fill = "white",
    position = position_dodge(0.5)
  ) +
  facet_grid(~depvar)+
  scale_y_discrete(
    labels = c(
      "Working Class (VI+VII)",
      "Intermediate class (III+IV+V)",
      "Class-based homogeneity"
    )
  )+
  scale_color_discrete(guide = guide_legend(reverse = TRUE))+
  labs(
    # caption = paste0("N=", nobs(int_homo), ", Countries=", lme4::ngrps(int_homo)),
    color = "Model",
    x="Coefficients (fixed effects)",y=NULL,
    title = "Preferences for Goverment redistribution") +
  theme(axis.text.y.left = element_text(face = "plain", size = 12))
ggsave(plot = last_plot(),filename = "coefplot-redist.png",device = "png",
       path = here::here("output/images"),width = 2,height = 1,units = "cm",scale = 15)
```

# Egalitarian economic preferences

```{r full-redist-est}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg,broom.mixed,ggplot2,haven)  
rm(list=ls())

load(here::here("output/fullsample-egalitarian-homclass-full-individual.RData")); fsample_egal_full_ind<- models
load(here::here("output/fullsample-egalitarian-homclass-strong-individual.RData"));fsample_egal_strong_ind<- models
load(here::here("output/fullsample-egalitarian-homclass-weak-individual.RData")); fsample_egal_weak_ind<- models

coef <- c("homclass","class3Intermediate class (III+IV)","class3Working Class (V+VI+VII)")

df_p_full<- 
bind_rows(
  fsample_egal_full_ind[[3]] %>% 
    broom.mixed::tidy(conf.int = T) %>% 
    filter(term %in% coef) %>% 
    mutate(group ="Full network"),
  fsample_egal_strong_ind[[3]] %>% 
    broom.mixed::tidy(conf.int = T) %>% 
    filter(term %in% coef) %>% 
    mutate(group ="Strong ties"),
  fsample_egal_weak_ind[[3]] %>%
    broom.mixed::tidy(conf.int = T) %>%
    filter(term %in% coef) %>%
    mutate(group ="Weak ties")
)
df_p_full$depvar <- "Full sample"
```

```{r nonconfusian-redist-est}
load(here::here("output/nonconfusian-egalitarian-homclass-full-individual.RData")); nconf_egal_full_ind<- models
load(here::here("output/nonconfusian-egalitarian-homclass-strong-individual.RData"));nconf_egal_strong_ind<- models
load(here::here("output/nonconfusian-egalitarian-homclass-weak-individual.RData")); nconf_egal_weak_ind<- models

coef <- c("homclass","class3Intermediate class (III+IV)","class3Working Class (V+VI+VII)")

df_p_nconf<- 
bind_rows(
  nconf_egal_full_ind[[3]] %>% 
    broom.mixed::tidy(conf.int = T) %>% 
    filter(term %in% coef) %>% 
    mutate(group ="Full network"),
  nconf_egal_strong_ind[[3]] %>% 
    broom.mixed::tidy(conf.int = T) %>% 
    filter(term %in% coef) %>% 
    mutate(group ="Strong ties"),
  nconf_egal_weak_ind[[3]] %>%
    broom.mixed::tidy(conf.int = T) %>%
    filter(term %in% coef) %>%
    mutate(group ="Weak ties")
)
df_p_nconf$depvar <- "Non-Confusian"
```

```{r oecd-redist-est}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
load(here::here("output/ocde-egalitarian-homclass-full-individual.RData")); ocde_egal_full_ind<- models
load(here::here("output/ocde-egalitarian-homclass-strong-individual.RData"));ocde_egal_strong_ind<- models
load(here::here("output/ocde-egalitarian-homclass-weak-individual.RData")); ocde_egal_weak_ind<- models

df_p_ocde<- 
bind_rows(
  ocde_egal_full_ind[[3]] %>% 
    broom.mixed::tidy(conf.int = T) %>% 
    filter(term %in% coef) %>% 
    mutate(group ="Full network"),
  ocde_egal_strong_ind[[3]] %>% 
    broom.mixed::tidy(conf.int = T) %>% 
    filter(term %in% coef) %>% 
    mutate(group ="Strong ties"),
  ocde_egal_weak_ind[[3]] %>%
    broom.mixed::tidy(conf.int = T) %>%
    filter(term %in% coef) %>%
    mutate(group ="Weak ties")
)
df_p_ocde$depvar <- "OECD"
```

```{r coefplot-egal}
df_p<- bind_rows(df_p_full,df_p_nconf,df_p_ocde)
df_p$term <- 
  factor(df_p$term,
         levels = c("class3Working Class (V+VI+VII)",
                    "class3Intermediate class (III+IV)",
                    "homclass"))
df_p$group <- factor(df_p$group,levels = c("Weak ties","Strong ties","Full network"))

ggplot(df_p,
       aes(x = estimate, y = term, color = group)) +
  geom_vline(xintercept = 0) +
  geom_errorbar(
    aes(xmin = conf.low, xmax = conf.high),
    size = 1,
    width = 0.2,
    position = position_dodge(0.5)
  ) +
  geom_point(
    shape = 21,
    size = 3,
    fill = "white",
    position = position_dodge(0.5)
  ) +
  facet_grid(~depvar)+
  scale_y_discrete(
    labels = c(
      "Working Class (VI+VII)",
      "Intermediate class (III+IV+V)",
      "Class-based homogeneity"
    )
  )+
  scale_color_discrete(guide = guide_legend(reverse = TRUE))+
  labs(
    # caption = paste0("N=", nobs(int_homo), ", Countries=", lme4::ngrps(int_homo)),
    color = "Model",
    x="Coefficients (fixed effects)",y=NULL,
    title = "Egalitarian Economic Preferences (Index)") +
  theme(axis.text.y.left = element_text(face = "plain", size = 12))

ggsave(plot = last_plot(),filename = "coefplot-egal.png",device = "png",
       path = here::here("output/images"),width = 2,height = 1,units = "cm",scale = 15)
```