### 2023-11-28 procesamiento/prod_analysis-tables-micro.Rmd

---
title: "Additional analyses"
subtitle: "Micro level"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
css: "../input/css/custom.css" # custom CSS para html
output: 
  bookdown::html_document2: 
    toc: false
    # toc_float: 
    #   collapsed: true
    #   smooth_scroll: false
    #   number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r echo=FALSE}
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
  echo=FALSE, results='asis'
)
options(scipen=9999) # desactivar notacion cientifica
```

# All countries

```{r full-lmer-micro, echo=FALSE, results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())

load(here::here("output/fullsample-numeric-homclass-full-individual.RData")); fsample_num_full_ind<- models
load(here::here("output/fullsample-numeric-homclass-strong-individual.RData"));fsample_num_strong_ind<- models
load(here::here("output/fullsample-numeric-homclass-weak-individual.RData")); fsample_num_weak_ind<- models

htmlreg(
  c(fsample_num_full_ind,fsample_num_strong_ind,fsample_num_weak_ind),
  caption = "Multilevel linear regression: full country sample",
  caption.above = T,
  custom.header = 
    list(
      "Full network"=1:4,
      "Strong ties"=5:8,
      "Weak ties"=9:12
      ))
```

```{r full-clmm-micro,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg,ordinal)  
rm(list=ls())

load(here::here("output/fullsample-ologit-homclass-full-individual.RData"));fsample_ologit_full_ind<- models
load(here::here("output/fullsample-ologit-homclass-strong-individual.RData"));fsample_ologit_strong_ind<- models
load(here::here("output/fullsample-ologit-homclass-weak-individual.RData"));fsample_ologit_weak_ind<- models

htmlreg(
  c(fsample_ologit_full_ind,
    fsample_ologit_strong_ind,
    fsample_ologit_weak_ind),
  include.threshold=F,
  caption = "Multilevel ordinal logistic regression: full country sample",
  caption.above = T,
  custom.header = 
    list(
      "Full network"=1:4,
      "Strong ties"=5:8,
      "Weak ties"=9:12
      ))
```

```{r full-lmer-micro-egalitarian,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())

load(here::here("output/fullsample-egalitarian-homclass-full-individual.RData")); fsample_num_full_ind<- models
load(here::here("output/fullsample-egalitarian-homclass-strong-individual.RData"));fsample_num_strong_ind<- models
load(here::here("output/fullsample-egalitarian-homclass-weak-individual.RData")); fsample_num_weak_ind<- models

coefmap <-  list(homclass ="Class-based network homogeneity",
                 'class3Intermediate class (III+IV)' ='Intermediate Class',
                 'class3Working Class (V+VI+VII)' ='Working Class',
                 "Q03pcmT02" = "Middle",
                 "Q03pcmT03" = "High",
                 "edyears" = 'Education in years',
                 "workstNot in paid work"="Not in labor force",
                 "unionYes" = 'Union membership: Yes',
                 'homclass:class3Intermediate class (III+IV)'='Homogeneity x Intermediate Class',
                 'homclass:class3Working Class (V+VI+VII)'='Homogeneity x Working Class'
                 )

htmlreg(
  c(fsample_num_full_ind,fsample_num_strong_ind,fsample_num_weak_ind),
  caption = "Multilevel linear regression: full country sample - egalitarian preferences index",
  caption.above = T,
  custom.header = 
    list(
      "Full network"=1:4,
      "Strong ties"=5:8,
      "Weak ties"=9:12
      ), 
  custom.coef.map = coefmap
  )
```

# Non-confusian

```{r nonconfusian-lmer-micro,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())

load(here::here("output/nonconfusian-numeric-homclass-full-individual.RData")); nconf_num_full_ind<- models
load(here::here("output/nonconfusian-numeric-homclass-strong-individual.RData"));nconf_num_strong_ind<- models
load(here::here("output/nonconfusian-numeric-homclass-weak-individual.RData")); nconf_num_weak_ind<- models

htmlreg(
  c(nconf_num_full_ind,nconf_num_strong_ind,nconf_num_weak_ind),
  caption = "Multilevel linear regression: Non-confusian country sample",
  caption.above = T,
  custom.header = 
    list(
      "Full network"=1:4,
      "Strong ties"=5:8,
      "Weak ties"=9:12
      ))
```

```{r nonconfusian-clmm-micro,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg,ordinal)  
rm(list=ls())

load(here::here("output/nonconfusian-ologit-homclass-full-individual.RData"));nconf_ologit_full_ind<- models
load(here::here("output/nonconfusian-ologit-homclass-strong-individual.RData"));nconf_ologit_strong_ind<- models
load(here::here("output/nonconfusian-ologit-homclass-weak-individual.RData"));nconf_ologit_weak_ind<- models

htmlreg(
  c(nconf_ologit_full_ind,
    nconf_ologit_strong_ind,
    nconf_ologit_weak_ind),
  include.threshold=F,
  caption = "Multilevel ordinal logistic regression: Non-confusian country sample",
  caption.above = T,
  custom.header = 
    list(
      "Full network"=1:4,
      "Strong ties"=5:8,
      "Weak ties"=9:12
      ))
```

```{r nonconfusian-lmer-micro-egalitarian,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())

load(here::here("output/nonconfusian-egalitarian-homclass-full-individual.RData")); fsample_num_full_ind<- models
load(here::here("output/nonconfusian-egalitarian-homclass-strong-individual.RData"));fsample_num_strong_ind<- models
load(here::here("output/nonconfusian-egalitarian-homclass-weak-individual.RData")); fsample_num_weak_ind<- models

knitreg(
  c(fsample_num_full_ind,fsample_num_strong_ind,fsample_num_weak_ind),
  caption = "Multilevel linear regression: Non-confusian sample - egalitarian preferences index",
  caption.above = T,
  custom.header = 
    list(
      "Full network"=1:4,
      "Strong ties"=5:8,
      "Weak ties"=9:12
      ))
```

# OECD

```{r ocde-lmer-micro,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())

load(here::here("output/ocde-numeric-homclass-full-individual.RData")); ocde_num_full_ind<- models
load(here::here("output/ocde-numeric-homclass-strong-individual.RData"));ocde_num_strong_ind<- models
load(here::here("output/ocde-numeric-homclass-weak-individual.RData")); ocde_num_weak_ind<- models

htmlreg(
  c(ocde_num_full_ind,ocde_num_strong_ind,ocde_num_weak_ind),
  caption = "Multilevel linear regression: OECD country sample",
  caption.above = T,
  custom.header = 
    list(
      "Full network"=1:4,
      "Strong ties"=5:8,
      "Weak ties"=9:12
      ))
```

```{r oecd-clmm-micro,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())

load(here::here("output/ocde-ologit-homclass-full-individual.RData")); ocde_ologit_full_ind<- models
load(here::here("output/ocde-ologit-homclass-strong-individual.RData"));ocde_ologit_strong_ind<- models
load(here::here("output/ocde-ologit-homclass-weak-individual.RData")); ocde_ologit_weak_ind<- models

htmlreg(
  c(ocde_ologit_full_ind,ocde_ologit_strong_ind,ocde_ologit_weak_ind),
  caption = "Multilevel ordinal logistic regression: OECD country sample",
  caption.above = T,
  include.threshold=F,
  custom.header = 
    list(
      "Full network"=1:4,
      "Strong ties"=5:8,
      "Weak ties"=9:12
      ))
```

```{r oecd-lmer-micro-egalitarian,results='asis'}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,texreg)  
rm(list=ls())

load(here::here("output/ocde-egalitarian-homclass-full-individual.RData")); fsample_num_full_ind<- models
load(here::here("output/ocde-egalitarian-homclass-strong-individual.RData"));fsample_num_strong_ind<- models
load(here::here("output/ocde-egalitarian-homclass-weak-individual.RData")); fsample_num_weak_ind<- models

knitreg(
  c(fsample_num_full_ind,fsample_num_strong_ind,fsample_num_weak_ind),
  caption = "Multilevel linear regression: OECD country sample - egalitarian preferences index",
  caption.above = T,
  custom.header = 
    list(
      "Full network"=1:4,
      "Strong ties"=5:8,
      "Weak ties"=9:12
      ))
```