### 2024-02-23 procesamiento/02-prod_prep-macro.Rmd
update de ginis

---
title: "Data preparation"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output: 
  html_document: 
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
      number_sections: yes
    code_folding: show  
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

# Setup

```{r setup}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE 
                      )
options(scipen=9999) # desactivar notacion cientifica
if (!require("pacman")) install.packages("pacman") # instalar pacman
load(here::here("input/data/proc/study1.RData"))
pacman::p_load(dplyr,sjmisc,sjPlot,sjlabelled,tidyverse,haven,MLMusingR)
```

# Country level

```{r}
countries <- 
df1 %>% 
  group_by(country) %>% 
  summarise(n=n())
countries$ccode <- as.numeric(countries$country)
countries$countryname <- as_character(countries$country)
```

Revisar:

https://wid.world/codes-dictionary#post-tax-income

```{r}
wid_gini <- haven::read_dta(here::here("input/data/original/wid_gini.dta"))
wid_ratios <- haven::read_dta(here::here("input/data/original/wid_ratios.dta"))
wid_shares <- haven::read_dta(here::here("input/data/original/wid_shares.dta"))

# wid_gini_2018 <- wid_gini %>% filter(year==2018)
# wid_gini_2018$year[wid_gini_2018$year==2018] <- 2016
# wid_gini_2018$country_yr <- stringi::stri_replace_all(str = wid_gini_2018$country_yr,replacement = "2016",regex = "2018")

wid_df <-wid_ratios  %>% 
  dplyr::left_join(wid_shares,by="country_yr") %>% 
  dplyr::left_join(wid_gini,by="country_yr")

names(wid_df)
# summary(wid_df)
wid_df <- 
wid_df %>% 
  select(country=country.x,year=year.x,country_yr,starts_with("wid_"))

wid_df$ccode <- countrycode::countrycode(sourcevar = wid_df$country,origin = "iso2c",destination = "iso3n")
wid_df <- wid_df %>% filter(year==2018)%>% select(ccode,wid_sharetop10,wid_rp90p50,wid_rd10d01,wid_gini_disp) 
```

**World Income Inequality Database (WIID)**

> UNU-WIDER, World Income Inequality Database (WIID) Companion dataset (wiidcountry and/or wiidglobal). Version 30 June 2022. https://doi.org/10.35188/UNU-WIDER/WIIDcomp-300622

```{r , eval=FALSE, include=FALSE}
wiid_long<- haven::read_dta(here::here("input/data/original/WIID_30JUN2022_0/WIID_30JUN2022_0.dta"))
wiid_long$ccode <- countrycode::countrycode(sourcevar = wiid_long$c3,origin = "iso3c",destination = "iso3n")

wiid_long1 <- 
wiid_long %>% 
  filter(ccode %in% countries$ccode) %>% 
  filter(year==2017)
```

```{r}
wiid <- 
  read.csv(file = here::here("input/data/original/wid-percentiles/wiid-data-2015-2020.csv"),sep = ";")
names(wiid) <- tolower(names(wiid))
names(wiid) <- gsub('\\.', '', names(wiid))
wiid$ccode <- countrycode::countrycode(sourcevar = wiid$iso,origin = "iso3c",destination = "iso3n")

wiid_d10d01 <- 
  read.csv(file = here::here("input/data/original/wid-percentiles/wiid-data-D10D01-ratio.csv"),sep = ";")

names(wiid_d10d01) <- tolower(names(wiid_d10d01))
names(wiid_d10d01) <- gsub('\\.', '', names(wiid_d10d01))
wiid_d10d01$ccode <- countrycode::countrycode(sourcevar = wiid_d10d01$iso,origin = "iso3c",destination = "iso3n")
wiid_d10d01 <- 
  wiid_d10d01 %>%  
  filter(ccode %in% countries$ccode) %>% 
  filter(year==2017) %>% 
  select(ccode,everything(),-iso,-country,-year,-giniindex)

wiid_s80s10 <- 
  read.csv(file = here::here("input/data/original/wid-percentiles/wiid-data-P80P20.csv"),sep = ";")
names(wiid_s80s10) <- tolower(names(wiid_s80s10))
names(wiid_s80s10) <- gsub('\\.', '', names(wiid_s80s10))
wiid_s80s10$ccode <- countrycode::countrycode(sourcevar = wiid_s80s10$iso,origin = "iso3c",destination = "iso3n")

wiid_s80s10 <- 
  wiid_s80s10 %>%  
  filter(ccode %in% countries$ccode) %>% 
  filter(year==2017) %>% 
  select(ccode,everything(),-iso,-country,-year,-giniindex)

wiid_2017 <- 
wiid %>% 
  filter(ccode %in% countries$ccode) %>% 
  filter(year==2017) %>% 
  select(ccode,everything(),-iso,-country,-year) %>% 
  left_join(wiid_d10d01,by="ccode") %>% 
  left_join(wiid_s80s10,by="ccode")
```

**Solt SWDIID**

```{r Solt}
load(file = here::here("input/data/original/swiid9_5/swiid9_5.rda"))
solt <- 
swiid_summary %>% dplyr::select(country,year,gini_disp,gini_mkt,abs_red,rel_red) %>% dplyr::filter(year %in% c(2016))
summary(solt)
solt$ccode <- countrycode::countrycode(sourcevar = solt$country,origin = "country.name",destination = "iso3n")
solt$country <- NULL
solt_2017 <-  solt %>% filter(ccode %in% countries$ccode) %>% 
  dplyr::select(ccode,gini_disp,gini_mkt)

solt_2017$abs_red <- solt_2017$gini_mkt - solt_2017$gini_disp 
solt_2017$rel_red <- solt_2017$abs_red/solt_2017$gini_mkt
summary(solt_2017)
```

**Otero et.al (2023)**

```{r otero23}
otero <- 
  read.csv(file = here::here("input/data/original/otero23.csv"),sep = ";")
otero$ccode <- countrycode::countrycode(sourcevar = otero$iso3n,origin = "iso2c",destination = "iso3n")
otero$gdp <- gsub('\\,', '',otero$gdp) 
otero<- otero %>% select(ccode,gv_spen,individualism)
```

**GDP: PEN data**
```{r pen-data}
pendata<- haven::read_dta(file = here::here("input/data/original/pen-world/pwt1001.dta"))
pendata$ccode <- countrycode::countrycode(sourcevar = pendata$countrycode,origin = "iso3c",destination = "iso3n")
pen_2017 <- pendata %>% filter(ccode %in% countries$ccode,year %in% 2017) %>% select(ccode,everything(),-country,-countrycode,-currency_unit,-year)
```

*OCDE data*
```{r ocde}
ocde<- 
  read.csv(file = here::here("input/data/original/ocde-countries.csv"),sep = ",")

ocde$ccode <- countrycode::countrycode(sourcevar = ocde$iso3n,origin = "iso3c",destination = "iso3n")
ocde <- ocde[,c("ccode","oecd")]
```

```{r merge}
macro_data <- 
  left_join(wiid_2017,solt_2017) %>%
  left_join(pen_2017) %>% 
  left_join(otero) %>%
  left_join(ocde) %>% 
  left_join(wid_df)

df1$ccode <- as.numeric(df1$country)
df2 <- left_join(df1,macro_data,by="ccode")
df2 <- 
df2 %>% mutate(country2=countrycode::countrycode(sourcevar = ccode,origin = "iso3n",destination = "iso3c"),
               country3=countrycode::countrycode(sourcevar = ccode,origin = "iso3n",destination = "country.name"),
               region=countrycode::countrycode(sourcevar = ccode,origin = "iso3n",destination = "region23")) %>% 
  select(ccode,country2,country3,everything())
```

```{r}
df2 %>% 
  select(country3,wid_rd10d01,d10d1) %>%
  group_by(country3) %>% 
  summarise(wid_rd10d01=mean(wid_rd10d01), d10d1=mean(d10d1)) %>% View
```

# centering micro level

```{r eval=FALSE, include=FALSE}
df2 <- 
df2 %>% 
  mutate(to_dummy(female),
         to_dummy(union),
         to_dummy(workst),
         # dummy for categorical variables-------------------------------------
         to_dummy(class3),to_dummy(class3spo),to_dummy(class3res),to_dummy(Q03pcm))
df2$female_gc = group_center(df2$female_2, grp = df2$country2)
df2$union_gc = group_center(df2$union_2, grp = df2$country2)
df2$workst_gc = group_center(df2$workst_2, grp = df2$country2)
df2$edyears_gc = group_center(df2$edyears, grp = df2$country2)
df2$agenum_gc = group_center(df2$agenum, grp = df2$country2)

# EGP: dominance
df2$class3_ser_gc = group_center(df2$class3_1, grp = df2$country2)
df2$class3_mid_gc = group_center(df2$class3_2, grp = df2$country2)
df2$class3_wor_gc = group_center(df2$class3_3, grp = df2$country2)
# EGP: Respondent
df2$class3res_ser_gc = group_center(df2$class3res_1, grp = df2$country2)
df2$class3res_mid_gc = group_center(df2$class3res_2, grp = df2$country2)
df2$class3res_wor_gc = group_center(df2$class3res_3, grp = df2$country2)
# EGP: Respondent
df2$class3spo_ser_gc = group_center(df2$class3spo_1, grp = df2$country2)
df2$class3spo_mid_gc = group_center(df2$class3spo_2, grp = df2$country2)
df2$class3spo_wor_gc = group_center(df2$class3spo_3, grp = df2$country2)
# Household Income Tercile
df2$Q03pcm_1_gc = group_center(df2$Q03pcm_1, grp = df2$country2)
df2$Q03pcm_2_gc = group_center(df2$Q03pcm_2, grp = df2$country2)
df2$Q03pcm_3_gc = group_center(df2$Q03pcm_3, grp = df2$country2)
df2$Q03pcm_NA_gc = group_center(df2$Q03pcm_4, grp = df2$country2)

names(df2)
# drop dummies
df2 <- 
  df2 %>% 
  select(-c(female_1,female_2,union_1,union_2,workst_1,workst_2,
            class3_1,class3_2,class3_3,class3res_1,class3res_2,class3res_3,
            class3spo_1,class3spo_2,class3spo_3,
            Q03pcm_1,Q03pcm_2,Q03pcm_2,Q03pcm_4))
names(df2)
```

# centering macro level

```{r eval=FALSE, include=FALSE}
df2$gini_disp_c <- center(df2$gini_disp)
df2$gini_mkt_c <- center(df2$gini_mkt)
df2$palmaratio_c <- center(df2$palmaratio)
df2$top10_c <- center(df2$top10)
df2$middle50_c <- center(df2$middle50)
df2$d10d1_c <- center(df2$d10d1)
df2$giniindex_c <- center(df2$giniindex)
df2$ttheilge1_c <- center(df2$ttheilge1)
df2$rgdpna_c <- center(df2$rgdpna)

df2$gini_disp_z <- center(df2$gini_disp)
df2$gini_mkt_z <- center(df2$gini_mkt)
df2$palmaratio_z <- center(df2$palmaratio)
df2$top10_z <- center(df2$top10)
df2$middle50_z <- center(df2$middle50)
df2$d10d1_z <- center(df2$d10d1)
df2$giniindex_z <- center(df2$giniindex)
df2$ttheilge1_z <- center(df2$ttheilge1)
df2$rgdpna_z <- center(log(df2$rgdpna))

df2 %>% 
  select(country3,ends_with("_c")) %>% group_by(country3) %>% summarise_all(.funs = mean) %>% View()
table(df2$country2)
```

```{r save}
save(df2,file = here::here("input/data/proc/study1_country.RData"))
# haven::write_dta(data = df2,path = here::here("input/data/proc/study1_country.dta"))
sjPlot::view_df(df2,file =here::here("input/data/proc/codebook-s1.html") )
```