### 2024-07-09 paper1_jiturra_supp-material.Rmd

---
title: "Class-based network segregation, Economic Inequality and Redistributive Preferences across societies"
subtitle: "Supplementary material"
css: "input/css/custom.css" # custom CSS para html
fontsize: 12pt
linestretch: '1.15'          # interlineado 
papersize: a4
link-citations: yes         # citas linkeadas
lang: en-US
# date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
date: "`r format(Sys.Date(), '%d %B %Y')`"
author:
- name: Julio Iturra-Sanhueza
  affiliation: Bremen International Graduate School of Social Sciences
  email: jiturra@uni-bremen.de
#   number: 1
# - name: Justin d'Ottawa
#   affiliation: University of Ottawa
#   number: 2
# - name: Pedro Torres
#   number: 1,2 
# Nota: Autores que comparten filiacion, poner el mismo number y declarar filiación una sola vez.  
# abstract: | 
#   # | **Keywords**: network segregation, redistributive preferences, economic inequality
output:
  bookdown::pdf_document2:
    template: null
    toc: false
    keep_tex: false
    number_sections: true
    pandoc_args:
      - --template=input/mytemplate.tex #custom template para usar autores con afiliacion  
  bookdown::html_document2:
    toc: yes
    code_folding: hide
    number_sections: false
    toc_float:
      collapsed: false
      smooth_scroll: false

always_allow_html: true      
linkcolor: gray                         # enlaces y citas en color azul
bibliography: input/bib/rgroup-redist.bib     # bibliografia en bibtex
csl: input/bib/apa6.csl
editor_options:
  chunk_output_type: console            # en RStudio, mostrar output en consola
geometry: "left=2.54cm,right=2.54cm,top=2.54cm,bottom=2.54cm" # márgenes de página
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \setlength\parindent{24pt} 
  - \usepackage{fancyhdr}
  - \fancyhead{} 
---

```{r setup, include=FALSE}
if (!require("pacman")) install.packages("pacman")  #si falta pacman, instalar
if (!require("tinytex")) install.packages("tinytex")#si falta tinytex, instalar
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)
knitr::opts_chunk$set(warning = FALSE,  # mensaje de warning
                      message = FALSE,  # mensajes/avisos de librerias  
                      cache = F,    # cache de los chunks,usar analisis pesados
                      out.width = '100%',
                      fig.pos= "H",     # posicion figuras H = HERE
                      fig.align = "center",
                      echo = FALSE      # incluir chunk en output
                      )
options(scipen=999) # notacion cientifica
rm(list=ls())       # limpiar workspace
options(knitr.kable.NA = '') # NA en kable = ''
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
```
\pagestyle{fancy}
\fancyhead[R]{\thepage}
\fancyhead[L]{Supplementary material}

\pagebreak

# Supplementary material

## Income inequality ratios

```{r tab2-ratio9010, results='asis', cache=TRUE}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2

df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)"))

dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  # starts_with("class3"),
  # starts_with("digclass3"),
  # starts_with("dclass3"),
  # starts_with("homclass"),
  dclass3res_V,
  # dclass3spo_V,
  homclass3_V_res,
  know_total,
  # socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  # religion,
  partner,
  # union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         gini_disp=scale(d10d1))

dfreg$wstate2 <- scale(scale(dfreg$abs_red*100) + scale(dfreg$ilo_taxrev) + scale(dfreg$ilo_govspe))

# Variable Group-Centering ------------------------------------------------$
dfreg <- 
  dfreg %>% 
  mutate(to_dummy(female),
         # to_dummy(union),
         to_dummy(workst),
         # dummy for categorical variables-------------------------------------$
         to_dummy(Q03pcm),
         # to_dummy(dclass3spo_V)
  )
dfreg$female_gc = group_center(dfreg$female_2, grp = dfreg$country2)
# dfreg$union_gc = group_center(dfreg$union_2, grp = dfreg$country2)
dfreg$workst_gc = group_center(dfreg$workst_2, grp = dfreg$country2)
dfreg$edyears_gc = group_center(dfreg$edyears, grp = dfreg$country2)
dfreg$agenum_gc = group_center(dfreg$agenum, grp = dfreg$country2)
dfreg$age2_gc = group_center(dfreg$age2, grp = dfreg$country2)
# dfreg$socialtrust_gc = group_center(dfreg$socialtrust, grp = dfreg$country2)
dfreg$homclass3_V_gc <-group_center(dfreg$homclass3_V_res, grp = dfreg$country2)
dfreg$know_total_gc = group_center(dfreg$know_total, grp = dfreg$country2)
# dfreg$religion_gc = group_center(as.numeric(dfreg$religion), grp = dfreg$country2)
dfreg$partner_gc = group_center(as.numeric(dfreg$partner), grp = dfreg$country2)

# Household Income Tercile
dfreg$Q03pcm_1_gc = group_center(dfreg$Q03pcm_1, grp = dfreg$country2)
dfreg$Q03pcm_2_gc = group_center(dfreg$Q03pcm_2, grp = dfreg$country2)
dfreg$Q03pcm_3_gc = group_center(dfreg$Q03pcm_3, grp = dfreg$country2)
dfreg$Q03pcm_NA_gc = group_center(dfreg$Q03pcm_4, grp = dfreg$country2)

# Partner Social Class
# dfreg$dclass3spo_V_1_gc = group_center(dfreg$dclass3spo_V_1, grp = dfreg$country2)
# dfreg$dclass3spo_V_2_gc = group_center(dfreg$dclass3spo_V_2, grp = dfreg$country2)
# dfreg$dclass3spo_V_3_gc = group_center(dfreg$dclass3spo_V_3, grp = dfreg$country2)
# dfreg$dclass3spo_V_NA_gc = group_center(dfreg$dclass3spo_V_4, grp = dfreg$country2)

# drop dummies
dfreg <- dplyr::select(dfreg,-c(female_1,female_2,
                                # union_1,union_2,
                                workst_1,
                                workst_2,Q03pcm_1,Q03pcm_2,Q03pcm_2,Q03pcm_4,
                                # dclass3spo_V_1,dclass3spo_V_2,dclass3spo_V_3,dclass3spo_V_4
                                ))

# dfreg$gini_disp <- scale(dfreg$gini_disp)

## Model -  EGP3 - DIGCLASS - V to Working -----------------------------$
homo_V_gini <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+
         gini_disp +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+
         gini_disp +loggdppercapita +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp_wstate2 <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2+
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp_controls <-
  lmer(egal~
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2 +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_full2 <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+         
         gini_disp +loggdppercapita + wstate2+
         homclass3_V_gc*
         dclass3res_V*
         gini_disp + 
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

digclas1_macro <- list(homo_V_gini,
                       homo_V_gini_gdp,
                       homo_V_gini_gdp_wstate2,
                       homo_V_gini_full2)

ccoef <- list(homclass3_V_gc="Class-based network homogeneity (CWC)",
              "dclass3res_VIntermediate class (III+IV)"="Intermediate Class",
              "dclass3res_VWorking Class (V+VI+VII)"="Working Class",
              "gini_disp" = "Income inequality (Ratio 90/10)",
              loggdppercapita = "GDP/capita",
              wstate2="Size of the welfare state",
              "homclass3_V_gc:dclass3res_VIntermediate class (III+IV)" = "Homogeneity*Intermediate Class",
              "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII)" = "Homogeneity*Working Class",
              "homclass3_V_gc:gini_disp"= "Homogeneity*Income Inequality",
              "homclass3_V_gc:dclass3res_VIntermediate class (III+IV):gini_disp" = "Homogeneity*Intermediate Class*Income Inequality",
              "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII):gini_disp" = "Homogeneity*Working Class *Income Inequality")

texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.50,
       caption = "Multilevel models for Income Inequality, Network segregation and Redistributive Preferences",
       custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars",
       caption.above = T,leading.zero = T,single.row = T,
       custom.coef.map = ccoef,
      groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Homogeneity*Social Class"=7:8, "Homogeneity * Social Class * Income Inequality"=10:11),
       float.pos = "h!",
      custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups",
                           "Var: Group (Intercept)",
                           "Var: Group Homogeneity",
                           "Var: Group Intermediate Class ",
                           "Var: Group Working Class",
                           "Cov: Group (Intercept), Homogeneity",
                           "Cov: Group (Intercept), Intermediate Class ",
                           "Cov: Group (Intercept), Working Class",
                           "Cov: Group Homogeneity, Intermediate Class",
                           "Cov: Group Homogeneity, Working Class",
                           "Cov: Group Intermediate Class, Working Class",
                           "Var: Residual"),
       use.packages = F)
```

```{r tab2-ratio9050, results='asis', cache=TRUE}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2

df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)"))

dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  # starts_with("class3"),
  # starts_with("digclass3"),
  # starts_with("dclass3"),
  # starts_with("homclass"),
  dclass3res_V,
  # dclass3spo_V,
  homclass3_V_res,
  know_total,
  # socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  # religion,
  partner,
  # union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         gini_disp=scale(wid_p90p50))

dfreg$wstate2 <- scale(scale(dfreg$abs_red*100) + scale(dfreg$ilo_taxrev) + scale(dfreg$ilo_govspe))

# Variable Group-Centering ------------------------------------------------$
dfreg <- 
  dfreg %>% 
  mutate(to_dummy(female),
         # to_dummy(union),
         to_dummy(workst),
         # dummy for categorical variables-------------------------------------$
         to_dummy(Q03pcm),
         # to_dummy(dclass3spo_V)
  )
dfreg$female_gc = group_center(dfreg$female_2, grp = dfreg$country2)
# dfreg$union_gc = group_center(dfreg$union_2, grp = dfreg$country2)
dfreg$workst_gc = group_center(dfreg$workst_2, grp = dfreg$country2)
dfreg$edyears_gc = group_center(dfreg$edyears, grp = dfreg$country2)
dfreg$agenum_gc = group_center(dfreg$agenum, grp = dfreg$country2)
dfreg$age2_gc = group_center(dfreg$age2, grp = dfreg$country2)
# dfreg$socialtrust_gc = group_center(dfreg$socialtrust, grp = dfreg$country2)
dfreg$homclass3_V_gc <-group_center(dfreg$homclass3_V_res, grp = dfreg$country2)
dfreg$know_total_gc = group_center(dfreg$know_total, grp = dfreg$country2)
# dfreg$religion_gc = group_center(as.numeric(dfreg$religion), grp = dfreg$country2)
dfreg$partner_gc = group_center(as.numeric(dfreg$partner), grp = dfreg$country2)

# Household Income Tercile
dfreg$Q03pcm_1_gc = group_center(dfreg$Q03pcm_1, grp = dfreg$country2)
dfreg$Q03pcm_2_gc = group_center(dfreg$Q03pcm_2, grp = dfreg$country2)
dfreg$Q03pcm_3_gc = group_center(dfreg$Q03pcm_3, grp = dfreg$country2)
dfreg$Q03pcm_NA_gc = group_center(dfreg$Q03pcm_4, grp = dfreg$country2)

# Partner Social Class
# dfreg$dclass3spo_V_1_gc = group_center(dfreg$dclass3spo_V_1, grp = dfreg$country2)
# dfreg$dclass3spo_V_2_gc = group_center(dfreg$dclass3spo_V_2, grp = dfreg$country2)
# dfreg$dclass3spo_V_3_gc = group_center(dfreg$dclass3spo_V_3, grp = dfreg$country2)
# dfreg$dclass3spo_V_NA_gc = group_center(dfreg$dclass3spo_V_4, grp = dfreg$country2)

# drop dummies
dfreg <- dplyr::select(dfreg,-c(female_1,female_2,
                                # union_1,union_2,
                                workst_1,
                                workst_2,Q03pcm_1,Q03pcm_2,Q03pcm_2,Q03pcm_4,
                                # dclass3spo_V_1,dclass3spo_V_2,dclass3spo_V_3,dclass3spo_V_4
                                ))

# dfreg$gini_disp <- scale(dfreg$gini_disp)

## Model -  EGP3 - DIGCLASS - V to Working -----------------------------$
homo_V_gini <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +loggdppercapita +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp_wstate2 <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +loggdppercapita + wstate2+
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp_controls <-
  lmer(egal~
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +loggdppercapita + wstate2 +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_full2 <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +loggdppercapita + wstate2+
         homclass3_V_gc*
         dclass3res_V*
         gini_disp + 
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

digclas1_macro <- list(homo_V_gini,
                       homo_V_gini_gdp,
                       homo_V_gini_gdp_wstate2,
                       homo_V_gini_full2)

ccoef <- list(homclass3_V_gc="Class-based network homogeneity (CWC)",
              "dclass3res_VIntermediate class (III+IV)"="Intermediate Class",
              "dclass3res_VWorking Class (V+VI+VII)"="Working Class",
              "gini_disp" = "Income inequality (Ratio 90/50)",
              loggdppercapita = "GDP/capita",
              wstate2="Size of the welfare state",
              "homclass3_V_gc:dclass3res_VIntermediate class (III+IV)" = "Homogeneity*Intermediate Class",
              "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII)" = "Homogeneity*Working Class",
              "homclass3_V_gc:gini_disp"= "Homogeneity*Income Inequality",
              "homclass3_V_gc:dclass3res_VIntermediate class (III+IV):gini_disp" = "Homogeneity*Intermediate Class*Income Inequality",
              "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII):gini_disp" = "Homogeneity*Working Class *Income Inequality")

texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.50,
       caption = "Multilevel models for Income Inequality, Network segregation and Redistributive Preferences",
       custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars",
       caption.above = T,leading.zero = T,single.row = T,
       custom.coef.map = ccoef,
      groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Homogeneity*Social Class"=7:8, "Homogeneity * Social Class * Income Inequality"=10:11),
       float.pos = "h!",
      custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups",
                           "Var: Group (Intercept)",
                           "Var: Group Homogeneity",
                           "Var: Group Intermediate Class ",
                           "Var: Group Working Class",
                           "Cov: Group (Intercept), Homogeneity",
                           "Cov: Group (Intercept), Intermediate Class ",
                           "Cov: Group (Intercept), Working Class",
                           "Cov: Group Homogeneity, Intermediate Class",
                           "Cov: Group Homogeneity, Working Class",
                           "Cov: Group Intermediate Class, Working Class",
                           "Var: Residual"),
       use.packages = F)
```

\pagebreak
## Income inequality groups

```{r table1-ineq-groups, results='asis'}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))

if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  dclass3res_V,
  # dclass3spo_V,
  homclass3_V_res,
  know_total,
  # socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  partner,
  # religion,
  # union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  mutate(logrgdpna = log(rgdpna),
         loggdppercapita=log(gdppercapita),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN")

# dfreg$dclass3spo_V <- car::recode(dfreg$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)"))
# df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)"))

# sjPlot::tab_xtab(df1$dclass3spo_V,df1$PARTLIV)

dfreg <- na.omit(dfreg)
tab_countries<- 
  dfreg %>% 
  group_by(country3) %>% 
  summarise(gini_disp=mean(gini_disp)) %>% 
  mutate(gini_q05=ntile(gini_disp,5),
         gini_d10=ntile(gini_disp,10)) %>% as.data.frame() %>% 
  arrange(gini_d10) %>% 
  dplyr::select(-gini_disp)

dfreg <- dfreg %>% left_join(tab_countries,by="country3")

## Model -  EGP3 - DIGCLASS - V to Working -----------------------------$

library(estimatr)

# Quintile 1 (low)
digclas1_VW_1<- lm(egal~1 + homclass3_V_res +know_total+female+agenum+ partner + factor(country2),data=subset(dfreg,gini_q05==1),weights = WEIGHT)
digclas1_VW_Q1_1 <- update(digclas1_VW_1, . ~ . + dclass3res_V+female+agenum+edyears+ Q03pcm+workst)
digclas1_VW_Q1_2 <- update(digclas1_VW_Q1_1, . ~ . +dclass3res_V*homclass3_V_res)

# Quintile 2 (mid-low)
digclas1_VW_1<- lm(egal~1 + homclass3_V_res +know_total+female+agenum+ partner + factor(country2),data=subset(dfreg,gini_q05==2),weights = WEIGHT)
digclas1_VW_Q2_1 <- update(digclas1_VW_1, . ~ . + dclass3res_V+female+agenum+age2+edyears+ Q03pcm+workst)
digclas1_VW_Q2_2 <- update(digclas1_VW_Q2_1, . ~ . +dclass3res_V*homclass3_V_res)

# Quintile 3 (middle)
digclas1_VW_1<- lm(egal~1 + homclass3_V_res +know_total+female+agenum+ partner + factor(country2),data=subset(dfreg,gini_q05==3),weights = WEIGHT)
digclas1_VW_Q03_1 <- update(digclas1_VW_1, . ~ . + dclass3res_V+female+agenum+edyears+ Q03pcm+workst)
digclas1_VW_Q03_2 <- update(digclas1_VW_Q03_1, . ~ . +dclass3res_V*homclass3_V_res)

# Quintile 4 (middle-high)
digclas1_VW_1<- lm(egal~1 + homclass3_V_res +know_total+female+agenum + partner +  factor(country2),data=subset(dfreg,gini_q05==4),weights = WEIGHT)
digclas1_VW_Q04_1 <- update(digclas1_VW_1, . ~ . + dclass3res_V+female+agenum+edyears+ Q03pcm+workst)
digclas1_VW_Q04_2 <- update(digclas1_VW_Q04_1, . ~ . +dclass3res_V*homclass3_V_res)

# Quintile 5 (high)
digclas1_VW_1<- lm(egal~1 + homclass3_V_res +know_total+female+agenum+ partner + factor(country2),data=subset(dfreg,gini_q05==5),weights = WEIGHT)
digclas1_VW_Q05_1 <- update(digclas1_VW_1, . ~ . + dclass3res_V+female+agenum+age2+edyears+ Q03pcm+workst)
digclas1_VW_Q05_2<- update(digclas1_VW_Q05_1, . ~ . +dclass3res_V*homclass3_V_res)

digclas1_Q05 <- 
list(digclas1_VW_Q1_1,digclas1_VW_Q1_2,digclas1_VW_Q2_1,digclas1_VW_Q2_2,
     digclas1_VW_Q03_1,digclas1_VW_Q03_2,digclas1_VW_Q04_1,digclas1_VW_Q04_2,
     digclas1_VW_Q05_1,digclas1_VW_Q05_2)

ccoef <- list(homclass3_V_res="Class-based network homogeneity",
              know_total="Network size",
              # socialtrust="Social Trust",
              "dclass3res_VIntermediate class (III+IV)"="Intermediate Class",
              "dclass3res_VWorking Class (V+VI+VII)"="Working Class",
              # "femaleFemale"="Female (Ref. = Male)",
              # "agenum"="Age",
              # "age2"="Age2",
              "edyears"="Year of Education",
              "Q03pcmT02"="Income (T2)","Q03pcmT03"="Income (T3)","Q03pcmMissing"="Income (No information)",
              "workstNot in paid work" = "Not in paid work (Ref. = In paid work)",
              # "unionYes" ="Union Membership (Ref. = Not Unionized)",
              # "dclass3spo_VIntermediate class (III+IV)" = "Intermediate Class (Partner)",
              # "dclass3spo_VWorking Class (V+VI+VII)" = "Working Class (Partner)",              
              # "dclass3spo_VNo information (Missing, No partner)" = "No information (Not available, No partner)",
              "homclass3_V_res:dclass3res_VIntermediate class (III+IV)" = "Homogeneity*Intermediate Class",
              "homclass3_V_res:dclass3res_VWorking Class (V+VI+VII)" = "Homogeneity*Working Class")

header <- list("Q1"=1:2,"Q2"=3:4,"Q3"=5:6,"Q4"=7:8,"Q5"=9:10)
texreg(digclas1_Q05,stars = c(0.001, 0.01, 0.05, 0.1),single.row = F,include.ci = FALSE,custom.header = header,
       caption = paste("(\\#tab:table1-groups)","Fixed effects linear regression models for Class-based network segregation and Redistributive Preferences by Income Inequality Quintiles"),
       custom.note = "Note: Models include sampling weights. Gender, age and marital status are included as controls. Standard errors in parentheses. %stars",
       caption.above = T,leading.zero = T,
       custom.coef.map = ccoef,
       symbol = "+",scalebox = 0.55,
      groups = list("Social Class (Ref.= Service Class)" = 3:4, "Household Income (Ref.= Tertile I)"= 6:8,  
                    # "Partner's Social Class (Ref.= Service Class)" = 12:14, 
                    "Homogeneity x Social Class"=10:11),
      custom.gof.rows = list("Country FE"=rep("Yes",10)),
      # include.loglik = FALSE,include.aic = FALSE,
      # custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual"),
       float.pos = "h!",
      # file = "table1.html"
       use.packages = F)
```

```{r fig1-marg-ineqgrup, fig.dim=c(15,5),fig.cap='Average Marginal Effects of Network homogeneity conditioned by Social class on Redistributive Preferences by Income Inequality Groups',warning=FALSE}
marginal <- bind_rows(
plot_slopes(digclas1_VW_Q1_2, variables = "homclass3_V_res", by = "dclass3res_V",draw = F) %>% mutate(quintile=1),
plot_slopes(digclas1_VW_Q2_2, variables = "homclass3_V_res", by = "dclass3res_V",draw = F)  %>% mutate(quintile=2),
plot_slopes(digclas1_VW_Q03_2, variables = "homclass3_V_res", by = "dclass3res_V",draw = F) %>% mutate(quintile=3),
plot_slopes(digclas1_VW_Q04_2, variables = "homclass3_V_res", by = "dclass3res_V",draw = F) %>% mutate(quintile=4),
plot_slopes(digclas1_VW_Q05_2, variables = "homclass3_V_res", by = "dclass3res_V",draw = F) %>% mutate(quintile=5))
marginal$quintile <- factor(marginal$quintile,labels = c("Gini (Q1)","Gini Q2","Gini Q3","Gini Q4","Gini Q5"))
marginal$dclass3res_V <- forcats::fct_rev(marginal$dclass3res_V)

ggplot(data=marginal,aes(x = dclass3res_V, y = estimate,ymin= conf.low, ymax=conf.high)) +
  geom_point(size=2) + 
  geom_errorbar(width=.2)+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))+
  geom_hline(yintercept = 0,color="red") +
  facet_grid(~quintile) +
  labs(x=NULL,y="Redistributive preferences"
       # title = "Average Marginal Effects for Class-based network segregation on Redistributive preferences by Social Class and Income Inequality groups"
       )

```

\pagebreak

<!-- ## OECD Countries -->

```{r tab2-ocde, eval=FALSE, cache=TRUE, include=FALSE, results='asis'}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1) 
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2

df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)"))

dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  # starts_with("class3"),
  # starts_with("digclass3"),
  # starts_with("dclass3"),
  # starts_with("homclass"),
  dclass3res_V,
  # dclass3spo_V,
  homclass3_V_res,
  know_total,
  # socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  # religion,
  partner,
  # union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         gini_disp=scale(gini_disp))

dfreg$wstate2 <- scale(scale(dfreg$abs_red*100) + scale(dfreg$ilo_taxrev) + scale(dfreg$ilo_govspe))

# Variable Group-Centering ------------------------------------------------$
dfreg <- 
  dfreg %>% 
  mutate(to_dummy(female),
         # to_dummy(union),
         to_dummy(workst),
         # dummy for categorical variables-------------------------------------$
         to_dummy(Q03pcm),
         # to_dummy(dclass3spo_V)
  )
dfreg$female_gc = group_center(dfreg$female_2, grp = dfreg$country2)
# dfreg$union_gc = group_center(dfreg$union_2, grp = dfreg$country2)
dfreg$workst_gc = group_center(dfreg$workst_2, grp = dfreg$country2)
dfreg$edyears_gc = group_center(dfreg$edyears, grp = dfreg$country2)
dfreg$agenum_gc = group_center(dfreg$agenum, grp = dfreg$country2)
dfreg$age2_gc = group_center(dfreg$age2, grp = dfreg$country2)
# dfreg$socialtrust_gc = group_center(dfreg$socialtrust, grp = dfreg$country2)
dfreg$homclass3_V_gc <-group_center(dfreg$homclass3_V_res, grp = dfreg$country2)
dfreg$know_total_gc = group_center(dfreg$know_total, grp = dfreg$country2)
# dfreg$religion_gc = group_center(as.numeric(dfreg$religion), grp = dfreg$country2)
dfreg$partner_gc = group_center(as.numeric(dfreg$partner), grp = dfreg$country2)

# Household Income Tercile
dfreg$Q03pcm_1_gc = group_center(dfreg$Q03pcm_1, grp = dfreg$country2)
dfreg$Q03pcm_2_gc = group_center(dfreg$Q03pcm_2, grp = dfreg$country2)
dfreg$Q03pcm_3_gc = group_center(dfreg$Q03pcm_3, grp = dfreg$country2)
dfreg$Q03pcm_NA_gc = group_center(dfreg$Q03pcm_4, grp = dfreg$country2)

# Partner Social Class
# dfreg$dclass3spo_V_1_gc = group_center(dfreg$dclass3spo_V_1, grp = dfreg$country2)
# dfreg$dclass3spo_V_2_gc = group_center(dfreg$dclass3spo_V_2, grp = dfreg$country2)
# dfreg$dclass3spo_V_3_gc = group_center(dfreg$dclass3spo_V_3, grp = dfreg$country2)
# dfreg$dclass3spo_V_NA_gc = group_center(dfreg$dclass3spo_V_4, grp = dfreg$country2)

# drop dummies
dfreg <- dplyr::select(dfreg,-c(female_1,female_2,
                                # union_1,union_2,
                                workst_1,workst_2,Q03pcm_1,Q03pcm_2,Q03pcm_2,Q03pcm_4,
                                # dclass3spo_V_1,dclass3spo_V_2,dclass3spo_V_3,dclass3spo_V_4
                                ))

# dfreg$gini_disp <- scale(dfreg$gini_disp)

## Model -  EGP3 - DIGCLASS - V to Working -----------------------------$
homo_V_gini <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +loggdppercapita +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp_wstate2 <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +loggdppercapita + wstate2+
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp_controls <-
  lmer(egal~
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +loggdppercapita + wstate2 +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_full2 <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +loggdppercapita + wstate2+
         homclass3_V_gc*
         dclass3res_V*
         gini_disp + 
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

digclas1_macro <- list(homo_V_gini,
                       homo_V_gini_gdp,
                       homo_V_gini_gdp_wstate2,
                       homo_V_gini_full2)

ccoef <- list(homclass3_V_gc="Class-based network homogeneity (CWC)",
              "dclass3res_VIntermediate class (III+IV)"="Intermediate Class",
              "dclass3res_VWorking Class (V+VI+VII)"="Working Class",
              "gini_disp" = "Income inequality (Gini index)",
              loggdppercapita = "GDP/capita",
              wstate2="Size of the welfare state",
              "homclass3_V_gc:dclass3res_VIntermediate class (III+IV)" = "Homogeneity*Intermediate Class",
              "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII)" = "Homogeneity*Working Class",
              "homclass3_V_gc:gini_disp"= "Homogeneity*Income Inequality",
              "homclass3_V_gc:dclass3res_VIntermediate class (III+IV):gini_disp" = "Homogeneity*Intermediate Class*Income Inequality",
              "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII):gini_disp" = "Homogeneity*Working Class *Income Inequality")

texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.60,
       caption = "Multilevel models for Income Inequality, Network segregation and Redistributive Preferences",
       custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars",
       caption.above = T,leading.zero = T,single.row = T,
       custom.coef.map = ccoef,
      groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Homogeneity*Social Class"=7:8, "Homogeneity * Social Class * Income Inequality"=10:11),
       float.pos = "h!",
      custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups",
                           "Var: Group (Intercept)",
                           "Var: Group Homogeneity",
                           "Var: Group Intermediate Class ",
                           "Var: Group Working Class",
                           "Cov: Group (Intercept), Homogeneity",
                           "Cov: Group (Intercept), Intermediate Class ",
                           "Cov: Group (Intercept), Working Class",
                           "Cov: Group Homogeneity, Intermediate Class",
                           "Cov: Group Homogeneity, Working Class",
                           "Cov: Group Intermediate Class, Working Class",
                           "Var: Residual"),
       use.packages = F)
```

```{r fig5-ocde, eval = FALSE,include = FALSE,fig.dim=c(12,7),fig.cap='Three-way interaccion effects for Redistributive Preferences, Network segregation, Social class and Income Inequality'}
### Predicted values homo*class*GiniD ---------------------------------------$
# ginid_max<- round(max(dfreg$gini_disp) ,2)
# ginid_min<- round(min(dfreg$gini_disp),2)

ginid_max<- round(mean(dfreg$gini_disp) + sd(dfreg$gini_disp)  ,2)
ginid_min<- round(mean(dfreg$gini_disp) -sd(dfreg$gini_disp) ,2)
ginid_mea<- round(mean(dfreg$gini_disp),2)

df_pred_giniD_gc <-
  predictions(
    homo_V_gini_full2,newdata = datagrid(
      gini_disp = c(ginid_min, ginid_mea, ginid_max),
      wstate2=mean(dfreg$wstate2),
      loggdppercapita=mean(dfreg$loggdppercapita),
      homclass3_V_gc = seq(min(dfreg$homclass3_V_gc)+0.01, max(dfreg$homclass3_V_gc)+0.01, by=0.1),
      dclass3res_V=levels(dfreg$dclass3res_V)
    )
  )

df_pred_giniD_gc <- as.data.frame(df_pred_giniD_gc)
df_pred_giniD_gc$gini_disp <- factor(df_pred_giniD_gc$gini_disp,labels = c("Gini (-1SD)","Gini (Mean)", "Gini (+1SD)"))
df_pred_giniD_gc <- df_pred_giniD_gc %>% filter(dclass3res_V!="Intermediate class (III+IV)")

df_pred_giniD_gc %>% 
  ggplot(aes(y=estimate,x=homclass3_V_gc, fill=dclass3res_V,color=dclass3res_V,ymin=conf.low, ymax=conf.high)) +
  geom_ribbon(alpha=0.8,size=0.7,linetype="solid",color="black") +
  geom_line(size=1,color="black") +
  geom_point(size=3, shape = 21,color="black") +
  facet_wrap(~gini_disp) +
  labs(y = "Preferences for Redistribution",
       x="Network homogeneity (CWC)",
       caption = paste0("N=",nobs(homo_V_gini_full2),", Countries=",lme4::ngrps(homo_V_gini_full2))
       ) +
  scale_x_continuous(sec.axis = sec_axis(~ . , name = "Income inequality (post-tax) \n ", breaks = NULL, labels = NULL),
                     ) +

   scale_fill_manual(values=c("#323232","#CCCCCC"))+
  scale_color_manual(values=c("#323232","#CCCCCC"))+
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        axis.text=element_text(size=15),
        legend.title = element_blank(),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black"))
```

\pagebreak

## Social Distance by ISEI

First, the average scores of the International Socio-Economic Index of Occupational Status (ISEI) [@ganzeboom_new_2010] for each occupation of the position generator are calculated. Second, the ISEI score of the respondent (R's) is subtracted from the average ISEI points of the personal network. For example, if the R's has an ISEI of 80 and the network ISEI is 50, the social distance will be 30 (80 - 50), a "upward" social distance. Another case could be 50 (R's) minus 80 (network), and the average social distance will be - 30 or "downward" social distance. In addition, when the distance is 0, the network is entirely homogeneous.

To facilitate the interpretation of the indicator, the homogeneity indicator based on social distance is calculated:

1. The absolute values are calculated to represent the total distance to occupations concerning R's ISEI score.
2. Since there are values of 0 representing absolute homogeneity, the variable is rescaled by summing 1.
3. The values were inverted to make higher values represent higher homogeneity.

Thus, higher values represent greater homogeneity regarding R's ISEI score in contrast to the average network ISEI score.

```{r desc-social, eval=FALSE, include=FALSE}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))

if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4,correlation)
rm(list=ls())

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  dclass3res_V,
  dclass3spo_V,
  homclass3_V_res=socdist,
  know_total,
  socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  partner,
  religion,
  union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  mutate(logrgdpna = log(rgdpna),
         loggdppercapita=log(gdppercapita),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2,
         homclass3_V_res=homclass3_V_res/10) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN")

dfreg$dclass3spo_V <- car::recode(dfreg$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)"))
df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)"))

# sjPlot::tab_xtab(df1$dclass3spo_V,df1$PARTLIV)

dfreg <- na.omit(dfreg)
```

```{r table1-socdist, results='asis'}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))

if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)"))

dummy_df <- df1 %>% dplyr::select(id,
  egal = egal2,dclass3res_V,
  # dclass3spo_V,
  homclass3_V_res,
  know_total,
  # socialtrust,
  Q03pcm,edyears,female,agenum,
  # religion,
  partner,
  # union,
  workst,WEIGHT,region,
  "gini_disp"=wid_gini_disp,"gini_mkt",gv_spen,rel_red,abs_red,ilo_taxrev,
  ilo_govspe,d10d1=wid_p90p10,wid_p90p50,top10=wid_sharetop10,rgdpna,gdppercapita,
  oecd,country2, country, country3) %>%
  filter(country2 != "SVN") %>% 
  # filter(oecd == "OECD") %>%
  na.omit() %>% 
  dplyr::select(id)

df1$isei08sp <- car::recode(df1$isei08sp,"NA=0")
df1$isei08 <- car::recode(df1$isei08,"NA=0")

dfreg <- df1 %>% dplyr::select(id,
  egal = egal2,
  dclass3res_V=isei08,
  # dclass3spo_V=isei08sp,
  homclass3_V_res=socdist,
  know_total,
  # socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  partner,
  # religion,
  # union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  mutate(logrgdpna = log(rgdpna),
         loggdppercapita=log(gdppercapita),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2,
         dclass3res_V=dclass3res_V/10,
         # dclass3spo_V=dclass3spo_V/10,
         homclass3_V_res=homclass3_V_res/10
         ) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN")  %>% 
  filter(id %in% dummy_df$id)
dfreg <- na.omit(dfreg)

# Step 1: Calculate absolute values
absolute_values <- abs(dfreg$homclass3_V_res)
# hist(absolute_values);summary(absolute_values)

# Step 2: Add 1 to all values to eliminate zeros
modified_values <- absolute_values + 0.1
# hist(modified_values);summary(modified_values)

# Step 3: Reverse the order of the scale
reversed_values <- max(modified_values) + 0.1 - modified_values
# hist(reversed_values);summary(reversed_values) 
dfreg$homclass3_V_res <- reversed_values

## Model -  EGP3 - DIGCLASS - V to Working -----------------------------$
digclas1_VW_1<- lmer(egal~1 + homclass3_V_res + female+agenum+ partner + (1|country2),data=dfreg,weights = WEIGHT)
digclas1_VW_2<- update(digclas1_VW_1, . ~ . +know_total)
digclas1_VW_3 <- update(digclas1_VW_2, . ~ . + dclass3res_V)

digclas1_VW_4 <- update(digclas1_VW_3, . ~ . + dclass3res_V+female+agenum)
digclas1_VW_5 <- update(digclas1_VW_3, . ~ . + dclass3res_V+female+agenum+edyears+ Q03pcm+workst)
digclas1_VW_6 <- update(digclas1_VW_3, . ~ . + dclass3res_V+female+agenum+edyears+ Q03pcm+workst)
digclas1_VW_7 <- update(digclas1_VW_6, . ~ . +dclass3res_V*homclass3_V_res)
digclas1_VW <- list(digclas1_VW_1,digclas1_VW_2,
                    digclas1_VW_3,
                    # digclas1_VW_4,
                    digclas1_VW_5,digclas1_VW_6,digclas1_VW_7)

ccoef <- list(homclass3_V_res="Social Distance /10",
              know_total="Network size",
              # socialtrust="Social Trust",
              # "dclass3res_VIntermediate class (III+IV)"="Intermediate Class",
              # "dclass3res_VWorking Class (V+VI+VII)"="Working Class",
              # "femaleFemale"="Female (Ref. = Male)",
              # "agenum"="Age",
              # "age2"="Age2",
              "dclass3res_V"="ISEI/10",
              "edyears"="Year of Education",
              "Q03pcmT02"="Income (T2)","Q03pcmT03"="Income (T3)","Q03pcmMissing"="Income (No information)",
              "workstNot in paid work" = "Not in paid work (Ref. = In paid work)",
              # "unionYes" ="Union Membership (Ref. = Not Unionized)",
              # "dclass3spo_V" = "ISEI/10 (Partner)",
              "homclass3_V_res:dclass3res_V" = "Social Distance*ISEI")

texreg(digclas1_VW,stars = c(0.001, 0.01, 0.05, 0.1),
       caption = paste("(\\#tab:table2)","Multilevel models for Social Distance and Redistributive Preferences"),
       custom.note = "Note: Models include sampling weights. Gender, age, marital status and religion are included as controls. Standard errors in parentheses. %stars",
       caption.above = T,leading.zero = T,
       custom.coef.map = ccoef,
       symbol = "+",scalebox = 0.60,
      custom.gof.rows = list("Controls"=rep("Yes",6)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual"),
       float.pos = "h!",
       use.packages = F)
```

\newpage

```{r fig1-marg-socdist, fig.dim=c(10,5),fig.cap='Average Marginal Effects of Social Distance conditioned by ISEI on Redistributive Preferences',warning=FALSE}
int_homo <- digclas1_VW_7
plot_slopes(digclas1_VW_7, variables = "homclass3_V_res", condition = list("dclass3res_V" = "threenum"))+
labs(caption = paste0("N=",nobs(int_homo),", Countries=",lme4::ngrps(int_homo)),
       x="ISEI",y="Redistributive preferences")
```

```{r fig1-pred-socdist, fig.dim=c(10,5),fig.cap='Linear Predictions for Social Distance on Redistributive Preferences by ISEI',warning=FALSE}
int_homo <- digclas1_VW_7
df_pred1 <- 
  predictions(int_homo, condition = c("homclass3_V_res","dclass3res_V"),
              newdata = datagrid(homclass3_V_res = seq(min(dfreg$homclass3_V_res), max(dfreg$homclass3_V_res), by= 0.5),
                                 dclass3res_V = c(min(dfreg$dclass3res_V),mean(dfreg$dclass3res_V),max(dfreg$dclass3res_V))))

df_pred1$dclass3res_V <- factor(df_pred1$dclass3res_V,labels = c("ISEI (min)","ISEI (Mean)", "ISEI (max)"))

# names(df_pred1)
df_pred1 %>% 
ggplot(aes(y =estimate,x=homclass3_V_res, fill=dclass3res_V, color=dclass3res_V,ymin=conf.low, ymax=conf.high)) +
  geom_ribbon(alpha=0.8,size=1,linetype="solid",color="black") +
  geom_line(size=1,color="black") +
  geom_point(size=2,color="black") +
  # facet_grid(~dclass3res_V) +
  scale_fill_manual(values=c("#323232","#989898","#CCCCCC"))+
  scale_color_manual(values=c("#323232","#989898","#CCCCCC"))+
  scale_y_continuous(limits = c(40,100))+
  labs(caption = paste0("N=",nobs(int_homo),", Countries=",lme4::ngrps(int_homo)),
       x="Social distance",y="Redistributive preferences") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

\pagebreak

```{r tab2-socdist, results='asis', cache=TRUE}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2

df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)"))
dummy_df <- df1 %>% dplyr::select(id,
  egal = egal2,dclass3res_V,
  # dclass3spo_V,
  homclass3_V_res,know_total,
  # socialtrust,
  Q03pcm,edyears,female,agenum,
  # religion,
  partner,
  # union,
  workst,WEIGHT,region,
  "gini_disp"=wid_gini_disp,"gini_mkt",gv_spen,rel_red,abs_red,ilo_taxrev,
  ilo_govspe,d10d1=wid_p90p10,wid_p90p50,top10=wid_sharetop10,rgdpna,gdppercapita,
  oecd,country2, country, country3) %>%
  filter(country2 != "SVN") %>% 
  # filter(oecd == "OECD") %>%
  na.omit() %>% 
  dplyr::select(id)

df1$isei08sp <- car::recode(df1$isei08sp,"NA=0")
dfreg <- df1 %>% dplyr::select(id,
  egal = egal2,
  # starts_with("class3"),
  # starts_with("digclass3"),
  # starts_with("dclass3"),
  # starts_with("homclass"),
  dclass3res_V=isei08,
  # dclass3spo_V=isei08sp,
  homclass3_V_res=socdist,
  know_total,
  # socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  # religion,
  partner,
  # union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         gini_disp=scale(gini_disp),
         dclass3res_V=dclass3res_V/10,
         # dclass3spo_V=dclass3spo_V/10,
         homclass3_V_res=homclass3_V_res/10
         ) %>% 
  filter(id %in% dummy_df$id)

dfreg$wstate2 <- scale(scale(dfreg$abs_red*100) + scale(dfreg$ilo_taxrev) + scale(dfreg$ilo_govspe))
# Step 1: Calculate absolute values
absolute_values <- abs(dfreg$homclass3_V_res)
# hist(absolute_values);summary(absolute_values)

# Step 2: Add 1 to all values to eliminate zeros
modified_values <- absolute_values + 0.1
# hist(modified_values);summary(modified_values)

# Step 3: Reverse the order of the scale
reversed_values <- max(modified_values) + 0.1 - modified_values
# hist(reversed_values);summary(reversed_values)

dfreg$homclass3_V_res <- reversed_values

# Variable Group-Centering ------------------------------------------------$
dfreg <- 
  dfreg %>% 
  mutate(to_dummy(female),
         # to_dummy(union),
         to_dummy(workst),
         # dummy for categorical variables-------------------------------------$
         to_dummy(Q03pcm),
         # to_dummy(dclass3spo_V)
  )
dfreg$female_gc = group_center(dfreg$female_2, grp = dfreg$country2)
# dfreg$union_gc = group_center(dfreg$union_2, grp = dfreg$country2)
dfreg$workst_gc = group_center(dfreg$workst_2, grp = dfreg$country2)
dfreg$edyears_gc = group_center(dfreg$edyears, grp = dfreg$country2)
dfreg$agenum_gc = group_center(dfreg$agenum, grp = dfreg$country2)
dfreg$age2_gc = group_center(dfreg$age2, grp = dfreg$country2)
# dfreg$socialtrust_gc = group_center(dfreg$socialtrust, grp = dfreg$country2)
dfreg$homclass3_V_gc <-group_center(dfreg$homclass3_V_res, grp = dfreg$country2)
dfreg$know_total_gc = group_center(dfreg$know_total, grp = dfreg$country2)
# dfreg$religion_gc = group_center(as.numeric(dfreg$religion), grp = dfreg$country2)
dfreg$partner_gc = group_center(as.numeric(dfreg$partner), grp = dfreg$country2)

dfreg$dclass3res_V = group_center(as.numeric(dfreg$dclass3res_V), grp = dfreg$country2)

# Household Income Tercile
dfreg$Q03pcm_1_gc = group_center(dfreg$Q03pcm_1, grp = dfreg$country2)
dfreg$Q03pcm_2_gc = group_center(dfreg$Q03pcm_2, grp = dfreg$country2)
dfreg$Q03pcm_3_gc = group_center(dfreg$Q03pcm_3, grp = dfreg$country2)
dfreg$Q03pcm_NA_gc = group_center(dfreg$Q03pcm_4, grp = dfreg$country2)

# Partner Social Class
# dfreg$dclass3spo_V_1_gc = group_center(dfreg$dclass3spo_V_1, grp = dfreg$country2)
# dfreg$dclass3spo_V_2_gc = group_center(dfreg$dclass3spo_V_2, grp = dfreg$country2)
# dfreg$dclass3spo_V_3_gc = group_center(dfreg$dclass3spo_V_3, grp = dfreg$country2)
# dfreg$dclass3spo_V_NA_gc = group_center(dfreg$dclass3spo_V_4, grp = dfreg$country2)
# dfreg$dclass3spo_V_gc = group_center(dfreg$dclass3spo_V, grp = dfreg$country2)

# drop dummies
dfreg <- dplyr::select(dfreg,-c(female_1,female_2,
                                # union_1,union_2,
                                workst_1,
                                workst_2,Q03pcm_1,Q03pcm_2,Q03pcm_2,Q03pcm_4,
                                # dclass3spo_V_1,dclass3spo_V_2,dclass3spo_V_3,dclass3spo_V_4
                                ))

# dfreg$gini_disp <- scale(dfreg$gini_disp)

## Model -  EGP3 - DIGCLASS - V to Working -----------------------------$
homo_V_gini <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +loggdppercapita +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp_wstate2 <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +loggdppercapita + wstate2+
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp_controls <-
  lmer(egal~
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +loggdppercapita + wstate2 +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_full2 <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         # socialtrust_gc+
         dclass3res_V+
         # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+
         female_gc+agenum_gc+
         # age2_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         # union_gc+
         workst_gc+ 
         gini_disp +loggdppercapita + wstate2+
         homclass3_V_gc*
         dclass3res_V*
         gini_disp + 
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

digclas1_macro <- list(homo_V_gini,
                       homo_V_gini_gdp,
                       homo_V_gini_gdp_wstate2,
                       homo_V_gini_full2)

ccoef <- list(homclass3_V_gc="Social Distance (CWC)",
              "dclass3res_V"="ISEI ",
              "gini_disp" = "Income inequality (Gini index)",
              loggdppercapita = "GDP/capita",
              wstate2="Size of the welfare state",
              "homclass3_V_gc:dclass3res_V" = "Distance*ISEI",
              "homclass3_V_gc:gini_disp"= "Distance*Income Inequality",
              "dclass3res_V:gini_disp" ="ISEI*Inequality",
              "homclass3_V_gc:dclass3res_V:gini_disp" = "Distance*Working Class*Income Inequality")

texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.60,
       caption = "Multilevel models for Income Inequality, Social Distance and Redistributive Preferences",
       custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars",
       caption.above = T,leading.zero = T,single.row = T,
       custom.coef.map = ccoef,
      # groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Distance*Social Class"=7:8, "Distance * Social Class * Income Inequality"=10:11),
       float.pos = "h!",
      custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups",
                           "Var: Group (Intercept)",
                           "Var: Group Soc. Distance",
                           "Var: Group ISEI",
                           "Cov: Group (Intercept), Soc. Distance",
                           "Cov: Group (Intercept), ISEI",                           
                           "Cov: Group Soc. Distance, ISEI",
                           "Var: Residual"),
       use.packages = F)
```

```{r fig2-socdist, fig.dim=c(12,7),fig.cap='Three-way interaccion effects for Redistributive Preferences, Social Distance, ISEI and Income Inequality'}
### Predicted values homo*class*GiniD ---------------------------------------$
# ginid_max<- round(max(dfreg$gini_disp) ,2)
# ginid_min<- round(min(dfreg$gini_disp),2)

ginid_max<- round(mean(dfreg$gini_disp) + sd(dfreg$gini_disp)  ,2)
ginid_min<- round(mean(dfreg$gini_disp) -sd(dfreg$gini_disp) ,2)
ginid_mea<- round(mean(dfreg$gini_disp),2)

df_pred_giniD_gc <-
  predictions(
    homo_V_gini_full2,newdata = datagrid(
      gini_disp = c(ginid_min, ginid_mea, ginid_max),
      wstate2=mean(dfreg$wstate2),
      loggdppercapita=mean(dfreg$loggdppercapita),
      homclass3_V_gc = seq(min(dfreg$homclass3_V_gc)+0.01, max(dfreg$homclass3_V_gc)+0.01, by=0.5),
      dclass3res_V=c(min(dfreg$dclass3res_V),max(dfreg$dclass3res_V))
    )
  )

df_pred_giniD_gc <- as.data.frame(df_pred_giniD_gc)
df_pred_giniD_gc$gini_disp <- factor(df_pred_giniD_gc$gini_disp,labels = c("Gini (-1SD)","Gini (Mean)", "Gini (+1SD)"))

df_pred_giniD_gc$dclass3res_V <- factor(df_pred_giniD_gc$dclass3res_V,labels = c("ISEI (min)","ISEI (max)"))
# df_pred_giniD_gc <- df_pred_giniD_gc %>% filter(dclass3res_V!="Intermediate class (III+IV)")

df_pred_giniD_gc$dclass3res_V <- as.factor(df_pred_giniD_gc$dclass3res_V)

df_pred_giniD_gc %>% 
  ggplot(aes(y=estimate,x=homclass3_V_gc, fill=dclass3res_V,color=dclass3res_V,ymin=conf.low, ymax=conf.high)) +
  geom_ribbon(alpha=0.8,size=0.7,linetype="solid",color="black") +
  geom_line(size=1,color="black") +
  geom_point(size=3, shape = 21,color="black") +
  facet_wrap(~gini_disp) +
  labs(y = "Preferences for Redistribution",
       x="Social Distance (CWC)",
       caption = paste0("N=",nobs(homo_V_gini_full2),", Countries=",lme4::ngrps(homo_V_gini_full2))
       ) +
  scale_x_continuous(sec.axis = sec_axis(~ . , name = "Income inequality (post-tax) \n ", breaks = NULL, labels = NULL),
                     ) +

   scale_fill_manual(values=c("#323232","#CCCCCC"))+
  scale_color_manual(values=c("#323232","#CCCCCC"))+
  theme(legend.position = "bottom",
        legend.direction = "horizontal",
        axis.text=element_text(size=15),
        legend.title = element_blank(),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black"))
```

```{r table1-alternative, eval=FALSE, include=FALSE, results='asis'}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))

if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(
  egal = egal2,
  dclass3res_V=dclass6res,
  # homclass3_V_res=homclass3_V_res_part,
  homclass3_V_res,
  know_total,
  # socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  partner,
  # religion,
  # union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  mutate(logrgdpna = log(rgdpna),
         loggdppercapita=log(gdppercapita),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN")

# dfreg$dclass3spo_V <- car::recode(dfreg$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)"))
# df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)"))

# sjPlot::tab_xtab(df1$dclass3spo_V,df1$PARTLIV)

dfreg <- na.omit(dfreg)

## Model -  EGP3 - DIGCLASS - V to Working -----------------------------$
digclas1_VW_1<- lmer(egal~1 + homclass3_V_res + female+agenum+(1|country2),data=dfreg,weights = WEIGHT)
digclas1_VW_2<- lmer(egal~1 + homclass3_V_res + female+agenum+ partner +  (1|country2),data=dfreg,weights = WEIGHT)
digclas1_VW_3<- lmer(egal~1 + homclass3_V_res + female+agenum+ partner +  (1|country2),data=dfreg,weights = WEIGHT)
digclas1_VW_4<- lmer(egal~1 + homclass3_V_res + female+agenum+ partner +  know_total + (1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_6<- lmer(egal~1 + homclass3_V_res + female+agenum+ 
                       partner +  know_total +   
                       dclass3res_V + (1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_7<- lmer(egal~1 + homclass3_V_res + female+agenum+
                       partner +   know_total +  
                       dclass3res_V +  edyears + (1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_8<- lmer(egal~1 + homclass3_V_res + female+agenum+
                       partner +  know_total + 
                       dclass3res_V +  edyears +  Q03pcm +(1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_9<- lmer(egal~1 + homclass3_V_res + female+agenum+
                       partner +  know_total +  
                       dclass3res_V +  edyears +  Q03pcm + workst +(1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_10<- lmer(egal~1 + homclass3_V_res + female+agenum+
                       partner +  know_total + 
                       dclass3res_V +  edyears +  Q03pcm + workst + (1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_10a<- lmer(egal~1 + homclass3_V_res + female+agenum+
                       partner +  know_total + 
                       dclass3res_V +  edyears +  Q03pcm + workst + (homclass3_V_res|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_11 <- update(digclas1_VW_10, . ~ . +dclass3res_V*homclass3_V_res)

digclas1_VW <- list(digclas1_VW_1,digclas1_VW_2,digclas1_VW_3,digclas1_VW_4,
                    digclas1_VW_6,digclas1_VW_7,digclas1_VW_8,digclas1_VW_10,digclas1_VW_10a,
                    digclas1_VW_11)

screenreg(digclas1_VW)
plot_slopes(digclas1_VW_11, variables = "homclass3_V_res", by = "dclass3res_V",draw = T)
```

<!-- # Additional analysis -->

<!-- ## OECD Countries -->

<!-- ```{r tab2-ocde, results='asis', cache=TRUE} -->
<!-- set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1") -->
<!-- options(ggplot2.discrete.fill = set1) -->
<!-- options(ggplot2.discrete.colour = set1) -->
<!-- ggplot2::theme_set(ggplot2::theme_bw()) -->
<!-- ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif")) -->
<!-- options(scipen=999) -->
<!-- if (!require("pacman")) install.packages("pacman") -->
<!-- pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4) -->
<!-- rm(list=ls()) -->
<!-- load(here::here("input/data/proc/study1_country.RData"));df1 <- df2 -->

<!-- df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)")) -->

<!-- dfreg <- df1 %>% dplyr::select( -->
<!--   egal = egal2, -->
<!--   # starts_with("class3"), -->
<!--   # starts_with("digclass3"), -->
<!--   # starts_with("dclass3"), -->
<!--   # starts_with("homclass"), -->
<!--   dclass3res_V, -->
<!--   dclass3spo_V, -->
<!--   homclass3_V_res, -->
<!--   know_total, -->
<!--   socialtrust, -->
<!--   Q03pcm, -->
<!--   edyears, -->
<!--   female, -->
<!--   agenum, -->
<!--   religion, -->
<!--   partner, -->
<!--   union, -->
<!--   workst, -->
<!--   WEIGHT, -->
<!--   region, -->
<!--   "gini_disp"=wid_gini_disp, -->
<!--   "gini_mkt", -->
<!--   gv_spen, -->
<!--   rel_red, -->
<!--   abs_red, -->
<!--   ilo_taxrev, -->
<!--   ilo_govspe, -->
<!--   d10d1=wid_p90p10, -->
<!--   wid_p90p50, -->
<!--   top10=wid_sharetop10, -->
<!--   rgdpna, -->
<!--   gdppercapita, -->
<!--   oecd, -->
<!--   country2, country, country3, -->
<!-- ) %>% -->
<!--   filter(oecd == "OECD") %>% -->
<!--   filter(country2 != "SVN") %>%  -->
<!--   na.omit() %>%  -->
<!--     mutate(loggdppercapita=scale(gdppercapita/1000), -->
<!--          edyears2 = edyears ^ 2, -->
<!--          age2 = agenum ^ 2,  -->
<!--          gini_disp=scale(gini_disp)) -->

<!-- dfreg$wstate2 <- scale(scale(dfreg$abs_red*100) + scale(dfreg$ilo_taxrev) + scale(dfreg$ilo_govspe)) -->

<!-- # Variable Group-Centering ------------------------------------------------ -->
<!-- dfreg <-  -->
<!--   dfreg %>%  -->
<!--   mutate(to_dummy(female), -->
<!--          to_dummy(union), -->
<!--          to_dummy(workst), -->
<!--          # dummy for categorical variables------------------------------------- -->
<!--          to_dummy(Q03pcm), -->
<!--          to_dummy(dclass3spo_V) -->
<!--   ) -->
<!-- dfreg$female_gc = group_center(dfreg$female_2, grp = dfreg$country2) -->
<!-- dfreg$union_gc = group_center(dfreg$union_2, grp = dfreg$country2) -->
<!-- dfreg$workst_gc = group_center(dfreg$workst_2, grp = dfreg$country2) -->
<!-- dfreg$edyears_gc = group_center(dfreg$edyears, grp = dfreg$country2) -->
<!-- dfreg$agenum_gc = group_center(dfreg$agenum, grp = dfreg$country2) -->
<!-- dfreg$age2_gc = group_center(dfreg$age2, grp = dfreg$country2) -->
<!-- dfreg$socialtrust_gc = group_center(dfreg$socialtrust, grp = dfreg$country2) -->
<!-- dfreg$homclass3_V_gc <-group_center(dfreg$homclass3_V_res, grp = dfreg$country2) -->
<!-- dfreg$know_total_gc = group_center(dfreg$know_total, grp = dfreg$country2) -->
<!-- dfreg$religion_gc = group_center(as.numeric(dfreg$religion), grp = dfreg$country2) -->
<!-- dfreg$partner_gc = group_center(as.numeric(dfreg$partner), grp = dfreg$country2) -->

<!-- # Household Income Tercile -->
<!-- dfreg$Q03pcm_1_gc = group_center(dfreg$Q03pcm_1, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_2_gc = group_center(dfreg$Q03pcm_2, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_3_gc = group_center(dfreg$Q03pcm_3, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_NA_gc = group_center(dfreg$Q03pcm_4, grp = dfreg$country2) -->

<!-- # Partner Social Class -->
<!-- dfreg$dclass3spo_V_1_gc = group_center(dfreg$dclass3spo_V_1, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_2_gc = group_center(dfreg$dclass3spo_V_2, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_3_gc = group_center(dfreg$dclass3spo_V_3, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_NA_gc = group_center(dfreg$dclass3spo_V_4, grp = dfreg$country2) -->

<!-- # drop dummies -->
<!-- dfreg <- dplyr::select(dfreg,-c(female_1,female_2,union_1,union_2,workst_1, -->
<!--                                 workst_2,Q03pcm_1,Q03pcm_2,Q03pcm_2,Q03pcm_4, -->
<!--                                 dclass3spo_V_1,dclass3spo_V_2,dclass3spo_V_3,dclass3spo_V_4)) -->

<!-- # dfreg$gini_disp <- scale(dfreg$gini_disp) -->

<!-- ## Model -  EGP3 - DIGCLASS - V to Working ----------------------------- -->
<!-- homo_V_gini <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp_wstate2 <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2+ -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp_controls <- -->
<!--   lmer(egal~ -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2 + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_full2 <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2+ -->
<!--          homclass3_V_gc* -->
<!--          dclass3res_V* -->
<!--          gini_disp +  -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- digclas1_macro <- list(homo_V_gini, -->
<!--                        homo_V_gini_gdp, -->
<!--                        homo_V_gini_gdp_wstate2, -->
<!--                        homo_V_gini_full2) -->

<!-- ccoef <- list(homclass3_V_gc="Class-based network homogeneity (CWC)", -->
<!--               "dclass3res_VIntermediate class (III+IV)"="Intermediate Class", -->
<!--               "dclass3res_VWorking Class (V+VI+VII)"="Working Class", -->
<!--               "gini_disp" = "Income inequality (Gini index)", -->
<!--               loggdppercapita = "GDP/capita", -->
<!--               wstate2="Size of the welfare state", -->
<!--               "homclass3_V_gc:dclass3res_VIntermediate class (III+IV)" = "Homogeneity*Intermediate Class", -->
<!--               "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII)" = "Homogeneity*Working Class", -->
<!--               "homclass3_V_gc:gini_disp"= "Homogeneity*Income Inequality", -->
<!--               "homclass3_V_gc:dclass3res_VIntermediate class (III+IV):gini_disp" = "Homogeneity*Intermediate Class*Income Inequality", -->
<!--               "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII):gini_disp" = "Homogeneity*Working Class *Income Inequality") -->

<!-- texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.60, -->
<!--        caption = "Multilevel models for Income Inequality, Network segregation and Redistributive Preferences", -->
<!--        custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars", -->
<!--        caption.above = T,leading.zero = T,single.row = T, -->
<!--        custom.coef.map = ccoef, -->
<!--       groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Homogeneity*Social Class"=7:8, "Homogeneity * Social Class * Income Inequality"=10:11), -->
<!--        float.pos = "h!", -->
<!--       custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE, -->
<!--       custom.gof.names = c(NA,NA,"Num. groups", -->
<!--                            "Var: Group (Intercept)", -->
<!--                            "Var: Group Homogeneity", -->
<!--                            "Var: Group Intermediate Class ", -->
<!--                            "Var: Group Working Class", -->
<!--                            "Cov: Group (Intercept), Homogeneity", -->
<!--                            "Cov: Group (Intercept), Intermediate Class ", -->
<!--                            "Cov: Group (Intercept), Working Class", -->
<!--                            "Cov: Group Homogeneity, Intermediate Class", -->
<!--                            "Cov: Group Homogeneity, Working Class", -->
<!--                            "Cov: Group Intermediate Class, Working Class", -->
<!--                            "Var: Residual"), -->
<!--        use.packages = F) -->
<!-- ``` -->

<!-- ```{r fig5-ocde, fig.dim=c(12,7),fig.cap='Three-way interaccion effects for Redistributive Preferences, Network segregation, Social class and Income Inequality'} -->
<!-- ### Predicted values homo*class*GiniD --------------------------------------- -->
<!-- # ginid_max<- round(max(dfreg$gini_disp) ,2) -->
<!-- # ginid_min<- round(min(dfreg$gini_disp),2) -->

<!-- ginid_max<- round(mean(dfreg$gini_disp) + sd(dfreg$gini_disp)  ,2) -->
<!-- ginid_min<- round(mean(dfreg$gini_disp) -sd(dfreg$gini_disp) ,2) -->
<!-- ginid_mea<- round(mean(dfreg$gini_disp),2) -->

<!-- df_pred_giniD_gc <- -->
<!--   predictions( -->
<!--     homo_V_gini_full2,newdata = datagrid( -->
<!--       gini_disp = c(ginid_min, ginid_mea, ginid_max), -->
<!--       wstate2=mean(dfreg$wstate2), -->
<!--       loggdppercapita=mean(dfreg$loggdppercapita), -->
<!--       homclass3_V_gc = seq(min(dfreg$homclass3_V_gc)+0.01, max(dfreg$homclass3_V_gc)+0.01, by=0.1), -->
<!--       dclass3res_V=levels(dfreg$dclass3res_V) -->
<!--     ) -->
<!--   ) -->

<!-- df_pred_giniD_gc <- as.data.frame(df_pred_giniD_gc) -->
<!-- df_pred_giniD_gc$gini_disp <- factor(df_pred_giniD_gc$gini_disp,labels = c("Gini (-1SD)","Gini (Mean)", "Gini (+1SD)")) -->
<!-- df_pred_giniD_gc <- df_pred_giniD_gc %>% filter(dclass3res_V!="Intermediate class (III+IV)") -->

<!-- df_pred_giniD_gc %>%  -->
<!--   ggplot(aes(y=estimate,x=homclass3_V_gc, fill=dclass3res_V,color=dclass3res_V,ymin=conf.low, ymax=conf.high)) + -->
<!--   geom_ribbon(alpha=0.8,size=0.7,linetype="solid",color="black") + -->
<!--   geom_line(size=1,color="black") + -->
<!--   geom_point(size=3, shape = 21,color="black") + -->
<!--   facet_wrap(~gini_disp) + -->
<!--   labs(y = "Preferences for Redistribution", -->
<!--        x="Network homogeneity (CWC)", -->
<!--        caption = paste0("N=",nobs(homo_V_gini_full2),", Countries=",lme4::ngrps(homo_V_gini_full2)) -->
<!--        ) + -->
<!--   scale_x_continuous(sec.axis = sec_axis(~ . , name = "Income inequality (post-tax) \n ", breaks = NULL, labels = NULL), -->
<!--                      ) + -->

<!--    scale_fill_manual(values=c("#323232","#CCCCCC"))+ -->
<!--   scale_color_manual(values=c("#323232","#CCCCCC"))+ -->
<!--   theme(legend.position = "bottom", -->
<!--         legend.direction = "horizontal", -->
<!--         axis.text=element_text(size=15), -->
<!--         legend.title = element_blank(), -->
<!--         legend.background = element_blank(), -->
<!--         legend.box.background = element_rect(colour = "black")) -->
<!-- ``` -->

<!-- ## Income Inequality Ratios (OECD) -->

<!-- ```{r tab2-ratio9010, results='asis', cache=TRUE} -->
<!-- set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1") -->
<!-- options(ggplot2.discrete.fill = set1) -->
<!-- options(ggplot2.discrete.colour = set1) -->
<!-- ggplot2::theme_set(ggplot2::theme_bw()) -->
<!-- ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif")) -->
<!-- options(scipen=999) -->
<!-- if (!require("pacman")) install.packages("pacman") -->
<!-- pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)   -->
<!-- rm(list=ls()) -->

<!-- load(here::here("input/data/proc/study1_country.RData"));df1 <- df2 -->

<!-- df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)")) -->

<!-- dfreg <- df1 %>% dplyr::select( -->
<!--   egal = egal2, -->
<!--   # starts_with("class3"), -->
<!--   # starts_with("digclass3"), -->
<!--   # starts_with("dclass3"), -->
<!--   # starts_with("homclass"), -->
<!--   dclass3res_V, -->
<!--   dclass3spo_V, -->
<!--   homclass3_V_res, -->
<!--   know_total, -->
<!--   socialtrust, -->
<!--   Q03pcm, -->
<!--   edyears, -->
<!--   female, -->
<!--   agenum, -->
<!--   religion, -->
<!--   partner, -->
<!--   union, -->
<!--   workst, -->
<!--   WEIGHT, -->
<!--   region, -->
<!--   "gini_disp"=wid_gini_disp, -->
<!--   "gini_mkt", -->
<!--   gv_spen, -->
<!--   rel_red, -->
<!--   abs_red, -->
<!--   ilo_taxrev, -->
<!--   ilo_govspe, -->
<!--   d10d1=wid_p90p10, -->
<!--   wid_p90p50, -->
<!--   top10=wid_sharetop10, -->
<!--   rgdpna, -->
<!--   gdppercapita, -->
<!--   oecd, -->
<!--   country2, country, country3, -->
<!-- ) %>% -->
<!--   filter(oecd == "OECD") %>% -->
<!--   filter(country2 != "SVN") %>%  -->
<!--   na.omit() %>%  -->
<!--     mutate(loggdppercapita=scale(gdppercapita/1000), -->
<!--          edyears2 = edyears ^ 2, -->
<!--          age2 = agenum ^ 2,  -->
<!--          gini_disp=scale(d10d1)) -->

<!-- dfreg$wstate2 <- scale(scale(dfreg$abs_red*100) + scale(dfreg$ilo_taxrev) + scale(dfreg$ilo_govspe)) -->

<!-- # Variable Group-Centering ------------------------------------------------ -->
<!-- dfreg <-  -->
<!--   dfreg %>%  -->
<!--   mutate(to_dummy(female), -->
<!--          to_dummy(union), -->
<!--          to_dummy(workst), -->
<!--          # dummy for categorical variables------------------------------------- -->
<!--          to_dummy(Q03pcm), -->
<!--          to_dummy(dclass3spo_V) -->
<!--   ) -->
<!-- dfreg$female_gc = group_center(dfreg$female_2, grp = dfreg$country2) -->
<!-- dfreg$union_gc = group_center(dfreg$union_2, grp = dfreg$country2) -->
<!-- dfreg$workst_gc = group_center(dfreg$workst_2, grp = dfreg$country2) -->
<!-- dfreg$edyears_gc = group_center(dfreg$edyears, grp = dfreg$country2) -->
<!-- dfreg$agenum_gc = group_center(dfreg$agenum, grp = dfreg$country2) -->
<!-- dfreg$age2_gc = group_center(dfreg$age2, grp = dfreg$country2) -->
<!-- dfreg$socialtrust_gc = group_center(dfreg$socialtrust, grp = dfreg$country2) -->
<!-- dfreg$homclass3_V_gc <-group_center(dfreg$homclass3_V_res, grp = dfreg$country2) -->
<!-- dfreg$know_total_gc = group_center(dfreg$know_total, grp = dfreg$country2) -->
<!-- dfreg$religion_gc = group_center(as.numeric(dfreg$religion), grp = dfreg$country2) -->
<!-- dfreg$partner_gc = group_center(as.numeric(dfreg$partner), grp = dfreg$country2) -->

<!-- # Household Income Tercile -->
<!-- dfreg$Q03pcm_1_gc = group_center(dfreg$Q03pcm_1, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_2_gc = group_center(dfreg$Q03pcm_2, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_3_gc = group_center(dfreg$Q03pcm_3, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_NA_gc = group_center(dfreg$Q03pcm_4, grp = dfreg$country2) -->

<!-- # Partner Social Class -->
<!-- dfreg$dclass3spo_V_1_gc = group_center(dfreg$dclass3spo_V_1, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_2_gc = group_center(dfreg$dclass3spo_V_2, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_3_gc = group_center(dfreg$dclass3spo_V_3, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_NA_gc = group_center(dfreg$dclass3spo_V_4, grp = dfreg$country2) -->

<!-- # drop dummies -->
<!-- dfreg <- dplyr::select(dfreg,-c(female_1,female_2,union_1,union_2,workst_1, -->
<!--                                 workst_2,Q03pcm_1,Q03pcm_2,Q03pcm_2,Q03pcm_4, -->
<!--                                 dclass3spo_V_1,dclass3spo_V_2,dclass3spo_V_3,dclass3spo_V_4)) -->

<!-- # dfreg$gini_disp <- scale(dfreg$gini_disp) -->

<!-- ## Model -  EGP3 - DIGCLASS - V to Working ----------------------------- -->
<!-- homo_V_gini <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp_wstate2 <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2+ -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp_controls <- -->
<!--   lmer(egal~ -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2 + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_full2 <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2+ -->
<!--          homclass3_V_gc* -->
<!--          dclass3res_V* -->
<!--          gini_disp +  -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- digclas1_macro <- list(homo_V_gini, -->
<!--                        homo_V_gini_gdp, -->
<!--                        homo_V_gini_gdp_wstate2, -->
<!--                        homo_V_gini_full2) -->

<!-- ccoef <- list(homclass3_V_gc="Class-based network homogeneity (CWC)", -->
<!--               "dclass3res_VIntermediate class (III+IV)"="Intermediate Class", -->
<!--               "dclass3res_VWorking Class (V+VI+VII)"="Working Class", -->
<!--               "gini_disp" = "Income inequality (Ratio 90/10)", -->
<!--               loggdppercapita = "GDP/capita", -->
<!--               wstate2="Size of the welfare state", -->
<!--               "homclass3_V_gc:dclass3res_VIntermediate class (III+IV)" = "Homogeneity*Intermediate Class", -->
<!--               "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII)" = "Homogeneity*Working Class", -->
<!--               "homclass3_V_gc:gini_disp"= "Homogeneity*Income Inequality", -->
<!--               "homclass3_V_gc:dclass3res_VIntermediate class (III+IV):gini_disp" = "Homogeneity*Intermediate Class*Income Inequality", -->
<!--               "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII):gini_disp" = "Homogeneity*Working Class *Income Inequality") -->

<!-- texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.50, -->
<!--        caption = "Multilevel models for Income Inequality, Network segregation and Redistributive Preferences", -->
<!--        custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars", -->
<!--        caption.above = T,leading.zero = T,single.row = T, -->
<!--        custom.coef.map = ccoef, -->
<!--       groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Homogeneity*Social Class"=7:8, "Homogeneity * Social Class * Income Inequality"=10:11), -->
<!--        float.pos = "h!", -->
<!--       custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE, -->
<!--       custom.gof.names = c(NA,NA,"Num. groups", -->
<!--                            "Var: Group (Intercept)", -->
<!--                            "Var: Group Homogeneity", -->
<!--                            "Var: Group Intermediate Class ", -->
<!--                            "Var: Group Working Class", -->
<!--                            "Cov: Group (Intercept), Homogeneity", -->
<!--                            "Cov: Group (Intercept), Intermediate Class ", -->
<!--                            "Cov: Group (Intercept), Working Class", -->
<!--                            "Cov: Group Homogeneity, Intermediate Class", -->
<!--                            "Cov: Group Homogeneity, Working Class", -->
<!--                            "Cov: Group Intermediate Class, Working Class", -->
<!--                            "Var: Residual"), -->
<!--        use.packages = F) -->
<!-- ``` -->

<!-- ```{r tab2-ratio9050, results='asis', cache=TRUE} -->
<!-- set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1") -->
<!-- options(ggplot2.discrete.fill = set1) -->
<!-- options(ggplot2.discrete.colour = set1) -->
<!-- ggplot2::theme_set(ggplot2::theme_bw()) -->
<!-- ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif")) -->
<!-- options(scipen=999) -->
<!-- if (!require("pacman")) install.packages("pacman") -->
<!-- pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)   -->
<!-- rm(list=ls()) -->

<!-- load(here::here("input/data/proc/study1_country.RData"));df1 <- df2 -->

<!-- df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)")) -->

<!-- dfreg <- df1 %>% dplyr::select( -->
<!--   egal = egal2, -->
<!--   # starts_with("class3"), -->
<!--   # starts_with("digclass3"), -->
<!--   # starts_with("dclass3"), -->
<!--   # starts_with("homclass"), -->
<!--   dclass3res_V, -->
<!--   dclass3spo_V, -->
<!--   homclass3_V_res, -->
<!--   know_total, -->
<!--   socialtrust, -->
<!--   Q03pcm, -->
<!--   edyears, -->
<!--   female, -->
<!--   agenum, -->
<!--   religion, -->
<!--   partner, -->
<!--   union, -->
<!--   workst, -->
<!--   WEIGHT, -->
<!--   region, -->
<!--   "gini_disp"=wid_gini_disp, -->
<!--   "gini_mkt", -->
<!--   gv_spen, -->
<!--   rel_red, -->
<!--   abs_red, -->
<!--   ilo_taxrev, -->
<!--   ilo_govspe, -->
<!--   d10d1=wid_p90p10, -->
<!--   wid_p90p50, -->
<!--   top10=wid_sharetop10, -->
<!--   rgdpna, -->
<!--   gdppercapita, -->
<!--   oecd, -->
<!--   country2, country, country3, -->
<!-- ) %>% -->
<!--   filter(oecd == "OECD") %>% -->
<!--   filter(country2 != "SVN") %>%  -->
<!--   na.omit() %>%  -->
<!--     mutate(loggdppercapita=scale(gdppercapita/1000), -->
<!--          edyears2 = edyears ^ 2, -->
<!--          age2 = agenum ^ 2,  -->
<!--          gini_disp=scale(wid_p90p50)) -->

<!-- dfreg$wstate2 <- scale(scale(dfreg$abs_red*100) + scale(dfreg$ilo_taxrev) + scale(dfreg$ilo_govspe)) -->

<!-- # Variable Group-Centering ------------------------------------------------ -->
<!-- dfreg <-  -->
<!--   dfreg %>%  -->
<!--   mutate(to_dummy(female), -->
<!--          to_dummy(union), -->
<!--          to_dummy(workst), -->
<!--          # dummy for categorical variables------------------------------------- -->
<!--          to_dummy(Q03pcm), -->
<!--          to_dummy(dclass3spo_V) -->
<!--   ) -->
<!-- dfreg$female_gc = group_center(dfreg$female_2, grp = dfreg$country2) -->
<!-- dfreg$union_gc = group_center(dfreg$union_2, grp = dfreg$country2) -->
<!-- dfreg$workst_gc = group_center(dfreg$workst_2, grp = dfreg$country2) -->
<!-- dfreg$edyears_gc = group_center(dfreg$edyears, grp = dfreg$country2) -->
<!-- dfreg$agenum_gc = group_center(dfreg$agenum, grp = dfreg$country2) -->
<!-- dfreg$age2_gc = group_center(dfreg$age2, grp = dfreg$country2) -->
<!-- dfreg$socialtrust_gc = group_center(dfreg$socialtrust, grp = dfreg$country2) -->
<!-- dfreg$homclass3_V_gc <-group_center(dfreg$homclass3_V_res, grp = dfreg$country2) -->
<!-- dfreg$know_total_gc = group_center(dfreg$know_total, grp = dfreg$country2) -->
<!-- dfreg$religion_gc = group_center(as.numeric(dfreg$religion), grp = dfreg$country2) -->
<!-- dfreg$partner_gc = group_center(as.numeric(dfreg$partner), grp = dfreg$country2) -->

<!-- # Household Income Tercile -->
<!-- dfreg$Q03pcm_1_gc = group_center(dfreg$Q03pcm_1, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_2_gc = group_center(dfreg$Q03pcm_2, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_3_gc = group_center(dfreg$Q03pcm_3, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_NA_gc = group_center(dfreg$Q03pcm_4, grp = dfreg$country2) -->

<!-- # Partner Social Class -->
<!-- dfreg$dclass3spo_V_1_gc = group_center(dfreg$dclass3spo_V_1, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_2_gc = group_center(dfreg$dclass3spo_V_2, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_3_gc = group_center(dfreg$dclass3spo_V_3, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_NA_gc = group_center(dfreg$dclass3spo_V_4, grp = dfreg$country2) -->

<!-- # drop dummies -->
<!-- dfreg <- dplyr::select(dfreg,-c(female_1,female_2,union_1,union_2,workst_1, -->
<!--                                 workst_2,Q03pcm_1,Q03pcm_2,Q03pcm_2,Q03pcm_4, -->
<!--                                 dclass3spo_V_1,dclass3spo_V_2,dclass3spo_V_3,dclass3spo_V_4)) -->

<!-- # dfreg$gini_disp <- scale(dfreg$gini_disp) -->

<!-- ## Model -  EGP3 - DIGCLASS - V to Working ----------------------------- -->
<!-- homo_V_gini <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp_wstate2 <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2+ -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp_controls <- -->
<!--   lmer(egal~ -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2 + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_full2 <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2+ -->
<!--          homclass3_V_gc* -->
<!--          dclass3res_V* -->
<!--          gini_disp +  -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- digclas1_macro <- list(homo_V_gini, -->
<!--                        homo_V_gini_gdp, -->
<!--                        homo_V_gini_gdp_wstate2, -->
<!--                        homo_V_gini_full2) -->

<!-- ccoef <- list(homclass3_V_gc="Class-based network homogeneity (CWC)", -->
<!--               "dclass3res_VIntermediate class (III+IV)"="Intermediate Class", -->
<!--               "dclass3res_VWorking Class (V+VI+VII)"="Working Class", -->
<!--               "gini_disp" = "Income inequality (Ratio 90/50)", -->
<!--               loggdppercapita = "GDP/capita", -->
<!--               wstate2="Size of the welfare state", -->
<!--               "homclass3_V_gc:dclass3res_VIntermediate class (III+IV)" = "Homogeneity*Intermediate Class", -->
<!--               "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII)" = "Homogeneity*Working Class", -->
<!--               "homclass3_V_gc:gini_disp"= "Homogeneity*Income Inequality", -->
<!--               "homclass3_V_gc:dclass3res_VIntermediate class (III+IV):gini_disp" = "Homogeneity*Intermediate Class*Income Inequality", -->
<!--               "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII):gini_disp" = "Homogeneity*Working Class *Income Inequality") -->

<!-- texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.50, -->
<!--        caption = "Multilevel models for Income Inequality, Network segregation and Redistributive Preferences", -->
<!--        custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars", -->
<!--        caption.above = T,leading.zero = T,single.row = T, -->
<!--        custom.coef.map = ccoef, -->
<!--       groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Homogeneity*Social Class"=7:8, "Homogeneity * Social Class * Income Inequality"=10:11), -->
<!--        float.pos = "h!", -->
<!--       custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE, -->
<!--       custom.gof.names = c(NA,NA,"Num. groups", -->
<!--                            "Var: Group (Intercept)", -->
<!--                            "Var: Group Homogeneity", -->
<!--                            "Var: Group Intermediate Class ", -->
<!--                            "Var: Group Working Class", -->
<!--                            "Cov: Group (Intercept), Homogeneity", -->
<!--                            "Cov: Group (Intercept), Intermediate Class ", -->
<!--                            "Cov: Group (Intercept), Working Class", -->
<!--                            "Cov: Group Homogeneity, Intermediate Class", -->
<!--                            "Cov: Group Homogeneity, Working Class", -->
<!--                            "Cov: Group Intermediate Class, Working Class", -->
<!--                            "Var: Residual"), -->
<!--        use.packages = F) -->
<!-- ``` -->

<!-- \pagebreak -->

<!-- ## Social distance (OECD) -->

<!-- ```{r desc-socdist,eval=FALSE, include=FALSE} -->
<!-- set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1") -->
<!-- options(ggplot2.discrete.fill = set1) -->
<!-- options(ggplot2.discrete.colour = set1) -->
<!-- ggplot2::theme_set(ggplot2::theme_bw()) -->
<!-- ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif")) -->

<!-- if (!require("pacman")) install.packages("pacman") -->
<!-- pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4,correlation) -->
<!-- rm(list=ls()) -->

<!-- load(here::here("input/data/proc/study1_country.RData"));df1 <- df2 -->
<!-- dfreg <- df1 %>% dplyr::select( -->
<!--   egal = egal2, -->
<!--   dclass3res_V, -->
<!--   dclass3spo_V, -->
<!--   homclass3_V_res=socdist, -->
<!--   know_total, -->
<!--   socialtrust, -->
<!--   Q03pcm, -->
<!--   edyears, -->
<!--   female, -->
<!--   agenum, -->
<!--   partner, -->
<!--   religion, -->
<!--   union, -->
<!--   workst, -->
<!--   WEIGHT, -->
<!--   region, -->
<!--   "gini_disp"=wid_gini_disp, -->
<!--   "gini_mkt", -->
<!--   gv_spen, -->
<!--   rel_red, -->
<!--   d10d1=wid_p90p10, -->
<!--   wid_p90p50, -->
<!--   top10=wid_sharetop10, -->
<!--   rgdpna, -->
<!--   gdppercapita, -->
<!--   oecd, -->
<!--   country2, country, country3, -->
<!-- ) %>% -->
<!--   mutate(logrgdpna = log(rgdpna), -->
<!--          loggdppercapita=log(gdppercapita), -->
<!--          edyears2 = edyears ^ 2, -->
<!--          age2 = agenum ^ 2, -->
<!--          homclass3_V_res=homclass3_V_res/10) %>% -->
<!--   filter(oecd == "OECD") %>% -->
<!--   filter(country2 != "SVN") -->

<!-- dfreg$dclass3spo_V <- car::recode(dfreg$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)")) -->
<!-- df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)")) -->

<!-- # sjPlot::tab_xtab(df1$dclass3spo_V,df1$PARTLIV) -->

<!-- dfreg <- na.omit(dfreg) -->
<!-- ``` -->

<!-- ### Social Distance by Social Class -->

<!-- ```{r table1-socdistclass, echo=FALSE, results='asis'} -->
<!-- set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1") -->
<!-- options(ggplot2.discrete.fill = set1) -->
<!-- options(ggplot2.discrete.colour = set1) -->
<!-- ggplot2::theme_set(ggplot2::theme_bw()) -->
<!-- ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif")) -->

<!-- if (!require("pacman")) install.packages("pacman") -->
<!-- pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4) -->
<!-- rm(list=ls()) -->

<!-- load(here::here("input/data/proc/study1_country.RData"));df1 <- df2 -->
<!-- dfreg <- df1 %>% dplyr::select( -->
<!--   egal = egal2, -->
<!--   dclass3res_V, -->
<!--   dclass3spo_V, -->
<!--   homclass3_V_res=socdist, -->
<!--   know_total, -->
<!--   socialtrust, -->
<!--   Q03pcm, -->
<!--   edyears, -->
<!--   female, -->
<!--   agenum, -->
<!--   partner, -->
<!--   religion, -->
<!--   union, -->
<!--   workst, -->
<!--   WEIGHT, -->
<!--   region, -->
<!--   "gini_disp"=wid_gini_disp, -->
<!--   "gini_mkt", -->
<!--   gv_spen, -->
<!--   rel_red, -->
<!--   d10d1=wid_p90p10, -->
<!--   wid_p90p50, -->
<!--   top10=wid_sharetop10, -->
<!--   rgdpna, -->
<!--   gdppercapita, -->
<!--   oecd, -->
<!--   country2, country, country3, -->
<!-- ) %>% -->
<!--   mutate(logrgdpna = log(rgdpna), -->
<!--          loggdppercapita=log(gdppercapita), -->
<!--          edyears2 = edyears ^ 2, -->
<!--          age2 = agenum ^ 2, -->
<!--          homclass3_V_res=homclass3_V_res/10) %>% -->
<!--   filter(oecd == "OECD") %>% -->
<!--   filter(country2 != "SVN") -->

<!-- dfreg$dclass3spo_V <- car::recode(dfreg$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)")) -->
<!-- df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)")) -->

<!-- # sjPlot::tab_xtab(df1$dclass3spo_V,df1$PARTLIV) -->

<!-- dfreg <- na.omit(dfreg) -->

<!-- # Step 1: Calculate absolute values -->
<!-- absolute_values <- abs(dfreg$homclass3_V_res) -->
<!-- # hist(absolute_values);summary(absolute_values) -->

<!-- # Step 2: Add 1 to all values to eliminate zeros -->
<!-- modified_values <- absolute_values + 0.1 -->
<!-- # hist(modified_values);summary(modified_values) -->

<!-- # Step 3: Reverse the order of the scale -->
<!-- reversed_values <- max(modified_values) + 0.1 - modified_values -->
<!-- dfreg$homclass3_V_res <- reversed_values -->

<!-- ## Model -  EGP3 - DIGCLASS - V to Working ----------------------------- -->
<!-- digclas1_VW_1<- lmer(egal~1 + homclass3_V_res + female+agenum+age2 + partner + religion + (1|country2),data=dfreg,weights = WEIGHT) -->
<!-- digclas1_VW_2<- update(digclas1_VW_1, . ~ . +know_total+socialtrust) -->
<!-- digclas1_VW_3 <- update(digclas1_VW_2, . ~ . + dclass3res_V) -->

<!-- digclas1_VW_4 <- update(digclas1_VW_3, . ~ . + dclass3res_V+female+agenum+age2) -->
<!-- digclas1_VW_5 <- update(digclas1_VW_3, . ~ . + dclass3res_V+female+agenum+age2+edyears+ Q03pcm+workst+union) -->
<!-- digclas1_VW_6 <- update(digclas1_VW_3, . ~ . + dclass3res_V+female+agenum+age2+edyears+ Q03pcm+workst+union+dclass3spo_V) -->
<!-- digclas1_VW_7 <- update(digclas1_VW_6, . ~ . +dclass3res_V*homclass3_V_res) -->
<!-- digclas1_VW <- list(digclas1_VW_1,digclas1_VW_2, -->
<!--                     digclas1_VW_3, -->
<!--                     # digclas1_VW_4, -->
<!--                     digclas1_VW_5,digclas1_VW_6,digclas1_VW_7) -->

<!-- ccoef <- list(homclass3_V_res="Social distance", -->
<!--               know_total="Network size", -->
<!--               socialtrust="Social Trust", -->
<!--               "dclass3res_VIntermediate class (III+IV)"="Intermediate Class", -->
<!--               "dclass3res_VWorking Class (V+VI+VII)"="Working Class", -->
<!--               # "femaleFemale"="Female (Ref. = Male)", -->
<!--               # "agenum"="Age", -->
<!--               # "age2"="Age2", -->
<!--               "edyears"="Year of Education", -->
<!--               "Q03pcmT02"="Income (T2)","Q03pcmT03"="Income (T3)","Q03pcmMissing"="Income (No information)", -->
<!--               "workstNot in paid work" = "Not in paid work (Ref. = In paid work)", -->
<!--               "unionYes" ="Union Membership (Ref. = Not Unionized)", -->
<!--               "dclass3spo_VIntermediate class (III+IV)" = "Intermediate Class (Partner)", -->
<!--               "dclass3spo_VWorking Class (V+VI+VII)" = "Working Class (Partner)", -->
<!--               "dclass3spo_VNo information (Missing, No partner)" = "No information (Not available, No partner)", -->
<!--               "homclass3_V_res:dclass3res_VIntermediate class (III+IV)" = "Distance*Intermediate Class", -->
<!--               "homclass3_V_res:dclass3res_VWorking Class (V+VI+VII)" = "Distance*Working Class") -->

<!-- texreg(digclas1_VW,stars = c(0.001, 0.01, 0.05, 0.1), -->
<!--        caption = paste("(\\#tab:table2)","Multilevel models for Social distance and Redistributive Preferences"), -->
<!--        custom.note = "Note: Models include sampling weights. Gender, age, marital status and religion are included as controls. Standard errors in parentheses. %stars", -->
<!--        caption.above = T,leading.zero = T, -->
<!--        custom.coef.map = ccoef, -->
<!--        symbol = "+",scalebox = 0.60, -->
<!--       groups = list("Social Class (Ref.= Service Class)" = 4:5, "Household Income (Ref.= Tertile I)"= 7:9,  "Partner's Social Class (Ref.= Service Class)" = 12:14, "Social distance x Social Class"=15:16), -->
<!--       custom.gof.rows = list("Controls"=rep("Yes",6)),include.loglik = FALSE,include.aic = FALSE, -->
<!--       custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual"), -->
<!--        float.pos = "h!", -->
<!--        use.packages = F) -->
<!-- ``` -->

<!-- ```{r fig1-marg-socdistclass, echo=FALSE, fig.cap='Average Marginal Effects of Social Distance conditionated by Social Class on Redistributive Preferences', fig.dim=c(10,5), warning=FALSE} -->
<!-- int_homo <- digclas1_VW_7 -->
<!-- plot_slopes(digclas1_VW_7, variables = "homclass3_V_res", condition = list("dclass3res_V"))+ -->
<!-- labs(caption = paste0("N=",nobs(int_homo),", Countries=",lme4::ngrps(int_homo)), -->
<!--        x="Social Distance",y="Redistributive preferences") -->
<!-- ``` -->

<!-- ```{r fig1-pred-socdistclass, echo=FALSE, fig.cap='Linear Predictions for Social distance on Redistributive Preferences by Social Class', fig.dim=c(10,5), warning=FALSE} -->
<!-- int_homo <- digclas1_VW_7 -->
<!-- df_pred1 <- -->
<!--   predictions(int_homo, condition = c("homclass3_V_res","dclass3res_V"), -->
<!--               newdata = datagrid(homclass3_V_res = seq(min(dfreg$homclass3_V_res), max(dfreg$homclass3_V_res), by= 1), -->
<!--                                  dclass3res_V = levels(dfreg$dclass3res_V))) -->
<!-- # names(df_pred1) -->
<!-- df_pred1 %>% -->
<!-- ggplot(aes(y =estimate,x=homclass3_V_res, fill=dclass3res_V, color=dclass3res_V,ymin=conf.low, ymax=conf.high)) + -->
<!--   geom_ribbon(alpha=0.8,size=1,linetype="solid",color="black") + -->
<!--   geom_line(size=1,color="black") + -->
<!--   geom_point(size=2,color="black") + -->
<!--   facet_grid(~factor(dclass3res_V, levels= -->
<!--                        c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"), -->
<!--                      c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"))) + -->
<!--   scale_fill_manual(values=c("#323232","#989898","#CCCCCC"))+ -->
<!--   scale_color_manual(values=c("#323232","#989898","#CCCCCC"))+ -->
<!--   scale_y_continuous(limits = c(50,100))+ -->
<!--   # scale_x_continuous(limits = c(-9,9))+ -->
<!--   labs(caption = paste0("N=",nobs(int_homo),", Countries=",lme4::ngrps(int_homo)), -->
<!--        x="Social Distance",y="Redistributive preferences") + -->
<!--   theme(legend.position = "none", legend.title = element_blank()) -->
<!-- ``` -->

<!-- \pagebreak -->

<!-- ```{r tab2-socdistclass, echo=FALSE, cache=TRUE, results='asis'} -->
<!-- set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1") -->
<!-- options(ggplot2.discrete.fill = set1) -->
<!-- options(ggplot2.discrete.colour = set1) -->
<!-- ggplot2::theme_set(ggplot2::theme_bw()) -->
<!-- ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif")) -->
<!-- options(scipen=999) -->
<!-- if (!require("pacman")) install.packages("pacman") -->
<!-- pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4) -->
<!-- rm(list=ls()) -->

<!-- load(here::here("input/data/proc/study1_country.RData"));df1 <- df2 -->

<!-- df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)")) -->

<!-- dfreg <- df1 %>% dplyr::select( -->
<!--   egal = egal2, -->
<!--   # starts_with("class3"), -->
<!--   # starts_with("digclass3"), -->
<!--   # starts_with("dclass3"), -->
<!--   # starts_with("homclass"), -->
<!--   dclass3res_V, -->
<!--   dclass3spo_V, -->
<!--   homclass3_V_res=socdist, -->
<!--   know_total, -->
<!--   socialtrust, -->
<!--   Q03pcm, -->
<!--   edyears, -->
<!--   female, -->
<!--   agenum, -->
<!--   religion, -->
<!--   partner, -->
<!--   union, -->
<!--   workst, -->
<!--   WEIGHT, -->
<!--   region, -->
<!--   "gini_disp"=wid_gini_disp, -->
<!--   "gini_mkt", -->
<!--   gv_spen, -->
<!--   rel_red, -->
<!--   abs_red, -->
<!--   ilo_taxrev, -->
<!--   ilo_govspe, -->
<!--   d10d1=wid_p90p10, -->
<!--   wid_p90p50, -->
<!--   top10=wid_sharetop10, -->
<!--   rgdpna, -->
<!--   gdppercapita, -->
<!--   oecd, -->
<!--   country2, country, country3, -->
<!-- ) %>% -->
<!--   filter(oecd == "OECD") %>% -->
<!--   filter(country2 != "SVN") %>% -->
<!--   na.omit() %>% -->
<!--     mutate(loggdppercapita=scale(gdppercapita/1000), -->
<!--          edyears2 = edyears ^ 2, -->
<!--          age2 = agenum ^ 2, -->
<!--          gini_disp=scale(gini_disp), -->
<!--          homclass3_V_res=homclass3_V_res/10) -->

<!-- dfreg$wstate2 <- scale(scale(dfreg$abs_red*100) + scale(dfreg$ilo_taxrev) + scale(dfreg$ilo_govspe)) -->
<!-- # Step 1: Calculate absolute values -->
<!-- absolute_values <- abs(dfreg$homclass3_V_res) -->
<!-- # hist(absolute_values);summary(absolute_values) -->

<!-- # Step 2: Add 1 to all values to eliminate zeros -->
<!-- modified_values <- absolute_values + 0.1 -->
<!-- # hist(modified_values);summary(modified_values) -->

<!-- # Step 3: Reverse the order of the scale -->
<!-- reversed_values <- max(modified_values) + 0.1 - modified_values -->
<!-- dfreg$homclass3_V_res <- reversed_values -->

<!-- # Variable Group-Centering ------------------------------------------------ -->
<!-- dfreg <- -->
<!--   dfreg %>% -->
<!--   mutate(to_dummy(female), -->
<!--          to_dummy(union), -->
<!--          to_dummy(workst), -->
<!--          # dummy for categorical variables------------------------------------- -->
<!--          to_dummy(Q03pcm), -->
<!--          to_dummy(dclass3spo_V) -->
<!--   ) -->
<!-- dfreg$female_gc = group_center(dfreg$female_2, grp = dfreg$country2) -->
<!-- dfreg$union_gc = group_center(dfreg$union_2, grp = dfreg$country2) -->
<!-- dfreg$workst_gc = group_center(dfreg$workst_2, grp = dfreg$country2) -->
<!-- dfreg$edyears_gc = group_center(dfreg$edyears, grp = dfreg$country2) -->
<!-- dfreg$agenum_gc = group_center(dfreg$agenum, grp = dfreg$country2) -->
<!-- dfreg$age2_gc = group_center(dfreg$age2, grp = dfreg$country2) -->
<!-- dfreg$socialtrust_gc = group_center(dfreg$socialtrust, grp = dfreg$country2) -->
<!-- dfreg$homclass3_V_gc <-group_center(dfreg$homclass3_V_res, grp = dfreg$country2) -->
<!-- dfreg$know_total_gc = group_center(dfreg$know_total, grp = dfreg$country2) -->
<!-- dfreg$religion_gc = group_center(as.numeric(dfreg$religion), grp = dfreg$country2) -->
<!-- dfreg$partner_gc = group_center(as.numeric(dfreg$partner), grp = dfreg$country2) -->

<!-- # Household Income Tercile -->
<!-- dfreg$Q03pcm_1_gc = group_center(dfreg$Q03pcm_1, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_2_gc = group_center(dfreg$Q03pcm_2, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_3_gc = group_center(dfreg$Q03pcm_3, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_NA_gc = group_center(dfreg$Q03pcm_4, grp = dfreg$country2) -->

<!-- # Partner Social Class -->
<!-- dfreg$dclass3spo_V_1_gc = group_center(dfreg$dclass3spo_V_1, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_2_gc = group_center(dfreg$dclass3spo_V_2, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_3_gc = group_center(dfreg$dclass3spo_V_3, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_NA_gc = group_center(dfreg$dclass3spo_V_4, grp = dfreg$country2) -->

<!-- # drop dummies -->
<!-- dfreg <- dplyr::select(dfreg,-c(female_1,female_2,union_1,union_2,workst_1, -->
<!--                                 workst_2,Q03pcm_1,Q03pcm_2,Q03pcm_2,Q03pcm_4, -->
<!--                                 dclass3spo_V_1,dclass3spo_V_2,dclass3spo_V_3,dclass3spo_V_4)) -->

<!-- # dfreg$gini_disp <- scale(dfreg$gini_disp) -->

<!-- ## Model -  EGP3 - DIGCLASS - V to Working ----------------------------- -->
<!-- homo_V_gini <- -->
<!--   lmer(egal~ -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp <- -->
<!--   lmer(egal~ -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp_wstate2 <- -->
<!--   lmer(egal~ -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2+ -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp_controls <- -->
<!--   lmer(egal~ -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2 + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_full2 <- -->
<!--   lmer(egal~ -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2+ -->
<!--          homclass3_V_gc* -->
<!--          dclass3res_V* -->
<!--          gini_disp + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- digclas1_macro <- list(homo_V_gini, -->
<!--                        homo_V_gini_gdp, -->
<!--                        homo_V_gini_gdp_wstate2, -->
<!--                        homo_V_gini_full2) -->

<!-- ccoef <- list(homclass3_V_gc="Social distance (CWC)", -->
<!--               "dclass3res_VIntermediate class (III+IV)"="Intermediate Class", -->
<!--               "dclass3res_VWorking Class (V+VI+VII)"="Working Class", -->
<!--               "gini_disp" = "Income inequality (Gini index)", -->
<!--               loggdppercapita = "GDP/capita", -->
<!--               wstate2="Size of the welfare state", -->
<!--               "homclass3_V_gc:dclass3res_VIntermediate class (III+IV)" = "Distance*Intermediate Class", -->
<!--               "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII)" = "Distance*Working Class", -->
<!--               "homclass3_V_gc:gini_disp"= "Distance*Income Inequality", -->
<!--               "homclass3_V_gc:dclass3res_VIntermediate class (III+IV):gini_disp" = "Distance*Intermediate Class*Income Inequality", -->
<!--               "homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII):gini_disp" = "Distance*Working Class*Income Inequality") -->

<!-- texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.60, -->
<!--        caption = "Multilevel models for Income Inequality, Network segregation and Redistributive Preferences", -->
<!--        custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars", -->
<!--        caption.above = T,leading.zero = T,single.row = T, -->
<!--        custom.coef.map = ccoef, -->
<!--       groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Distance x Social Class"=7:8, "Distance * Social Class x Income Inequality"=10:11), -->
<!--        float.pos = "h!", -->
<!--       custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE, -->
<!--       custom.gof.names = c(NA,NA,"Num. groups", -->
<!--                            "Var: Group (Intercept)", -->
<!--                            "Var: Group Homogeneity", -->
<!--                            "Var: Group Intermediate Class ", -->
<!--                            "Var: Group Working Class", -->
<!--                            "Cov: Group (Intercept), Homogeneity", -->
<!--                            "Cov: Group (Intercept), Intermediate Class ", -->
<!--                            "Cov: Group (Intercept), Working Class", -->
<!--                            "Cov: Group Homogeneity, Intermediate Class", -->
<!--                            "Cov: Group Homogeneity, Working Class", -->
<!--                            "Cov: Group Intermediate Class, Working Class", -->
<!--                            "Var: Residual"), -->
<!--        use.packages = F) -->
<!-- ``` -->

<!-- ```{r fig5-socdistclass, echo=FALSE, fig.cap='Three-way interaccion effects for Redistributive Preferences, Social Distance, Social Class and Income Inequality', fig.dim=c(12,7)} -->
<!-- ### Predicted values homo*class*GiniD --------------------------------------- -->
<!-- # ginid_max<- round(max(dfreg$gini_disp) ,2) -->
<!-- # ginid_min<- round(min(dfreg$gini_disp),2) -->

<!-- ginid_max<- round(mean(dfreg$gini_disp) + sd(dfreg$gini_disp)  ,2) -->
<!-- ginid_min<- round(mean(dfreg$gini_disp) -sd(dfreg$gini_disp) ,2) -->
<!-- ginid_mea<- round(mean(dfreg$gini_disp),2) -->

<!-- df_pred_giniD_gc <- -->
<!--   predictions( -->
<!--     homo_V_gini_full2,newdata = datagrid( -->
<!--       gini_disp = c(ginid_min, ginid_mea, ginid_max), -->
<!--       wstate2=mean(dfreg$wstate2), -->
<!--       loggdppercapita=mean(dfreg$loggdppercapita), -->
<!--       homclass3_V_gc = seq(min(dfreg$homclass3_V_gc)+0.01, max(dfreg$homclass3_V_gc)+0.01, by=0.5), -->
<!--       dclass3res_V=levels(dfreg$dclass3res_V) -->
<!--     ) -->
<!--   ) -->

<!-- df_pred_giniD_gc <- as.data.frame(df_pred_giniD_gc) -->
<!-- df_pred_giniD_gc$gini_disp <- factor(df_pred_giniD_gc$gini_disp,labels = c("Gini (-1SD)","Gini (Mean)", "Gini (+1SD)")) -->
<!-- df_pred_giniD_gc <- df_pred_giniD_gc %>% filter(dclass3res_V!="Intermediate class (III+IV)") -->

<!-- df_pred_giniD_gc %>% -->
<!--   ggplot(aes(y=estimate,x=homclass3_V_gc, fill=dclass3res_V,color=dclass3res_V,ymin=conf.low, ymax=conf.high)) + -->
<!--   geom_ribbon(alpha=0.8,size=0.7,linetype="solid",color="black") + -->
<!--   geom_line(size=1,color="black") + -->
<!--   geom_point(size=3, shape = 21,color="black") + -->
<!--   facet_wrap(~gini_disp) + -->
<!--   labs(y = "Preferences for Redistribution", -->
<!--        x="Social Distance (CWC)", -->
<!--        caption = paste0("N=",nobs(homo_V_gini_full2),", Countries=",lme4::ngrps(homo_V_gini_full2)) -->
<!--        ) + -->
<!--   scale_x_continuous(sec.axis = sec_axis(~ . , name = "Income inequality (post-tax) \n ", breaks = NULL, labels = NULL), -->
<!--                      ) + -->

<!--    scale_fill_manual(values=c("#323232","#CCCCCC"))+ -->
<!--   scale_color_manual(values=c("#323232","#CCCCCC"))+ -->
<!--   theme(legend.position = "bottom", -->
<!--         legend.direction = "horizontal", -->
<!--         axis.text=element_text(size=15), -->
<!--         legend.title = element_blank(), -->
<!--         legend.background = element_blank(), -->
<!--         legend.box.background = element_rect(colour = "black")) -->
<!-- ``` -->

<!-- \pagebreak -->

<!-- ### Social Distance by ISEI  -->

<!-- ```{r table1-socdist, results='asis'} -->
<!-- set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1") -->
<!-- options(ggplot2.discrete.fill = set1) -->
<!-- options(ggplot2.discrete.colour = set1) -->
<!-- ggplot2::theme_set(ggplot2::theme_bw()) -->
<!-- ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif")) -->

<!-- if (!require("pacman")) install.packages("pacman") -->
<!-- pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)   -->
<!-- load(here::here("input/data/proc/study1_country.RData"));df1 <- df2 -->
<!-- df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)")) -->

<!-- dummy_df <- df1 %>% dplyr::select(id, -->
<!--   egal = egal2,dclass3res_V,dclass3spo_V,homclass3_V_res,know_total,socialtrust, -->
<!--   Q03pcm,edyears,female,agenum,religion,partner,union,workst,WEIGHT,region, -->
<!--   "gini_disp"=wid_gini_disp,"gini_mkt",gv_spen,rel_red,abs_red,ilo_taxrev, -->
<!--   ilo_govspe,d10d1=wid_p90p10,wid_p90p50,top10=wid_sharetop10,rgdpna,gdppercapita, -->
<!--   oecd,country2, country, country3) %>% -->
<!--   filter(country2 != "SVN") %>%  -->
<!--   filter(oecd == "OECD") %>% -->
<!--   na.omit() %>%  -->
<!--   dplyr::select(id) -->

<!-- df1$isei08sp <- car::recode(df1$isei08sp,"NA=0") -->
<!-- df1$isei08 <- car::recode(df1$isei08,"NA=0") -->

<!-- dfreg <- df1 %>% dplyr::select(id, -->
<!--   egal = egal2, -->
<!--   dclass3res_V=isei08, -->
<!--   dclass3spo_V=isei08sp, -->
<!--   homclass3_V_res=socdist, -->
<!--   know_total, -->
<!--   socialtrust, -->
<!--   Q03pcm, -->
<!--   edyears, -->
<!--   female, -->
<!--   agenum, -->
<!--   partner, -->
<!--   religion, -->
<!--   union, -->
<!--   workst, -->
<!--   WEIGHT, -->
<!--   region, -->
<!--   "gini_disp"=wid_gini_disp, -->
<!--   "gini_mkt", -->
<!--   gv_spen, -->
<!--   rel_red, -->
<!--   d10d1=wid_p90p10, -->
<!--   wid_p90p50, -->
<!--   top10=wid_sharetop10, -->
<!--   rgdpna, -->
<!--   gdppercapita, -->
<!--   oecd, -->
<!--   country2, country, country3, -->
<!-- ) %>% -->
<!--   mutate(logrgdpna = log(rgdpna), -->
<!--          loggdppercapita=log(gdppercapita), -->
<!--          edyears2 = edyears ^ 2, -->
<!--          age2 = agenum ^ 2, -->
<!--          dclass3res_V=dclass3res_V/10, -->
<!--          dclass3spo_V=dclass3spo_V/10, -->
<!--          homclass3_V_res=homclass3_V_res/10) %>% -->
<!--   filter(oecd == "OECD") %>% -->
<!--   filter(country2 != "SVN")  %>%  -->
<!--   filter(id %in% dummy_df$id) -->
<!-- dfreg <- na.omit(dfreg) -->
<!-- # Step 1: Calculate absolute values -->
<!-- absolute_values <- abs(dfreg$homclass3_V_res) -->
<!-- # hist(absolute_values);summary(absolute_values) -->

<!-- # Step 2: Add 1 to all values to eliminate zeros -->
<!-- modified_values <- absolute_values + 0.1 -->
<!-- # hist(modified_values);summary(modified_values) -->

<!-- # Step 3: Reverse the order of the scale -->
<!-- reversed_values <- max(modified_values) + 0.1 - modified_values -->

<!-- dfreg$homclass3_V_res <- reversed_values -->

<!-- ## Model -  EGP3 - DIGCLASS - V to Working ----------------------------- -->
<!-- digclas1_VW_1<- lmer(egal~1 + homclass3_V_res + female+agenum+age2 + partner + religion + (1|country2),data=dfreg,weights = WEIGHT) -->
<!-- digclas1_VW_2<- update(digclas1_VW_1, . ~ . +know_total+socialtrust) -->
<!-- digclas1_VW_3 <- update(digclas1_VW_2, . ~ . + dclass3res_V) -->

<!-- digclas1_VW_4 <- update(digclas1_VW_3, . ~ . + dclass3res_V+female+agenum+age2) -->
<!-- digclas1_VW_5 <- update(digclas1_VW_3, . ~ . + dclass3res_V+female+agenum+age2+edyears+ Q03pcm+workst+union) -->
<!-- digclas1_VW_6 <- update(digclas1_VW_3, . ~ . + dclass3res_V+female+agenum+age2+edyears+ Q03pcm+workst+union+dclass3spo_V) -->
<!-- digclas1_VW_7 <- update(digclas1_VW_6, . ~ . +dclass3res_V*homclass3_V_res) -->
<!-- digclas1_VW <- list(digclas1_VW_1,digclas1_VW_2, -->
<!--                     digclas1_VW_3, -->
<!--                     # digclas1_VW_4, -->
<!--                     digclas1_VW_5,digclas1_VW_6,digclas1_VW_7) -->

<!-- ccoef <- list(homclass3_V_res="Social Distance /10", -->
<!--               know_total="Network size", -->
<!--               socialtrust="Social Trust", -->
<!--               # "dclass3res_VIntermediate class (III+IV)"="Intermediate Class", -->
<!--               # "dclass3res_VWorking Class (V+VI+VII)"="Working Class", -->
<!--               # "femaleFemale"="Female (Ref. = Male)", -->
<!--               # "agenum"="Age", -->
<!--               # "age2"="Age2", -->
<!--               "dclass3res_V"="ISEI/10", -->
<!--               "edyears"="Year of Education", -->
<!--               "Q03pcmT02"="Income (T2)","Q03pcmT03"="Income (T3)","Q03pcmMissing"="Income (No information)", -->
<!--               "workstNot in paid work" = "Not in paid work (Ref. = In paid work)", -->
<!--               "unionYes" ="Union Membership (Ref. = Not Unionized)", -->
<!--               "dclass3spo_V" = "ISEI/10 (Partner)", -->
<!--               "homclass3_V_res:dclass3res_V" = "Social Distance*ISEI") -->

<!-- texreg(digclas1_VW,stars = c(0.001, 0.01, 0.05, 0.1), -->
<!--        caption = paste("(\\#tab:table2)","Multilevel models for Social Distance and Redistributive Preferences"), -->
<!--        custom.note = "Note: Models include sampling weights. Gender, age, marital status and religion are included as controls. Standard errors in parentheses. %stars", -->
<!--        caption.above = T,leading.zero = T, -->
<!--        custom.coef.map = ccoef, -->
<!--        symbol = "+",scalebox = 0.60, -->
<!--       custom.gof.rows = list("Controls"=rep("Yes",6)),include.loglik = FALSE,include.aic = FALSE, -->
<!--       custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual"), -->
<!--        float.pos = "h!", -->
<!--        use.packages = F) -->
<!-- ``` -->

<!-- ```{r fig1-marg-socdist, fig.dim=c(10,5),fig.cap='Average Marginal Effects of Social Distance conditionated by ISEI on Redistributive Preferences',warning=FALSE} -->
<!-- int_homo <- digclas1_VW_7 -->
<!-- plot_slopes(digclas1_VW_7, variables = "homclass3_V_res", condition = list("dclass3res_V" = "threenum"))+ -->
<!-- labs(caption = paste0("N=",nobs(int_homo),", Countries=",lme4::ngrps(int_homo)), -->
<!--        x="ISEI",y="Redistributive preferences") -->
<!-- ``` -->

<!-- ```{r fig1-pred-socdist, fig.dim=c(10,5),fig.cap='Linear Predictions for Social Distance on Redistributive Preferences by ISEI',warning=FALSE} -->
<!-- int_homo <- digclas1_VW_7 -->
<!-- df_pred1 <-  -->
<!--   predictions(int_homo, condition = c("homclass3_V_res","dclass3res_V"), -->
<!--               newdata = datagrid(homclass3_V_res = seq(min(dfreg$homclass3_V_res), max(dfreg$homclass3_V_res), by= 0.5), -->
<!--                                  dclass3res_V = c(min(dfreg$dclass3res_V),mean(dfreg$dclass3res_V),max(dfreg$dclass3res_V)))) -->

<!-- df_pred1$dclass3res_V <- factor(df_pred1$dclass3res_V,labels = c("ISEI (min)","ISEI (Mean)", "ISEI (max)")) -->

<!-- # names(df_pred1) -->
<!-- df_pred1 %>%  -->
<!-- ggplot(aes(y =estimate,x=homclass3_V_res, fill=dclass3res_V, color=dclass3res_V,ymin=conf.low, ymax=conf.high)) + -->
<!--   geom_ribbon(alpha=0.8,size=1,linetype="solid",color="black") + -->
<!--   geom_line(size=1,color="black") + -->
<!--   geom_point(size=2,color="black") + -->
<!--   # facet_grid(~dclass3res_V) + -->
<!--   scale_fill_manual(values=c("#323232","#989898","#CCCCCC"))+ -->
<!--   scale_color_manual(values=c("#323232","#989898","#CCCCCC"))+ -->
<!--   scale_y_continuous(limits = c(40,100))+ -->
<!--   labs(caption = paste0("N=",nobs(int_homo),", Countries=",lme4::ngrps(int_homo)), -->
<!--        x="Social distance",y="Redistributive preferences") + -->
<!--   theme(legend.position = "bottom", legend.title = element_blank()) -->
<!-- ``` -->

<!-- \pagebreak -->

<!-- ```{r tab2-socdist, results='asis', cache=TRUE} -->
<!-- set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1") -->
<!-- options(ggplot2.discrete.fill = set1) -->
<!-- options(ggplot2.discrete.colour = set1) -->
<!-- ggplot2::theme_set(ggplot2::theme_bw()) -->
<!-- ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif")) -->
<!-- options(scipen=999) -->
<!-- if (!require("pacman")) install.packages("pacman") -->
<!-- pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)   -->
<!-- rm(list=ls()) -->

<!-- load(here::here("input/data/proc/study1_country.RData"));df1 <- df2 -->

<!-- df1$dclass3spo_V <- car::recode(df1$dclass3spo_V,"NA='No information (Missing, No partner)'",levels = c("Service Class (I+II)","Intermediate class (III+IV)","Working Class (V+VI+VII)","No information (Missing, No partner)")) -->
<!-- dummy_df <- df1 %>% dplyr::select(id, -->
<!--   egal = egal2,dclass3res_V,dclass3spo_V,homclass3_V_res,know_total,socialtrust, -->
<!--   Q03pcm,edyears,female,agenum,religion,partner,union,workst,WEIGHT,region, -->
<!--   "gini_disp"=wid_gini_disp,"gini_mkt",gv_spen,rel_red,abs_red,ilo_taxrev, -->
<!--   ilo_govspe,d10d1=wid_p90p10,wid_p90p50,top10=wid_sharetop10,rgdpna,gdppercapita, -->
<!--   oecd,country2, country, country3) %>% -->
<!--   filter(country2 != "SVN") %>%  -->
<!--   filter(oecd == "OECD") %>% -->
<!--   na.omit() %>%  -->
<!--   dplyr::select(id) -->

<!-- df1$isei08sp <- car::recode(df1$isei08sp,"NA=0") -->
<!-- dfreg <- df1 %>% dplyr::select(id, -->
<!--   egal = egal2, -->
<!--   # starts_with("class3"), -->
<!--   # starts_with("digclass3"), -->
<!--   # starts_with("dclass3"), -->
<!--   # starts_with("homclass"), -->
<!--   dclass3res_V=isei08, -->
<!--   dclass3spo_V=isei08sp, -->
<!--   homclass3_V_res=socdist, -->
<!--   know_total, -->
<!--   socialtrust, -->
<!--   Q03pcm, -->
<!--   edyears, -->
<!--   female, -->
<!--   agenum, -->
<!--   religion, -->
<!--   partner, -->
<!--   union, -->
<!--   workst, -->
<!--   WEIGHT, -->
<!--   region, -->
<!--   "gini_disp"=wid_gini_disp, -->
<!--   "gini_mkt", -->
<!--   gv_spen, -->
<!--   rel_red, -->
<!--   abs_red, -->
<!--   ilo_taxrev, -->
<!--   ilo_govspe, -->
<!--   d10d1=wid_p90p10, -->
<!--   wid_p90p50, -->
<!--   top10=wid_sharetop10, -->
<!--   rgdpna, -->
<!--   gdppercapita, -->
<!--   oecd, -->
<!--   country2, country, country3, -->
<!-- ) %>% -->
<!--   filter(oecd == "OECD") %>% -->
<!--   filter(country2 != "SVN") %>%  -->
<!--   na.omit() %>%  -->
<!--     mutate(loggdppercapita=scale(gdppercapita/1000), -->
<!--          edyears2 = edyears ^ 2, -->
<!--          age2 = agenum ^ 2,  -->
<!--          gini_disp=scale(gini_disp), -->
<!--          dclass3res_V=dclass3res_V/10, -->
<!--          dclass3spo_V=dclass3spo_V/10, -->
<!--          homclass3_V_res=homclass3_V_res/10) %>%  -->
<!--   filter(id %in% dummy_df$id) -->

<!-- dfreg$wstate2 <- scale(scale(dfreg$abs_red*100) + scale(dfreg$ilo_taxrev) + scale(dfreg$ilo_govspe)) -->
<!-- # Step 1: Calculate absolute values -->
<!-- absolute_values <- abs(dfreg$homclass3_V_res) -->
<!-- # hist(absolute_values);summary(absolute_values) -->

<!-- # Step 2: Add 1 to all values to eliminate zeros -->
<!-- modified_values <- absolute_values + 0.1 -->
<!-- # hist(modified_values);summary(modified_values) -->

<!-- # Step 3: Reverse the order of the scale -->
<!-- reversed_values <- max(modified_values) + 0.1 - modified_values -->
<!-- dfreg$homclass3_V_res <- reversed_values -->

<!-- # Variable Group-Centering ------------------------------------------------ -->
<!-- dfreg <-  -->
<!--   dfreg %>%  -->
<!--   mutate(to_dummy(female), -->
<!--          to_dummy(union), -->
<!--          to_dummy(workst), -->
<!--          # dummy for categorical variables------------------------------------- -->
<!--          to_dummy(Q03pcm), -->
<!--          # to_dummy(dclass3spo_V) -->
<!--   ) -->
<!-- dfreg$female_gc = group_center(dfreg$female_2, grp = dfreg$country2) -->
<!-- dfreg$union_gc = group_center(dfreg$union_2, grp = dfreg$country2) -->
<!-- dfreg$workst_gc = group_center(dfreg$workst_2, grp = dfreg$country2) -->
<!-- dfreg$edyears_gc = group_center(dfreg$edyears, grp = dfreg$country2) -->
<!-- dfreg$agenum_gc = group_center(dfreg$agenum, grp = dfreg$country2) -->
<!-- dfreg$age2_gc = group_center(dfreg$age2, grp = dfreg$country2) -->
<!-- dfreg$socialtrust_gc = group_center(dfreg$socialtrust, grp = dfreg$country2) -->
<!-- dfreg$homclass3_V_gc <-group_center(dfreg$homclass3_V_res, grp = dfreg$country2) -->
<!-- dfreg$know_total_gc = group_center(dfreg$know_total, grp = dfreg$country2) -->
<!-- dfreg$religion_gc = group_center(as.numeric(dfreg$religion), grp = dfreg$country2) -->
<!-- dfreg$partner_gc = group_center(as.numeric(dfreg$partner), grp = dfreg$country2) -->

<!-- dfreg$dclass3res_V = group_center(as.numeric(dfreg$dclass3res_V), grp = dfreg$country2) -->

<!-- # Household Income Tercile -->
<!-- dfreg$Q03pcm_1_gc = group_center(dfreg$Q03pcm_1, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_2_gc = group_center(dfreg$Q03pcm_2, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_3_gc = group_center(dfreg$Q03pcm_3, grp = dfreg$country2) -->
<!-- dfreg$Q03pcm_NA_gc = group_center(dfreg$Q03pcm_4, grp = dfreg$country2) -->

<!-- # Partner Social Class -->
<!-- # dfreg$dclass3spo_V_1_gc = group_center(dfreg$dclass3spo_V_1, grp = dfreg$country2) -->
<!-- # dfreg$dclass3spo_V_2_gc = group_center(dfreg$dclass3spo_V_2, grp = dfreg$country2) -->
<!-- # dfreg$dclass3spo_V_3_gc = group_center(dfreg$dclass3spo_V_3, grp = dfreg$country2) -->
<!-- # dfreg$dclass3spo_V_NA_gc = group_center(dfreg$dclass3spo_V_4, grp = dfreg$country2) -->
<!-- dfreg$dclass3spo_V_gc = group_center(dfreg$dclass3spo_V, grp = dfreg$country2) -->

<!-- # drop dummies -->
<!-- dfreg <- dplyr::select(dfreg,-c(female_1,female_2,union_1,union_2,workst_1, -->
<!--                                 workst_2,Q03pcm_1,Q03pcm_2,Q03pcm_2,Q03pcm_4, -->
<!--                                 # dclass3spo_V_1,dclass3spo_V_2,dclass3spo_V_3,dclass3spo_V_4 -->
<!--                                 )) -->

<!-- # dfreg$gini_disp <- scale(dfreg$gini_disp) -->

<!-- ## Model -  EGP3 - DIGCLASS - V to Working ----------------------------- -->
<!-- homo_V_gini <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+ -->
<!--          # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          dclass3spo_V_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+ -->
<!--          # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          dclass3spo_V_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp_wstate2 <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+ -->
<!--          # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          dclass3spo_V_gc+         female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2+ -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_gdp_controls <- -->
<!--   lmer(egal~ -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+ -->
<!--          # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          dclass3spo_V_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2 + -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- homo_V_gini_full2 <-  -->
<!--   lmer(egal~  -->
<!--          homclass3_V_gc+ -->
<!--          know_total_gc+ -->
<!--          socialtrust_gc+ -->
<!--          dclass3res_V+ -->
<!--          # dclass3spo_V_2_gc+dclass3spo_V_3_gc+dclass3spo_V_NA_gc+ -->
<!--          dclass3spo_V_gc+ -->
<!--          female_gc+agenum_gc+age2_gc+partner_gc+religion_gc+ -->
<!--          edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+ -->
<!--          union_gc+workst_gc+ -->
<!--          gini_disp +loggdppercapita + wstate2+ -->
<!--          homclass3_V_gc* -->
<!--          dclass3res_V* -->
<!--          gini_disp +  -->
<!--          (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT) -->

<!-- digclas1_macro <- list(homo_V_gini, -->
<!--                        homo_V_gini_gdp, -->
<!--                        homo_V_gini_gdp_wstate2, -->
<!--                        homo_V_gini_full2) -->

<!-- ccoef <- list(homclass3_V_gc="Social Distance (CWC)", -->
<!--               "dclass3res_V"="ISEI ", -->
<!--               "gini_disp" = "Income inequality (Gini index)", -->
<!--               loggdppercapita = "GDP/capita", -->
<!--               wstate2="Size of the welfare state", -->
<!--               "homclass3_V_gc:dclass3res_V" = "Distance*ISEI", -->
<!--               "homclass3_V_gc:gini_disp"= "Distance*Income Inequality", -->
<!--               "dclass3res_V:gini_disp" ="ISEI*Inequality", -->
<!--               "homclass3_V_gc:dclass3res_V:gini_disp" = "Distance*Working Class*Income Inequality") -->

<!-- texreg(digclas1_macro,stars = c(0.001, 0.01, 0.05, 0.1),symbol = "+",scalebox = 0.60, -->
<!--        caption = "Multilevel models for Income Inequality, Network segregation and Redistributive Preferences", -->
<!--        custom.note = "Note: Models include sampling weights and individual level controls centered within cluster (group mean). Standard errors in parentheses. %stars", -->
<!--        caption.above = T,leading.zero = T,single.row = T, -->
<!--        custom.coef.map = ccoef, -->
<!--       # groups = list("Social Class (Ref.= Service Class)" = 2:3,"Macro-level factors"=4:6, "Distance*Social Class"=7:8, "Distance * Social Class * Income Inequality"=10:11), -->
<!--        float.pos = "h!", -->
<!--       custom.gof.rows = list("Controls"=rep("Yes",4)),include.loglik = FALSE,include.aic = FALSE, -->
<!--       custom.gof.names = c(NA,NA,"Num. groups", -->
<!--                            "Var: Group (Intercept)", -->
<!--                            "Var: Group Soc. Distance", -->
<!--                            "Var: Group ISEI", -->
<!--                            "Cov: Group (Intercept), Soc. Distance", -->
<!--                            "Cov: Group (Intercept), ISEI",                            -->
<!--                            "Cov: Group Soc. Distance, ISEI", -->
<!--                            "Var: Residual"), -->
<!--        use.packages = F) -->
<!-- ``` -->

<!-- ```{r fig2-socdist, fig.dim=c(12,7),fig.cap='Three-way interaccion effects for Redistributive Preferences, Social Distance, ISEI and Income Inequality'} -->
<!-- ### Predicted values homo*class*GiniD --------------------------------------- -->
<!-- # ginid_max<- round(max(dfreg$gini_disp) ,2) -->
<!-- # ginid_min<- round(min(dfreg$gini_disp),2) -->

<!-- ginid_max<- round(mean(dfreg$gini_disp) + sd(dfreg$gini_disp)  ,2) -->
<!-- ginid_min<- round(mean(dfreg$gini_disp) -sd(dfreg$gini_disp) ,2) -->
<!-- ginid_mea<- round(mean(dfreg$gini_disp),2) -->

<!-- df_pred_giniD_gc <- -->
<!--   predictions( -->
<!--     homo_V_gini_full2,newdata = datagrid( -->
<!--       gini_disp = c(ginid_min, ginid_mea, ginid_max), -->
<!--       wstate2=mean(dfreg$wstate2), -->
<!--       loggdppercapita=mean(dfreg$loggdppercapita), -->
<!--       homclass3_V_gc = seq(min(dfreg$homclass3_V_gc)+0.01, max(dfreg$homclass3_V_gc)+0.01, by=0.5), -->
<!--       dclass3res_V=c(min(dfreg$dclass3res_V),max(dfreg$dclass3res_V)) -->
<!--     ) -->
<!--   ) -->

<!-- df_pred_giniD_gc <- as.data.frame(df_pred_giniD_gc) -->
<!-- df_pred_giniD_gc$gini_disp <- factor(df_pred_giniD_gc$gini_disp,labels = c("Gini (-1SD)","Gini (Mean)", "Gini (+1SD)")) -->

<!-- df_pred_giniD_gc$dclass3res_V <- factor(df_pred_giniD_gc$dclass3res_V,labels = c("ISEI (min)","ISEI (max)")) -->
<!-- # df_pred_giniD_gc <- df_pred_giniD_gc %>% filter(dclass3res_V!="Intermediate class (III+IV)") -->

<!-- df_pred_giniD_gc$dclass3res_V <- as.factor(df_pred_giniD_gc$dclass3res_V) -->

<!-- df_pred_giniD_gc %>%  -->
<!--   ggplot(aes(y=estimate,x=homclass3_V_gc, fill=dclass3res_V,color=dclass3res_V,ymin=conf.low, ymax=conf.high)) + -->
<!--   geom_ribbon(alpha=0.8,size=0.7,linetype="solid",color="black") + -->
<!--   geom_line(size=1,color="black") + -->
<!--   geom_point(size=3, shape = 21,color="black") + -->
<!--   facet_wrap(~gini_disp) + -->
<!--   labs(y = "Preferences for Redistribution", -->
<!--        x="Social Distance (CWC)", -->
<!--        caption = paste0("N=",nobs(homo_V_gini_full2),", Countries=",lme4::ngrps(homo_V_gini_full2)) -->
<!--        ) + -->
<!--   scale_x_continuous(sec.axis = sec_axis(~ . , name = "Income inequality (post-tax) \n ", breaks = NULL, labels = NULL), -->
<!--                      ) + -->

<!--    scale_fill_manual(values=c("#323232","#CCCCCC"))+ -->
<!--   scale_color_manual(values=c("#323232","#CCCCCC"))+ -->
<!--   theme(legend.position = "bottom", -->
<!--         legend.direction = "horizontal", -->
<!--         axis.text=element_text(size=15), -->
<!--         legend.title = element_blank(), -->
<!--         legend.background = element_blank(), -->
<!--         legend.box.background = element_rect(colour = "black")) -->
<!-- ``` -->