### 2025-05-13 procesamiento/05-prod_prep-complementary-analysis.Rmd
avance review

---
title: "Complementary analysis"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output: 
  html_document: 
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
      number_sections: yes
    code_folding: show  
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

# Setup

```{r setup}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE 
                      )
options(scipen=9999) # desactivar notacion cientifica
if (!require("pacman")) install.packages("pacman") # instalar pacman
pacman::p_load(dplyr,sjmisc,sjPlot,sjlabelled,tidyverse,haven,MLMusingR)
```

# Individual-level

```{r, results='asis'}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))

if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(
  prefineq,redist,
  egal = egal2,
  dclass3res_V,
  n_middle,n_low,n_high,
  homclass3_V_res,
  know_total,
  Q03pcm,
  edyears,
  female,
  agenum,
  partner,
  workst, 
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  mutate(logrgdpna = log(rgdpna),
         loggdppercapita=log(gdppercapita),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2) %>%
  filter(country2 != "SVN")

dfreg <- na.omit(dfreg)

## Model -  EGP3 - DIGCLASS - V to Working -----------------------------$

digclas1_VW_7 <- lmer(egal~1 + dclass3res_V+n_low+female+agenum+partner+workst+ (1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_8 <- lmer(egal~1 + dclass3res_V+n_low+n_middle+female+agenum+partner+workst+ (1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_10 <- lmer(egal~1 + dclass3res_V+n_low+n_middle+n_high+female+agenum+partner+workst+ (1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_10.1 <- lmer(egal~1 + dclass3res_V*n_low+n_middle+n_high+female+agenum+partner+workst+ (1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_10.2 <- lmer(egal~1 + dclass3res_V*n_middle+n_high+n_low+female+agenum+partner+workst+ (1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_10.3 <- lmer(egal~1 + dclass3res_V*n_high+n_low+n_middle+female+agenum+partner+workst+ (1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW_11 <- lmer(egal~1 + dclass3res_V*n_low + dclass3res_V*n_middle + dclass3res_V*n_high +female+agenum+partner+workst+ (1|country2),data=dfreg,weights = WEIGHT)

digclas1_VW <- list(digclas1_VW_7,digclas1_VW_8,digclas1_VW_10,
                    digclas1_VW_10.1,digclas1_VW_10.2,digclas1_VW_10.3,
                    digclas1_VW_11)

ccoef <- list(homclass3_V_res="Class-based network homogeneity",
              know_total="Network size",
              "dclass3res_VIntermediate class (III+IV)"="Intermediate Class",
              "dclass3res_VWorking Class (V+VI+VII)"="Working Class",
              "n_low" = "Tie to working class",
              "n_middle" = "Tie to Intermediate class",
              "n_high" = "Tie to Service class",
              "dclass3res_VIntermediate class (III+IV):n_low" = "Intermediate x Tie to Working class",
              "dclass3res_VWorking Class (V+VI+VII):n_low"    = "Working x Tie to Working class",
              "dclass3res_VIntermediate class (III+IV):n_middle" = "Intermediate x Tie to Intermediate class",
              "dclass3res_VWorking Class (V+VI+VII):n_middle" = "Working x Tie to Intermediate class",
              "dclass3res_VIntermediate class (III+IV):n_high" = "Intermediate x Tie to Service class",
              "dclass3res_VWorking Class (V+VI+VII):n_high"= "Working x Tie to Service class"
              )

screenreg(digclas1_VW,stars = c(0.001, 0.01, 0.05, 0.1),single.row = F,symbol = ".",
       caption = paste("Multilevel models for network homogeneity and redistributive preferences"),
       custom.note = "Note: Standard errors in parentheses. %stars",
       threeparttable = T,
       caption.above = T,leading.zero = T,
       custom.coef.map = ccoef,
      groups = list("Social Class (Ref.= Service Class)" = 1:2,
                    "Network class-profile" = 3:5,
                    "Interactions Class x Class-profile" = 6:11),
      # custom.gof.rows = list("Controls"=c("Yes",rep("Yes",1))),
      include.loglik = FALSE,
      include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual")
      )
```

# Country-level

```{r}
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
data_corruption <- xlsx::read.xlsx("input/data/original/coruption-data/CPI2017_Full_DataSet-1801.xlsx",sheetIndex = 1)
data_corruption$year <- 2017

data_unionden <- haven::read_dta("input/data/original/union-density/union-density.dta") 
data_migration <- data.table::fread("input/data/original/migration/migrant-stock-share.csv")
data_unemployment <- data.table::fread("input/data/original/unemployment-world-bank/unemployment-rate.csv")
data_qog <- haven::read_dta("input/data/original/QOG/qog_std_ts_jan25.dta") 
data_vdem <- haven::read_dta("input/data/original/QOG/V-Dem-CY-Full+Others-v15.dta")

union_den <- 
data_unionden %>% 
  dplyr::select(country=ref_area_label,year=time,union_den=obs_value) %>% 
  mutate(ccode=countrycode::countrycode(country,origin = "country.name",destination = "iso3n"),
         country=countrycode::countrycode(ccode,origin = "iso3n",destination = "country.name"),
         year=as.numeric(year))

migration<- 
data_migration %>% dplyr::select(country=Entity,year=Year,mig_share=`Share of the population that was born in another country`) %>% 
  mutate(ccode=countrycode::countrycode(country,origin = "country.name",destination = "iso3n"),
         country=countrycode::countrycode(ccode,origin = "iso3n",destination = "country.name"),
         year=as.numeric(year))

corruption <- 
data_corruption %>% 
  dplyr::select(country=Country,year,iso3=ISO3,cpi=CPI.Score.2017) %>% 
  mutate(ccode=countrycode::countrycode(country,origin = "country.name",destination = "iso3n"),
         country=countrycode::countrycode(ccode,origin = "iso3n",destination = "country.name"),
         year=as.numeric(year))

unemployment <- 
data_unemployment %>% 
  dplyr::select(country=Entity,year=Year,unemployment_rate=`Unemployment, total (% of total labor force) (modeled ILO estimate)`) %>% 
  mutate(ccode=countrycode::countrycode(country,origin = "country.name",destination = "iso3n"),
         country=countrycode::countrycode(ccode,origin = "iso3n",destination = "country.name"),
         year=as.numeric(year))

democracy <- 
data_qog %>% 
  dplyr::select(country=cname,year=year,van_index,p_polity2) %>% 
  mutate(ccode=countrycode::countrycode(country,origin = "country.name",destination = "iso3n"),
         country=countrycode::countrycode(ccode,origin = "iso3n",destination = "country.name"),
         year=as.numeric(year))

vdem <- 
data_vdem %>% 
  dplyr::select(country=country_name,year=year,v2x_delibdem,v2x_polyarchy,v2x_egal) %>% 
  mutate(ccode=countrycode::countrycode(country,origin = "country.name",destination = "iso3n"),
         country=countrycode::countrycode(ccode,origin = "iso3n",destination = "country.name"),
         year=as.numeric(year))

additional_macro <- 
union_den %>% 
  full_join(migration,by=c("ccode","country","year")) %>% 
  full_join(corruption,by=c("ccode","country","year")) %>% 
  full_join(unemployment,by=c("ccode","country","year")) %>%   
  full_join(democracy,by=c("ccode","country","year")) %>%
  full_join(vdem,by=c("ccode","country","year")) %>%  
  arrange(year)

summary(additional_macro)

country_code<- as.numeric(levels(as.factor(as.character(df1$ccode))))

additional_macro_2000to2018 <- 
  additional_macro %>% 
  filter(year %in% c(2015:2017)) %>% 
  filter(ccode %in% country_code)

additional_macro_avg <- 
additional_macro_2000to2018 %>% 
  group_by(ccode) %>% 
  summarise(union_den=mean(union_den,na.rm=T),
            mig_share=mean(mig_share,na.rm=T),
            cpi=mean(cpi,na.rm=T),
            unemployment_rate=mean(unemployment_rate,na.rm=T),
            van_index=mean(van_index,na.rm=T),
            p_polity2=mean(p_polity2,na.rm=T),            
            v2x_delibdem=mean(v2x_delibdem ,na.rm=T),
            v2x_polyarchy=mean(v2x_polyarchy ,na.rm=T),
            v2x_egal = mean(v2x_egal,na.rm=T)
            )
summary(additional_macro_avg)
save(additional_macro_avg,file ="input/data/proc/additional_macro_avg.RData")
```

```{r}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
# rm(list=ls())

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
load(file ="input/data/proc/additional_macro_avg.RData")

# Merge new macro-data
df1 <- df1 %>% left_join(.,additional_macro_avg,by ="ccode")
dfreg <- df1 %>% dplyr::select(
  redist,prefineq,
  egal = egal2,
  dclass3res_V,
  homclass3_V_res,
  know_total,
  Q03pcm,
  edyears,
  female,
  agenum,
  partner,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  cpi,
  mig_share,union_den,van_index,v2x_delibdem,v2x_polyarchy,v2x_egal,p_polity2,
  country2, country, country3,
) %>%
  filter(country2 != "SVN") %>% 
  # na.omit() %>%
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         gini_disp=scale(gini_disp))

dfreg$wstate2 <- scale(scale(dfreg$abs_red) + scale(dfreg$ilo_taxrev) + scale(dfreg$ilo_govspe))

# Variable Group-Centering ------------------------------------------------$
dfreg <- 
  dfreg %>% 
  mutate(to_dummy(female),
         to_dummy(workst),
         to_dummy(Q03pcm)
         )
dfreg$female_gc = group_center(dfreg$female_2, grp = dfreg$country2)
dfreg$workst_gc = group_center(dfreg$workst_2, grp = dfreg$country2)
dfreg$edyears_gc = group_center(dfreg$edyears, grp = dfreg$country2)
dfreg$agenum_gc = group_center(dfreg$agenum, grp = dfreg$country2)
dfreg$age2_gc = group_center(dfreg$age2, grp = dfreg$country2)
dfreg$homclass3_V_gc <-group_center(dfreg$homclass3_V_res, grp = dfreg$country2)
dfreg$know_total_gc = group_center(dfreg$know_total, grp = dfreg$country2)
dfreg$partner_gc = group_center(as.numeric(dfreg$partner), grp = dfreg$country2)

# Household Income Tercile
dfreg$Q03pcm_1_gc = group_center(dfreg$Q03pcm_1, grp = dfreg$country2)
dfreg$Q03pcm_2_gc = group_center(dfreg$Q03pcm_2, grp = dfreg$country2)
dfreg$Q03pcm_3_gc = group_center(dfreg$Q03pcm_3, grp = dfreg$country2)
dfreg$Q03pcm_NA_gc = group_center(dfreg$Q03pcm_4, grp = dfreg$country2)

# drop dummies
dfreg <- dplyr::select(dfreg,-c(female_1,female_2,
                                workst_2,Q03pcm_1,Q03pcm_2,Q03pcm_2,Q03pcm_4,
                                ))

## Model -  EGP3 - DIGCLASS - V to Working -----------------------------$
homo_V_gini <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         workst_gc+
         gini_disp +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

screenreg(homo_V_gini)
homo_V_gini_gdp <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         workst_gc+
         gini_disp +loggdppercapita +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

screenreg(homo_V_gini_gdp)

homo_V_gini_gdp_wstate2 <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2+
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

screenreg(homo_V_gini_gdp_wstate2)

homo_V_gini_gdp_cpi <-
  lmer(egal~
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2 + cpi +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

screenreg(homo_V_gini_gdp_cpi)

homo_V_gini_gdp_mig <-
  lmer(egal~
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2 + mig_share +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

screenreg(homo_V_gini_gdp_mig)

homo_V_gini_gdp_union <-
  lmer(egal~
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2 +union_den + 
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

screenreg(homo_V_gini_gdp_union)

homo_V_gini_gdp_van_index <-
  lmer(egal~
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2 + van_index +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)
screenreg(homo_V_gini_gdp_van_index)

homo_V_gini_gdp_v2x_delibdem <-
  lmer(egal~
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2 + v2x_delibdem +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)
screenreg(homo_V_gini_gdp_v2x_delibdem)
homo_V_gini_gdp_v2x_polyarchy <-
  lmer(egal~
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2 + v2x_polyarchy +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

homo_V_gini_gdp_v2x_egal <-
  lmer(egal~
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2 + v2x_egal +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

screenreg(homo_V_gini_gdp_v2x_egal)

homo_V_gini_gdp_p_polity2 <-
  lmer(egal~
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2 + p_polity2 +
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

screenreg(homo_V_gini_gdp_p_polity2)

homo_V_gini_full2 <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2+
         homclass3_V_gc*
         dclass3res_V*
         gini_disp + 
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

screenreg(homo_V_gini_full2)

homo_V_gini_full2.1 <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2+
         homclass3_V_gc*
         dclass3res_V*
         gini_disp + 
         homclass3_V_gc*
         dclass3res_V*
         wstate2 +          
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

screenreg(homo_V_gini_full2.1)

homo_V_gini_full3 <- 
  lmer(egal~ 
         homclass3_V_gc+
         know_total_gc+
         dclass3res_V+
         female_gc+agenum_gc+
         partner_gc+
         # edyears_gc+Q03pcm_2_gc+Q03pcm_3_gc+Q03pcm_NA_gc+
         workst_gc+
         gini_disp +loggdppercapita + wstate2+
         homclass3_V_gc*
         dclass3res_V*
         gini_disp + mig_share + union_den + p_polity2 + cpi+ 
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)
screenreg(homo_V_gini_full3)
```