### 2025-09-30 processing/99-prod_power-analysis.Rmd

---
title: "Power analysis"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output: 
  html_document: 
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
      number_sections: yes
    code_folding: show  
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

# Power analysis

```{r}
set1 <-  RColorBrewer::brewer.pal(n = 8, name = "Set1")
options(ggplot2.discrete.fill = set1)
options(ggplot2.discrete.colour = set1)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(text=ggplot2::element_text(size=15,  family="serif"))
options(scipen=999)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(simr,dplyr,sjmisc,ggplot2,interplot,marginaleffects,sjlabelled,haven,stringr,ordinal,texreg,MLMusingR,lme4)  
rm(list=ls())
# setwd("~/paper1")
# load("study1_country.RData");df1 <- df2
load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
# load("additional_macro_avg.RData")
load(here::here("input/data/proc/additional_macro_avg.RData"))

df1$idnum <- rownames(df1)

dfid <- df1 %>% dplyr::select(
  idnum,
  redist,prefineq,
  egal = egal2,
  dclass3res_V,
  homclass3_V_res,
  know_total,
  # Q03pcm,
  # edyears,
  female,
  agenum,
  # partner,
  # workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  country2, country, country3,
) %>% na.omit()

# Merge new macro-data
df1 <- df1 %>% left_join(.,additional_macro_avg,by ="ccode")
dfreg <- df1 %>% dplyr::select(
  idnum,
  redist,prefineq,
  egal = egal2,
  dclass3res_V,
  homclass3_V_res,
  know_total,
  # Q03pcm,
  # edyears,
  female,
  agenum,
  # partner,
  # workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  cpi,
  mig_share,
  union_den,
  v2x_polyarchy,
  unemployment_rate,
  v2catrauni,
  country2, country, country3,
) %>%
  filter(country2 != "SVN") %>% 
  # na.omit() %>%
  filter(idnum %in%  dfid$idnum) %>% 
  mutate(loggdppercapita=scale(gdppercapita/1000),
         gini_disp=scale(gini_disp))

fit_gov<- psych::pca(dfreg[,c("ilo_taxrev","ilo_govspe","abs_red")])
dfreg$wstate<- as.numeric(fit_gov$scores)
dfreg$wstate2 <- ((dfreg$wstate-min(dfreg$wstate,na.rm = T))/(max(dfreg$wstate,na.rm = T)-min(dfreg$wstate,na.rm = T)))*100
dfreg$wstate2 <- as.numeric(scale(dfreg$wstate2))
dfreg$gini_disp <- as.numeric(scale(dfreg$gini_disp))

# Variable Group-Centering ------------------------------------------------$
dfreg <- 
  dfreg %>% 
  mutate(to_dummy(female)
         )
dfreg$female_gc = group_center(dfreg$female_2, grp = dfreg$country2)
dfreg$agenum_gc = group_center(dfreg$agenum, grp = dfreg$country2)
dfreg$homclass3_V_gc <-group_center(dfreg$homclass3_V_res, grp = dfreg$country2)
dfreg$know_total_gc = group_center(dfreg$know_total, grp = dfreg$country2)
dfreg$egal_sd <- scale(dfreg$egal) #standardized DV

# Create Model with the double cross-level interactions: Gini + Welfare State Size------
model_re <- 
  lmer(egal_sd~homclass3_V_gc+know_total_gc+dclass3res_V+female_gc+agenum_gc+
         homclass3_V_gc*dclass3res_V*gini_disp+homclass3_V_gc*dclass3res_V*wstate2+
         (homclass3_V_gc+dclass3res_V|country2),data=dfreg,weights = WEIGHT)

model <- 
  makeLmer(egal_sd~homclass3_V_gc+know_total_gc+dclass3res_V+female_gc+agenum_gc+
             homclass3_V_gc*dclass3res_V*gini_disp +homclass3_V_gc*dclass3res_V*wstate2+
             (homclass3_V_gc+dclass3res_V|country2), 
         fixef=as.numeric(summary(model_re)$coef[,"Estimate"]), 
         VarCorr=lme4::VarCorr(model_re), 
         sigma=as.data.frame(lme4::VarCorr(model_re))$sdcor[11], 
         data=dfreg)

## Simulation for N clusters ----------------------------------------------
n_clusters <- c(50,200) # set the number of clusters
n_sim <- 500 # set the number of simulations

# Simulation: Homogeneity*Working*Gini ------------------------------------
psim_work <- list(NA,NA) # set empty list 
model_psim <- extend(model, along = "country2", n =200)

psim_work_50 <-
    powerCurve(model_psim,
             test=simr::fixed("homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII):gini_disp",method = c("t")),
             breaks = 50,along = "country2",
             nsim=100,
             alpha=.05,
             progress=TRUE,
             seed = 123)

psim_work_200 <-
    powerCurve(model_psim,
             test=simr::fixed("homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII):gini_disp",method = c("t")),
             breaks = 200,along = "country2",
             nsim=20,
             alpha=.05,
             progress=TRUE,
             seed = 123)

model_psim <- extend(model, along = "country2", n =200)
  psim_work[[2]] <-
    powerSim(model_psim,
             test=simr::fixed("homclass3_V_gc:dclass3res_VWorking Class (V+VI+VII):gini_disp",method = c("t")),
             nsim=n_sim,
             alpha=.05,
             progress=TRUE)

psim_work
save(psim_work,file = "psim_xlvl_working.RData")

# Simulation: Homogeneity*Intermediate*Gini -------------------------------
psim_inte <- list() # set empty list 
for (i in n_clusters) {
  model_psim <- extend(model, along = "country2", n =i)
  psim_inte[[i]] <-powerSim(model_psim, 
                            test=simr::fixed("homclass3_V_gc:dclass3res_VIntermediate class (III+IV):gini_disp",method = c("t")),
                            nsim=n_sim,
                            alpha=.05,
                            progress=TRUE)
}
psim_inte
save(psim_inte,file = "psim_xlvl_middle.RData")

```